### 2024-01-31 ms.qmd
First steps

---
docx: 
bibliography: "./bib/refs.bib"
csl: "./bib/biology-letters.csl"
reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

## Introduction
Behavioural flexibility, **the ability to adapt behaviour to changing environments** (**REFS**), is considered crucial for invasion success [@sol2002behavioural]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats (**REFS**) as behaving flexibly enhances their capacity to exploit new resources and face new challenges (**REFS**).

adaptive flexibility hypothesis says that flexibility may be relevant in early and late stages of invassion (Wright et al 2010). But at the same time behavioural flexibility  it is related to relative brain size (Sol et al 2002, 2008), that may be more fixed than we thought (controversy with Wright et al??). In contrast, other hyoptheses predict this is constant, or there is individual variation in beh flexibility (Wright et al 2010). Since some aspects of early environment can play a prominent role in shaping neural substrate, we hypothesize that beh flexibility may be affected by early environment. Also, since beh flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early developmental constraints. CORT can be essential in this process as it is associated to vertebrates response to environmental changes, and it is known to impact brain development (Zhu et al 2004). Also, temperature may be relevant, since it affects brain development in ectotherms, it can be one of the main environmental changes to face by animals, andvertebrates reponse to changes in temeprature is also associated with the production of CORT (Crino et al 2023).

Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context. To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe (**REFS**). Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_ (**REFS**), and are likely key for its invassive success (**REFS**). Our goal in this study was to investigate if behavioural flexibility is affected by the prenatal environment and the species specific response to it. Particularly, we were interested in the interactive effects of prenatal temperature and Corticosterone (CORT) - the main type of GC in reptiles [@Crino_2023] - because CORT is usually associated to vertebrates response to changes in temeprature [@Crino_2023], and both factors are known to impact brain development [@coomber_independent_1997; @zhu_prenatal_2004]. We hypothesized that, since behavioural flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early

## Methods

#### Experimental design
We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.  
To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A. domestica*) as a reward (see @fig-Methods). The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B). In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase. The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top. Lizards were tested once daily between 11 to 12 am, when lizards are active. The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days. We recorded and analysed manually all trials form the learning phases using a standard video player (IINA). PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses. We used as the main variable whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. Here, we analysed the data of the reversal as a measure of behavioural flexibility. As such, we only included in our analyses those individuals who passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.

```{r, fig-Methods}
#| label: fig-Methods
#| fig.cap: "Experimental design of early environment manipulation (**A**) and learning tasks (**B**). Stages 1-3 indicate the different phases of the habituation process. In the associative and reversal tasks, white lids show the ramps where the food reward was not attainable."

knitr::include_graphics("./Others/BEHFLEX_FIG_1.svg")

```

#### Statistical analyses

We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
**As such, we run a total of different 8 models employing Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). Each model consisted of four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations.'Choice', i.e. whether the individual chose correct (1) or not (0) was used as the response variable. The fixed effects of the model included a triple interaction between: trial ('Associative trial' or 'Reversal trial') as a numeric variable, and hormone treatmet ('CORT' versus 'Control') and the temperature at which eggs were incubated ('Cold' versus 'Hot') as factors. For the random effects, we employed lizard identity as a random intercept, and as a random slope we included the variable trial ('Associative' or 'Reversal') within each level of lizard identity. We used the resulting posterior of these models to evaluate learning differences between treatments within and between species and colour assigned. More specifically, we calculated learning slopes by using the estimates of the trial variable per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between 'Treatments' (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.**

## Results
```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
specie <- c("delicata", "guichenoti")
groups <- c("Red", "Blue")
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(i in 1:length(specie)){
  for(j in 1:length(groups)){
    for(k in 1:length(hormone)){
      for(l in 1:length(temperature)){
        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
      }
    }
  }
}
```

Originally, we started with 96 lizards, 48 per species and 12 per treatment per species. However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards. Final sample sizes per treatment and species are disclosed on @tbl-data.

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
source(here("R", "func.R"))
# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
## A.1) L. delicata
deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
deli_asso_blue <- fit_m("asso", "deli", "blue", refit = FALSE)
write.csv(deli_asso_blue, file= "./output/Checking/deli_asso_blue.csv")
## A.2) L. guichenoti
guich_asso_red <- fit_m("asso", "guich", "red", refit = FALSE)
write.csv(guich_asso_red, file= "./output/Checking/guich_asso_red.csv")
guich_asso_blue <- fit_m("asso", "guich", "blue", refit = FALSE)
write.csv(guich_asso_blue, file= "./output/Checking/guich_asso_blue.csv")
# B) Reversal task. Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
## B.1) L. delicata
deli_rev_red <- fit_m("rev", "deli", "red", refit = FALSE)
write.csv(deli_rev_red, file= "./output/Checking/deli_rev_red.csv")
deli_rev_blue <- fit_m("rev", "deli", "blue", refit = FALSE)
write.csv(deli_rev_blue, file= "./output/Checking/deli_rev_blue.csv")
## B.2) L. guichenoti
guich_rev_red <- fit_m("rev", "guich", "red", refit = FALSE)
write.csv(guich_rev_red, file= "./output/Checking/guich_rev_red.csv")
guich_rev_blue <- fit_m("rev", "guich", "blue", refit = FALSE)
write.csv(guich_rev_blue, file= "./output/Checking/guich_rev_blue.csv")
```

```{r, resultsRev}
# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
## 1) L. delicata
### Group = red
drr_CORTCold <- deli_rev_red$b_trial_reversal
drr_ControlCold <- (deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$b_trial_reversal)
drr_CORTHot <- (deli_rev_red$'b_trial_reversal:tempHot' + deli_rev_red$b_trial_reversal)
drr_ControlHot <- (deli_rev_red$'b_trial_reversal:cortControl:tempHot' + deli_rev_red$b_trial_reversal + deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$'b_trial_reversal:tempHot')
### Group = blue
drb_CORTCold <- deli_rev_blue$b_trial_reversal
drb_ControlCold <- (deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$b_trial_reversal)
drb_CORTHot <- (deli_rev_blue$'b_trial_reversal:tempHot' + deli_rev_blue$b_trial_reversal)
drb_ControlHot <- (deli_rev_blue$'b_trial_reversal:cortControl:tempHot' + deli_rev_blue$b_trial_reversal + deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$'b_trial_reversal:tempHot')
## 2) L. guichenoti
### Group = red
grr_CORTCold <- guich_rev_red$b_trial_reversal
grr_ControlCold <- (guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$b_trial_reversal)
grr_CORTHot <- (guich_rev_red$'b_trial_reversal:tempHot' + guich_rev_red$b_trial_reversal)
grr_ControlHot <- (guich_rev_red$'b_trial_reversal:cortControl:tempHot' + guich_rev_red$b_trial_reversal + guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$'b_trial_reversal:tempHot')
### Group = blue
grb_CORTCold <- guich_rev_blue$b_trial_reversal
grb_ControlCold <- (guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$b_trial_reversal)
grb_CORTHot <- (guich_rev_blue$'b_trial_reversal:tempHot' + guich_rev_blue$b_trial_reversal)
grb_ControlHot <- (guich_rev_blue$'b_trial_reversal:cortControl:tempHot' + guich_rev_blue$b_trial_reversal + guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$'b_trial_reversal:tempHot')
```

```{r, tbl-data}
#| label: tbl-data
#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group. Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero. In bold, those values that are significant (p-value <0.05)"
source(here("R", "func.R"))
#
############################## CREATING BIG DF FOR TABLE ##############################
# Building the vectors for titles of rows and columns
specie <- c("L. delicata", "L. guichenoti")
groups <- c("Red", "Blue")
test <- c("Associative", "Reversal")
treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
values <- c("Mean", "95% CI", "p-value")
# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors. p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
#First get estimates for both tasks
estimates_asso <- list(
  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
)
#
estimates_rev <- list(
  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
)
# Then get the mean, co.intervals(95%), and p-values
asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
#
rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
#
# Building the df for the associative
asso_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(asso_mean, each = 1),
  CI = rep(asso_intervals, each = 1),
  PValue = rep(asso_pvalue, each = 1),
  Task = rep("Associative", length(asso_mean))
)
# Building the df for the reversal
rev_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(rev_mean, each = 1),
  CI = rep(rev_intervals, each = 1),
  PValue = rep(rev_pvalue, each = 1),
  Task = rep("Reversal", length(rev_mean))
)
# Joining both dfs
table_data <- rbind(asso_df, rev_df)
table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
#
############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
# Make n_list into a df
n_df <- as.data.frame(do.call(rbind, n_list)) %>%
  rename("n" = V1) %>%
  rownames_to_column("model") %>%
  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
  unite("Treatment", c("cort", "temp"), sep = "-") %>%
  mutate(Specie = factor(Specie,
                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
        Treatment = factor(Treatment,
                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
new_table_data <- merge(table_data, n_df) %>%
  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
  mutate(Specie = factor(Specie,
                  levels = c("L. delicata", "L. guichenoti")),
        Group = factor(Group,
                  levels = c("Red", "Blue")),
        Treatment = factor(Treatment, 
                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
  mutate(Treatment = paste0(Treatment, ")"))
write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
#
############################## MAKING THE TABLE ##############################
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 fint.size = 10)
# Split the table_data df by task
real_table <- flextable(new_table_data) %>%
    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
    set_table_properties(width = 1) %>%
    align(align="center", part="all") %>% 
    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
    set_header_labels(Mean_Associative = "Mean",
                      `95% CI_Associative` = "95% CI",
                      `p-value_Associative` = "p-value",
                      Mean_Reversal = "Mean",
                      `95% CI_Reversal` = "95% CI",
                      `p-value_Reversal` = "p-value") %>%
    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
    autofit() 
real_table
```

In the reversal tesk, we decided to pool both groups 'Blue' and 'Red' to make the between-treatment comparisons for two reasons. First, we did not find any significant differences caused group on the learning slopes (_L. delicata_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`, p = `r format_dec(pmcmc((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`; _L. guichenoti_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`, p = `r format_dec(pmcmc((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`). Second, we were expecting an acquired bias towards the colour assigned in the associative task as part of the design. Nonetheless, we did not find any significant differences between treatments in _L. delicata_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`) (see @Figdeli) or _L. guichenoti_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`) (see @Figguich). We did not find any effect of incubation temperature (_L. delicata_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`; _L. guichenoti_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`) or the hormone (_L. delicata_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`;_L. guichenoti_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`) when groups were pooled. Finally, we did not find any significant differences when we compared the estimated slopes between species for the reversal task (_L. delicata_ - _L. guichenoti_ = `r format_dec(mean((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`).

## Discussion

## References
<div id="refs"></div>

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

# Suplementary Material

#### Colour preference in the Associative task
To test if lizards were biased towards the assigned colour as our preliminary analyses suggested, we employed the intercepts  from our posterior distributions. We first estimated the predicted probability of choosing the correct feeder first in the first trial, by using the formula: 
$$\text{Probability} = \frac{1}{1 + e^{-Intercept}}$$

Second, we tested the hypothesis that the estimated probability was higher than 0.33 (the probability expected by chance of choosing the correct feeder) using pmcmc. If the estimated probability is above 0.33. we consider it as an indication that there was a preference towards that colour that could be affecting learning slopes. The results per treatment are summarized in @tbl-bias. 
```{r, tbl-bias}
#| label: tbl-bias
#| tbl-cap: "Probability of choosing the correct feeder in the first trial when the correct feeder was blue (Prob Blue) or red (Prob Red) for each species and each treatment. p-values were obtained using pmcmc and test the hypothesis that the probability is >0.33. In bold, those values that are significant (p-value <0.05)"
source(here("R", "func.R"))
# First we estimate the probability of choosing right in the first trial using the intercepts from the posteriors
## 1) L. delicata
### Group = red
probright_drCORTCold <- 1/(1+exp(-deli_asso_red$b_Intercept))
probright_drControlCold <- 1/(1+exp(-(deli_asso_red$b_cortControl + deli_asso_red$b_Intercept)))
probright_drCORTHot <- 1/(1+exp(-(deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
probright_drControlHot <- 1/(1+exp(-(deli_asso_red$'b_cortControl:tempHot' + deli_asso_red$b_cortControl + deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
### Group = blue
probright_dbCORTCold <- 1/(1+exp(-(deli_asso_blue$b_Intercept)))
probright_dbControlCold <- 1/(1+exp(-(deli_asso_blue$b_cortControl + deli_asso_blue$b_Intercept)))
probright_dbCORTHot <- 1/(1+exp(-(deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
probright_dbControlHot <- 1/(1+exp(-(deli_asso_blue$'b_cortControl:tempHot' + deli_asso_blue$b_cortControl + deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
## 2) L. guichenoti
### Group = red
probright_grCORTCold <- 1/(1+exp(-guich_asso_red$b_Intercept))
probright_grControlCold <- 1/(1+exp(-(guich_asso_red$b_cortControl + guich_asso_red$b_Intercept)))
probright_grCORTHot <- 1/(1+exp(-(guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
probright_grControlHot <- 1/(1+exp(-(guich_asso_red$'b_cortControl:tempHot' + guich_asso_red$b_cortControl + guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
### Group = blue
probright_gbCORTCold <- 1/(1+exp(-guich_asso_blue$b_Intercept))
probright_gbControlCold <- 1/(1+exp(-(guich_asso_blue$b_cortControl + guich_asso_blue$b_Intercept)))
probright_gbCORTHot <- 1/(1+exp(-(guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
probright_gbControlHot <- 1/(1+exp(-(guich_asso_blue$'b_cortControl:tempHot' + guich_asso_blue$b_cortControl + guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
# Second we build a df with all the mean probabilities of choosing right in the first trial for each treatment and species
prob_df <- data.frame(
  probright_drCORTCold_t = format_dec(mean(probright_drCORTCold), 3),
  probright_drControlCold_t = format_dec(mean(probright_drControlCold), 3),
  probright_drCORTHot_t = format_dec(mean(probright_drCORTHot), 3),
  probright_drControlHot_t = format_dec(mean(probright_drControlHot), 3),
  probright_dbCORTCold_t = format_dec(mean(probright_dbCORTCold), 3),
  probright_dbControlCold_t = format_dec(mean(probright_dbControlCold), 3),
  probright_dbCORTHot_t = format_dec(mean(probright_dbCORTHot), 3),
  probright_dbControlHot_t = format_dec(mean(probright_dbControlHot), 3),
  probright_grCORTCold_t = format_dec(mean(probright_grCORTCold), 3),
  probright_grControlCold_t = format_dec(mean(probright_grControlCold), 3),
  probright_grCORTHot_t = format_dec(mean(probright_grCORTHot), 3),
  probright_grControlHot_t = format_dec(mean(probright_grControlHot), 3),
  probright_gbCORTCold_t = format_dec(mean(probright_gbCORTCold), 3),
  probright_gbControlCold_t = format_dec(mean(probright_gbControlCold), 3),
  probright_gbCORTHot_t = format_dec(mean(probright_gbCORTHot), 3),
  probright_gbControlHot_t = format_dec(mean(probright_gbControlHot), 3)) %>%
gather(key = "column", value = "value", .) %>%
rename("prob_choice" = "value") %>%
data.frame()
write.csv(prob_df, file= "./output/Checking/prob_df.csv")
# Then we build a df with the statistical test (one-tailed pmcmc) that the probability is >0.33 (the probability of choosing right by chance)
comp_pval <- data.frame(
  probright_drCORTCold_t = format_dec(pmcmc(probright_drCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drControlCold_t = format_dec(pmcmc(probright_drControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drCORTHot_t = format_dec(pmcmc(probright_drCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_drControlHot_t = format_dec(pmcmc(probright_drControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbCORTCold_t = format_dec(pmcmc(probright_dbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbControlCold_t = format_dec(pmcmc(probright_dbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbCORTHot_t = format_dec(pmcmc(probright_dbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_dbControlHot_t = format_dec(pmcmc(probright_dbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grCORTCold_t = format_dec(pmcmc(probright_grCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grControlCold_t = format_dec(pmcmc(probright_grControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grCORTHot_t = format_dec(pmcmc(probright_grCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_grControlHot_t = format_dec(pmcmc(probright_grControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbCORTCold_t = format_dec(pmcmc(probright_gbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbControlCold_t = format_dec(pmcmc(probright_gbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbCORTHot_t = format_dec(pmcmc(probright_gbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
  probright_gbControlHot_t = format_dec(pmcmc(probright_gbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3)) %>%
gather(key = "column", value = "value", .) %>%
rename("p_value" = "value") %>%
data.frame()
write.csv(comp_pval, file= "./output/Checking/comp_pval.csv")
# Now we merge them and organise them to have the final dt for the table
table_df <- merge(prob_df, comp_pval, by ="column") %>%
  mutate(Specie = gsub(".*_(d|g).*", "\\1",column)) %>%
  mutate(Group = gsub(".*_(dr|db|gr|gb).*", "\\1",column)) %>%
  mutate(Treatment = gsub(".*(CORTCold|ControlCold|CORTHot|ControlHot).*", "\\1",column)) %>% #Last three commands used to get the species, group and treatment from the column names
  mutate(Treatment = factor(Treatment, 
                    levels = c("CORTCold", "ControlCold", "CORTHot", "ControlHot"),
                    labels = c("CORTCold" = "CORT-Cold", "ControlCold" = "Control-Cold", "CORTHot" = "CORT-Hot", "ControlHot" = "Control-Hot"))) %>%
  mutate(Specie = factor(Specie,
                levels = c("d", "g"),
                labels = c("d" = "L. delicata", "g" = "L. guichenoti"))) %>%
  mutate(Group = recode(Group, "dr" = "Red", "db" = "Blue", "gr" = "Red", "gb" = "Blue")) %>%
  select(-column) %>%
  pivot_wider(names_from = Group, values_from = c(prob_choice, p_value)) %>% #To split the df in Red and Blue
  select(Specie, Treatment, prob_choice_Blue, p_value_Blue, prob_choice_Red, p_value_Red) %>%
  arrange(Specie, Treatment) %>% # To arrange the rows the way I wantdata.frame()
data.frame()
write.csv(table_df, file= "./output/Checking/table_df.csv")
#################
colour_table <- flextable(table_df) %>%
  bold(~ `p_value_Blue` < 0.05, ~ `p_value_Blue` + prob_choice_Blue) %>%
  bold(~ `p_value_Red` < 0.05, ~ `p_value_Red` + prob_choice_Red) %>%
  set_table_properties(width = 1) %>%
  align(align="center", part="all") %>% 
  set_header_labels(prob_choice_Blue = "Prob Blue", 
                    p_value_Blue = "p-value Blue", 
                    prob_choice_Red = "Prob Red", 
                    p_value_Red = "p-value Red") %>%
  italic(j = 1, italic = TRUE, part = "body") %>% # To have names of species in italics
  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  hline(i = c(4), j = c(1:6), part = "body") %>% # To make some horizontal lines
  vline(j = c(2,4), part = "all") %>% # To make some vertical lines on body
autofit()
colour_table
```

On average, we found that, for both species, the proportion of correct choices in the first trial was significantly above chance when the correct feeder was blue (_L. delicata_: mean Prob choice = `r format_dec(mean((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L. guichenoti_: mean Prob choice = `r format_dec(mean((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`), but not when it was red (_L. delicata_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L. guichenoti_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`).

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Husbandry  
*Breeding colony* -- We tested juveniles coming from a breeding colony established in the lab since 2019. There is a total of 270 and 180 adults of *L. delicata* and *L. guichenoti* respectively, housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. Enclosures are provided with non-stick matting, shelter, and several small water dishes. Water is given daily, and they are fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle. Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.  
*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.  
*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).  
Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other. These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below). The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*. We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The rooms temperature was set to between 22-24 Celsius.

#### CORT and Temperature manipulation  
To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A). We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle. CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data). Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching. The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.

### Learning task
The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A. domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig. @fig-Methods B, Stage 3). Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away. We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials. This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.  
In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig. @fig-Methods B). As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp. The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids. The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results. One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. Lizards were tested in this task once a day for 35 days. After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B). Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward. This test was done once a day for 40 days.  
We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active. Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual. Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not. We considered a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. We excluded from our analyses those individuals with more than 15 trials failed (i.e. they did not make a choice), and we considered the first trial to be the first one where the individual made a choice. We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

#### Checking the models plots  
```{r, residuals}
# Chunk for calling all the models and making the residuals
# Associative task
## L. delicata
### Red
mod_dar <- readRDS(here("output/models/asso_deli_red.rds"))
resid_dar <- residuals(mod_dar)
### Blue
mod_dab <- readRDS(here("output/models/asso_deli_blue.rds"))
resid_dab <- residuals(mod_dab)
## L. guichenoti
### Red
mod_gar <- readRDS(here("output/models/asso_guich_red.rds"))
resid_gar <- residuals(mod_gar)
### Blue
mod_gab <- readRDS(here("output/models/asso_guich_blue.rds"))
resid_gab <- residuals(mod_gab)
# Reversal task
## L. delicata
### Red
mod_drr <- readRDS(here("output/models/rev_deli_red.rds"))
resid_drr <- residuals(mod_drr)
### Blue
mod_drb <- readRDS(here("output/models/rev_deli_blue.rds"))
resid_drb <- residuals(mod_drb)
## L. guichenoti
### Red
mod_grr <- readRDS(here("output/models/rev_guich_red.rds"))
resid_grr <- residuals(mod_grr)
### Blue
mod_grb <- readRDS(here("output/models/rev_guich_blue.rds"))
resid_grb <- residuals(mod_grb)
```

Model formula for the reversal task is:  
Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)  
Plots for the different models of the associative task:  
1.- _L. delicata_  
  1.a.- Red  
```{r , plotmod_drr, out.width="70%", fig.align="center"}
plot(mod_drr)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

  1.b.- Blue  
```{r , plotmod_drb, out.width="70%", fig.align="center"}  
plot(mod_drb)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

2.- _L. guichenoti_  
  2.a.- Red  
```{r , plotmod_grr, out.width="70%", fig.align="center"}
plot(mod_grr)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

  2.b.- Blue  
```{r , plotmod_grb, out.width="70%", fig.align="center"}
plot(mod_grb)
```

### 2024-01-31 R/data_process.R
Update data_process.R

@@ -1,12 +1,20 @@
 ####################################
 # 1_data_process
 ####################################
-
-# Packages
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, forcats)
-
-# Load data
-data  <-  read.csv("./data/Learning.csv")
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
+# A) We extract the lizards we are using from the main data_base
+meta_data <- read.csv("./data/lizard_database.csv") %>%
+  filter(lizard_id %in% c()) %>%
+data.frame("LD3158_23", "LD3177_23", "LD3104_23", "LD3100_23", "LD3027_23", "LD3196_23", "LD3025_23", "LD2911_23",
+"LD2930_23", "LD2953_23", "LD2904_23", "LD2981_23", "LD3078_23", "LD3102_23", "LD2997_23", "LD3090_23", "LD3098_23",
+"LD3041_23", "LD2875_23", "LD2950_23", "LD3023_23", "LD2993_23", "LD2958_23", "LD3031_23", "LD3065_23", "LD3038_23",
+"LD3198_23", "LD2851_23", "LD3193_23", "LD2952_23", "LD2995_23", "LD3033_23", "LD2928_23", "LD3081_23", "LD2982_23",
+"LD3103_23", "LD2924_23", "LD3195_23", "LD3028_23", "LD3176_23", "LD3099_23", "LD3079_23", "LD3159_23", "LD3082_23",
+"LD3093_23", "LD2983_23", "LD3039_23", "LD2960_23", "LD3101_23", "LD2926_23", "LD2959_23", "LD3042_23", "LD2908_23",
+"LD3068_23", "LD2925_23", "LD3014_23", "LD3051_23", "LD2984_23", "LD2968_23", "LD2967_23", "LD3009_23", "LD2922_23",
+"LD3034_23", "LD3095_23", "LD2971_23", "LD3191_23", "LD2873_23", "LD3109_23", "LD3178_23", "LD3188_23", "LD3187_23",
+"LD3154_23", "LD3157_23", "LD3179_23", "LD3084_23", "LD3201_23")
+write.csv(meta_data, "./data/spaL.csv") 
 
 # Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
 data_asso <- data %>%

### 2024-01-31 R/data_process.R
Updates

@@ -4,16 +4,16 @@
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
 # A) We extract the lizards we are using from the main data_base
 meta_data <- read.csv("./data/lizard_database.csv") %>%
-  filter(lizard_id %in% c()) %>%
-data.frame("LD3158_23", "LD3177_23", "LD3104_23", "LD3100_23", "LD3027_23", "LD3196_23", "LD3025_23", "LD2911_23",
+  filter(lizard_id %in% c("LD3158_23", "LD3177_23", "LD3104_23", "LD3100_23", "LD3027_23", "LD3196_23", "LD3025_23",
 "LD2930_23", "LD2953_23", "LD2904_23", "LD2981_23", "LD3078_23", "LD3102_23", "LD2997_23", "LD3090_23", "LD3098_23",
 "LD3041_23", "LD2875_23", "LD2950_23", "LD3023_23", "LD2993_23", "LD2958_23", "LD3031_23", "LD3065_23", "LD3038_23",
 "LD3198_23", "LD2851_23", "LD3193_23", "LD2952_23", "LD2995_23", "LD3033_23", "LD2928_23", "LD3081_23", "LD2982_23",
 "LD3103_23", "LD2924_23", "LD3195_23", "LD3028_23", "LD3176_23", "LD3099_23", "LD3079_23", "LD3159_23", "LD3082_23",
 "LD3093_23", "LD2983_23", "LD3039_23", "LD2960_23", "LD3101_23", "LD2926_23", "LD2959_23", "LD3042_23", "LD2908_23",
-"LD3068_23", "LD2925_23", "LD3014_23", "LD3051_23", "LD2984_23", "LD2968_23", "LD2967_23", "LD3009_23", "LD2922_23",
+"LD3086_23", "LD2925_23", "LD3014_23", "LD3051_23", "LD2984_23", "LD2968_23", "LD2967_23", "LD3009_23", "LD2922_23",
 "LD3034_23", "LD3095_23", "LD2971_23", "LD3191_23", "LD2873_23", "LD3109_23", "LD3178_23", "LD3188_23", "LD3187_23",
-"LD3154_23", "LD3157_23", "LD3179_23", "LD3084_23", "LD3201_23")
+"LD3154_23", "LD3157_23", "LD3179_23", "LD3084_23", "LD2911_23", "LD3201_23")) %>%
+data.frame()
[…]

### 2024-07-15 ms.qmd
Updates

@@
 ## Introduction
-Behavioural flexibility, **the ability to adapt behaviour to changing environments** (**REFS**), is considered crucial for invasion success [@sol2002behavioural].
-In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats (**REFS**) as behaving flexibly enhances their capacity to exploit new resources and face new challenges (**REFS**).
-
-adaptive flexibility hypothesis says that flexibility may be relevant in early and late stages of invassion (Wright et al 2010).
-But at the same time behavioural flexibility  it is related to relative brain size (Sol et al 2002, 2008), that may be more fixed than we thought (controversy with Wright et al??).
-In contrast, other hyoptheses predict this is constant, or there is individual variation in beh flexibility (Wright et al 2010).
-Since some aspects of early environment can play a prominent role in shaping neural substrate, we hypothesize that beh flexibility may be affected by early environment.
-Also, since beh flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations.
-In other words, there may be some mechanisms in invassive species to avoid early developmental constraints.
-CORT can be essential in this process as it is associated to vertebrates response to environmental changes, and it is known to impact brain development (Zhu et al 2004).
-Also, temperature may be relevant, since it affects brain development in ectotherms, it can be one of the main environmental changes to face by animals, andvertebrates reponse to changes in temeprature is also associated with the production of CORT (Crino et al 2023).
-
-Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context.
-To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L.
-guichenoti_).
-These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L.
-delicata_ has been succesful in colonising different areas around the globe (**REFS**).
-Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L.
-delicata_ (**REFS**), and are likely key for its invassive success (**REFS**).
-Our goal in this study was to investigate if behavioural flexibility is affected by the prenatal environment and the species specific response to it.
-Particularly, we were interested in the interactive effects of prenatal temperature and Corticosterone (CORT) - the main type of GC in reptiles [@Crino_2023] - because CORT is usually associated to vertebrates response to changes in temeprature [@Crino_2023], and both factors are known to impact brain development [@coomber_independent_1997; @zhu_prenatal_2004].
-We hypothesized that, since behavioural flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations.
-In other words, there may be some mechanisms in invassive species to avoid early
 
@@
 
-#### Experimental design
-We employed juveniles of both species, _L.
-delicata_ and _L.
-guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material).
-To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.
-To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A.
-domestica*) as a reward (see @fig-Methods).
-The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B).
-In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase.
-The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision].
-To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal.
-In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association.
-To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top.
-Lizards were tested once daily between 11 to 12 am, when lizards are active.
-The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days.
-We recorded and analysed manually all trials form the learning phases using a standard video player (IINA).
-PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses.
-We used as the main variable whether the animal chose the correct ramp first or not.
-We considered that a choice was made if the head of the lizard was inside the well of one of the ramps.
-We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses.
-Here, we analysed the data of the reversal as a measure of behavioural flexibility.
-As such, we only included in our analyses those individuals who passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.
+#### Husbandry  
+*Breeding colony* -- The study was conducted with juveniles of the Delicate skink (_Lamprophpolis delicata_) from a breeding colony kept in the lab since 2019.
+The colony is composed by a total of 270 adults housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
+Enclosures are provided with non-stick matting, shelter, and several small water dishes.
+Water is given daily, and they are fed approx.
+40 mid-size crickets (*Acheta domestica*) per enclosure three days a week.
+Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
+To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
+Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.
+
+*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above).
+We checked for the presence of eggs in the boxes three days a week.
+After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
+Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.
+
+*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings.
+After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
+We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish.
+During this period, they were sprayed water every day and received 3-6 small *A.
+domestica* crickets three times a week.
+All care otherwise follows similar protocols to adults (see above).
+Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation.
+The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other.
+These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below).
+The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
+During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
+We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
+The rooms temperature was set to between 22-24 Celsius.
+
+#### CORT and Temperature manipulation  
+To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
+We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle.
+CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data).
+Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching.
+The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
+
+#### Learning
+
+```{r, fig-Methods}
 
@@
 #| label: fig-Methods
-#| fig.cap: "Experimental design of early environment manipulation (**A**) and learning tasks (**B**).
-Stages 1-3 indicate the different phases of the habituation process.
-In the associative and reversal tasks, white lids show the ramps where the food reward was not attainable."
+#| fig.cap: "Panel **A** shows the experimental design of the study.
+On top, the treatments applied to the eggs.
+On the bottom the learning task and the brain region extracted together with the physiological analyses performed.
+In panel **B** the details and measurements of the spatial learning maze."
 
-knitr::include_graphics("./Others/BEHFLEX_FIG_1.svg")
+knitr::include_graphics("./Others/SPAL_METH.svg")
 

### 2024-07-15 R/extract.R

@@ -1,0 +1,86 @@
+# extraction
+# extracting data
+
+# clear workspace
+rm(list = ls())
+
+# load libraries
+library(tidyverse)
+library(here)
+library(brms)
+library(rstan)
+
+# set working directory to top level
+setwd(here())
+
+# source custom functions
+source("./r/func.R")
+
+# make a path to your data
+path <- "./data/"
+
+# make an empty list to hold the data
+data <- list()
+
+# make paths to each file
+paths <- list.files(path = path,
+                    pattern = "\\.fcs",
+                    full.names = T,
+                    recursive = T)
+
+# for each file, read in a simple fcs
+for(i in 1:length(paths)){
+  data[[i]] <- paths[i] %>%
+    read.fcs() %>%
+    data.fcs() %>%
+    process.fcs() %>%
+    summary.fcs()
+  names(data)[[i]] <- paths[i]
+}
+
+# prepare data for compensation
+df <- plyr::ldply(data) %>%
+  # tidy up .id names
+  mutate(.id = str_remove(.id, path),
+         .id = str_remove(.id, "/"),
+         .id = str_remove(.id, "\\.fcs$")) %>%
+  separate(col = .id,
+           into = c("experiment", "plate", "sample"),
+           sep = "/") %>%
+  separate(col = sample,
+           into = c("specimen", "spec.no", "tube", "tube.no", ".no"),
+           sep = "_") %>%
+  # tidy up the acquisition number
+  mutate(.no = as.numeric(.no),
+         .no = factor(.no))
+
+# r crashes if you try to do too much...
+# i'll break this into chunks for now
+# starting with only the area values from the medial cortex experiment
+df.mc <- df %>%
+  dplyr::filter(experiment == "medial cortex" &
+                  param == "a")
+comp.mc <- comp.fcs(df.mc)
+
+# now olfactory bulbs
+df.ob <- df %>%
+  dplyr::filter(experiment == "olfactory bulb" &
+                  param == "a")
+comp.ob <- comp.fcs(df.ob)
+
+# now optical tecta
+df.ot <- df %>%
+  dplyr::filter(experiment == "optical tecta" &
+                  param == "a")
+comp.ot <- comp.fcs(df.ot)
+
+# bind and tidy everything up
+comp <- bind_rows(comp.mc, comp.ob, comp.ot) %>%
+  right_join(df) %>%
+  unite(specimen, c(specimen, spec.no), sep = "_") %>%
[…]

### 2024-07-15 R/extrction_func.R

@@ -1,0 +1,554 @@
+# func.R - a script for custom functions
+
+#-------------------
+# bookkeeping
+#-------------------
+pacman::p_load(tidyverse, here, brms, rstan, flowCore)
+#-------------------
+# statistics
+#-------------------
+#----
+#' @title geo.mean
+#' @param x a vector
+#' @description calculates the geometric mean a robust metric of the central tendency of a population. applicable to values on a logicle scale. is robust to outliers. will only ever be at most equal to the arithmetic mean
+geo.mean <- function(x, na.rm = T){
+  10^base::mean(log.sc(x), na.rm = na.rm)
+}
+#----
+#' @title se
+#' @param x a numeric vector
+#' @description calculates the standard error
+se <- function(x, na.rm = T){
+  stats::sd(x, na.rm = na.rm) / base::sqrt(base::length(x))
+}
+#----
+#' @title r.sd
+#' @param x a numeric vector
+#' @description calculates the robust standard deviation; based on median absolute deviation.
+r.sd <- function(x, na.rm = T) {
+  1.4826 * stats::mad(x, constant = 1.4826, na.rm = na.rm)
+}
+#----
+#' @title r.se
+#' @param x a numeric vector
+#' @description calculates the robust standard error
+r.se <- function(x, na.rm = T) {
+  r.se = r.sd(x, na.rm = na.rm) / sqrt(n())
+}
+#----
+#' @title log.sc
+#' @param x a numeric vector
+#' @description scales a vector to logicle. allows for negative or zero values
+log.sc <- function(x) {
+  base::suppressWarnings(base::ifelse(x <= 0,
+                                      -base::log10(base::abs(x) + 1),
+                                      base::log10(x + 1)))
+}
+#----
+#' @title pmcmc
+#' @param x The vector for the posterior distribution. Note that this will test the null hypothesis that the parameter of interest is significantly different from 0.
+#' @param null A numeric value decsribing what the null hypothesis should be
+#' @param twotail Whether to conduct a one-tailed hypothesis or a two-tailed hypotheses. Default = true indicating a two-tailed test will be done.
+#' @description calculates the pmcmc value for a given posterior distribution (or any combination of posterior distributions) as a metric of significance.
+pmcmc <- function(x, null = 0, twotail = TRUE){
+  if(twotail){
+    2*(1 - max(table(x<=null) / length(x)))
[…]

### 2024-07-15 ms.qmd
Updates

@@
 ## Results
+
+```{r, data_processing}
+#| label: data_processing
+# The result will be the final df with the data for the analysis.
+To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder).
+The final df will be saved in (here("output/databases_clean/data_complete.csv")
+refit = FALSE
+source(here("R", "data_process.R"))
+```
+
 ```{r, sampleSize}
 #| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task.
+# List with the sample sizes from the database (here("output/databases_clean/data_complete.csv") as the sample size per species and group is the same on each task.
 We used function sample (see func.R) to estimate the sample size per treatment and species.

### 2024-07-15 R/data_process.R

@@ -2,29 +2,13 @@
 # 1_data_process
 ####################################
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
-# A) We extract the lizards we are using from the main data_base
-meta_data <- read.csv("./data/lizard_database.csv") %>%
-  filter(lizard_id %in% c("LD3158_23", "LD3177_23", "LD3104_23", "LD3100_23", "LD3027_23", "LD3196_23", "LD3025_23",
-"LD2930_23", "LD2953_23", "LD2904_23", "LD2981_23", "LD3078_23", "LD3102_23", "LD2997_23", "LD3090_23", "LD3098_23",
-"LD3041_23", "LD2875_23", "LD2950_23", "LD3023_23", "LD2993_23", "LD2958_23", "LD3031_23", "LD3065_23", "LD3038_23",
-"LD3198_23", "LD2851_23", "LD3193_23", "LD2952_23", "LD2995_23", "LD3033_23", "LD2928_23", "LD3081_23", "LD2982_23",
-"LD3103_23", "LD2924_23", "LD3195_23", "LD3028_23", "LD3176_23", "LD3099_23", "LD3079_23", "LD3159_23", "LD3082_23",
-"LD3093_23", "LD2983_23", "LD3039_23", "LD2960_23", "LD3101_23", "LD2926_23", "LD2959_23", "LD3042_23", "LD2908_23",
-"LD3086_23", "LD2925_23", "LD3014_23", "LD3051_23", "LD2984_23", "LD2968_23", "LD2967_23", "LD3009_23", "LD2922_23",
-"LD3034_23", "LD3095_23", "LD2971_23", "LD3191_23", "LD2873_23", "LD3109_23", "LD3178_23", "LD3188_23", "LD3187_23",
-"LD3154_23", "LD3157_23", "LD3179_23", "LD3084_23", "LD2911_23", "LD3201_23")) %>%
-data.frame()
-write.csv(meta_data, "./data/spaL.csv") 
-
-# Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
-data_asso <- data %>%
+# A) Processing the learning df
+data <- read.csv(here("./data/Spatial_learn.csv"))
[…]

### 2024-07-15 R/extract.R

@@ -14,7 +14,7 @@ library(rstan)
 setwd(here())
 
 # source custom functions
-source("./r/func.R")
+source("./r/extraction_func.R")
 
 # make a path to your data
 path <- "./data/"

### 2024-07-16 R/data_process.R
Updates

@@ -24,31 +24,32 @@ data_spal <- data %>%
   ungroup() %>% 
 data.frame()
 #### A.2) Extract the slope of each individual
-  #Fit the models only if they have not been fit yet (if refit=TRUE)
-  if(refit){
-    # Formula models
-    formula <- choice ~ 1 + (1 + day|lizard_id) + (1|clutch)
-    # Fit the models
-    # Choice
-    model_choice <- brm(formula,
+#Fit the models only if they have not been fit yet (if refit = TRUE)
+if(refit==TRUE){
+# Fit the models
+  # Choice
+  formula_choice <- choice ~ 1 + (1 + day|lizard_id) + (1|clutch)
+  model_choice <- brm(formula_choice,
                 data = data_spal,
                 family = bernoulli(link = "logit"),
-                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99))
-    # Errors
-    model_errors <- brm(formula,
+                chains = 4, cores = 4, iter = 5000, warmup = 1000, control = list(adapt_delta = 0.99))
+  # Errors
+  formula_errors <- errors ~ 1 + (1 + day|lizard_id) + (1|clutch)
+  model_errors <- brm(formula_errors,
                 data = data_spal,
                 negbinomial(link = "log"),
-                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99))
-    # Write the models to two files
-    saveRDS(model_choice, file = paste0(here("output/models/model_choice.rds")))
-    saveRDS(model_errors, file = paste0(here("output/models/model_errors.rds")))
-
+                chains = 4, cores = 4, iter = 8000, warmup = 2000, prior = ,
+                control = list(adapt_delta = 0.999, max_treedepth = 15))
+  # Write the models to two files
+  saveRDS(model_choice, file = paste0(here("output/models/model_choice.rds")))
+  saveRDS(model_errors, file = paste0(here("output/models/model_errors.rds")))
   } else {
-      # Read the model from a file
-      model_choice <- readRDS(file = paste0(here("output/models/model_choice.rds")))
-      model_errors <- readRDS(file = paste0(here("output/models/model_errors.rds")))
+    # Read the model from a file
+    model_choice <- readRDS(file = paste0(here("output/models/model_choice.rds")))
[…]

### 2024-07-17 ms.qmd
Updates

@@
 ---
+
+```{r}
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
+```
 
@@
 
+```{r, data_metrics}
+#| label: data_metrics
+# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
+data <- read.csv(here("./data/Spatial_learn.csv"))
+#
+# Total number of clutches
+total_clutches <- data %>% 
+  distinct(clutch) %>% 
+  count() %>% 
+  pull(n)
+# Calculate total number of individuals
+total_individuals <- data %>% 
+  distinct(lizard_id) %>% 
+  count() %>% 
+  pull(n)
+# Calculate number of individuals per clutch
+individuals_per_clutch <- data %>%
+  distinct(lizard_id, clutch) %>%
+  group_by(clutch) %>%
+  summarize(individual_count = n())
+```
+
 We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
@@
 We considered statistical significance if p-value < 0.05.**
+
+There are a total of  `r total_individuals`, and `r total_clutches` clutches in the database.
+The average number of individuals per clutch is `r mean(individuals_per_clutch)`, with a minimum of `r min(individuals_per_clutch)` individuals per clutch (`r sum(individuals_per_clutch$individual_count == max(individuals_per_clutch$individual_count))` clutches) and a maximum of `r sum(individuals_per_clutch$individual_count == min(individuals_per_clutch$individual_count))` ( clutches).
 

### 2024-07-17 .vdoc.r

@@ -1,0 +1,513 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+
+```{r, fig-Methods}
+#| label: fig-Methods
+#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
+
+knitr::include_graphics("./Others/SPAL_METH.svg")
+
+#
+#
+#
+#
+#
+#| label: data_metrics
+# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
+data <- read.csv(here("./data/Spatial_learn.csv"))
+#
+# Total number of clutches
+total_clutches <- data %>% 
+  distinct(clutch) %>% 
+  count() %>% 
+  pull(n)
+# Calculate total number of individuals
+total_individuals <- data %>% 
+  distinct(lizard_id) %>% 
+  count() %>% 
+  pull(n)
+# Calculate number of individuals per clutch
+individuals_per_clutch <- data %>%
+  distinct(lizard_id, clutch) %>%
+  group_by(clutch) %>%
+  summarize(individual_count = n())
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: data_processing
+# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
+refit = FALSE
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: sampleSize
[…]

### 2024-07-18 ms.qmd
Updates

@@
 There are a total of  `r total_individuals`, and `r total_clutches` clutches in the database.
-The average number of individuals per clutch is `r mean(individuals_per_clutch)`, with a minimum of `r min(individuals_per_clutch)` individuals per clutch (`r sum(individuals_per_clutch$individual_count == max(individuals_per_clutch$individual_count))` clutches) and a maximum of `r sum(individuals_per_clutch$individual_count == min(individuals_per_clutch$individual_count))` ( clutches).
+The average number of individuals per clutch is `r mean(individuals_per_clutch$individual_count)`, with a minimum of `r min(individuals_per_clutch$individual_count)` individuals per clutch (`r sum(individuals_per_clutch$individual_count == min(individuals_per_clutch$individual_count))` clutches) and a maximum of `r max(individuals_per_clutch$individual_count)` (`r sum(individuals_per_clutch$individual_count == max(individuals_per_clutch$individual_count))` clutches).
+Clutches with only one individual do not contribute to the estimation of within-clutch variability.
+Including these clutches might lead to misleading variance estimates and convergence issues in the hierarchical model.
+Therefore, we initially fit the model without the clutch effect to obtain stable estimates at the individual level.
 

### 2024-07-18 R/data_process.R

@@ -17,29 +17,26 @@ data_spal <- data %>%
     mutate(cort = factor(cort,
       levels = c("B", "A"),
       labels = c("B" = "CORT", "A" = "Control"))) %>%
-    mutate( # Standarize data by trial (i.e. make the first trial where each individual participated their trial 1)
-      first_non_na = min(which(!is.na(choice))),
-      day = ifelse(!is.na(first_non_na),choice - first_non_na + 1, day))%>% 
-      filter(day >= 1) %>%
   ungroup() %>% 
 data.frame()
+#
+#
 #### A.2) Extract the slope of each individual
-#Fit the models only if they have not been fit yet (if refit = TRUE)
+#
+## Fit the models only if they have not been fit yet (if refit = TRUE)
 if(refit==TRUE){
-# Fit the models
   # Choice
-  formula_choice <- choice ~ 1 + (1 + day|lizard_id) + (1|clutch)
-  model_choice <- brm(formula_choice,
+  model_choice <- brm(choice ~ 1 + (1 + day|lizard_id) + (1|clutch),
                 data = data_spal,
                 family = bernoulli(link = "logit"),
-                chains = 4, cores = 4, iter = 5000, warmup = 1000, control = list(adapt_delta = 0.99))
+                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
   # Errors
-  formula_errors <- errors ~ 1 + (1 + day|lizard_id) + (1|clutch)
-  model_errors <- brm(formula_errors,
+  model_errors <- brm(errors ~ 1 + (1 + day|lizard_id) + (1|clutch),
                 data = data_spal,
                 negbinomial(link = "log"),
-                chains = 4, cores = 4, iter = 8000, warmup = 2000, prior = ,
-                control = list(adapt_delta = 0.999, max_treedepth = 15))
+                chains = 4, cores = 4, iter = 8000, warmup = 2000,
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
   # Write the models to two files
   saveRDS(model_choice, file = paste0(here("output/models/model_choice.rds")))
   saveRDS(model_errors, file = paste0(here("output/models/model_errors.rds")))
@@ -48,8 +45,68 @@ if(refit==TRUE){
     model_choice <- readRDS(file = paste0(here("output/models/model_choice.rds")))
[…]

### 2024-07-22 ms.qmd
Updates

@@
 
-In the reversal tesk, we decided to pool both groups 'Blue' and 'Red' to make the between-treatment comparisons for two reasons.
-First, we did not find any significant differences caused group on the learning slopes (_L.
-delicata_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`, p = `r format_dec(pmcmc((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`; _L.
-guichenoti_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`, p = `r format_dec(pmcmc((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`).
-Second, we were expecting an acquired bias towards the colour assigned in the associative task as part of the design.
-Nonetheless, we did not find any significant differences between treatments in _L.
-delicata_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`) (see @Figdeli) or _L.
-guichenoti_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`) (see @Figguich).
-We did not find any effect of incubation temperature (_L.
-delicata_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`; _L.
-guichenoti_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`) or the hormone (_L.
-delicata_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`;_L.
-guichenoti_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`) when groups were pooled.
-Finally, we did not find any significant differences when we compared the estimated slopes between species for the reversal task (_L.
-delicata_ - _L.
-guichenoti_ = `r format_dec(mean((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`).
+```{r, fig-learning}
+#| label: fig-learning
+#| fig.cap: "Results for L.
+delicata (A,B) and L.
+guichenoti (C, D).
+Panels A and C show the predicted probability of choosing the correct feeder first over trials.
+The lines represent the mean predicted probability of choosing the correct feeder first on each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained by using the slope and intercept estimates from the posterior distributions.
+The different colours indicate the different treatments.
+Panels B and D show the distribution of the estimates of slopes per each treatment.
+The x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
+The different colours indicate the different treatments.
+Points and bars represent the mean and standard deviation of the mean of the estimates, respectively."
+#
+#
+####### A) Probability of choosing correct figure
+## A.1) Get the df with the estimated probability of choosing the correct feeder first over trials (using learning_df from data_process.R)
+# Get the df for the intercepts
+source(here("R", "func.R"))
+post_choice_intercept <- tidy_post_df("choice", "intercept")
+# Merge slope and intercept dfs for choice 
+learning_df_prob <- merge(learning_df, post_choice_intercept, by = "lizard_id", all = TRUE)
+# Create a vector with treatments
+treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
+# Create new matrix and new dfs
+treat_matrix <- matrix(NA, nrow = 20, ncol = 40)
+fig_df <- data.frame()
+# Loop through treatments
+for(t in treatments){
+  if(t == "CORT-Cold"){
+    df <- learning_df_prob %>%
+      filter(cort=="CORT", temp=="Cold") %>%
+      data.frame()
+  } else if(t == "Control-Cold"){
+    df <- learning_df_prob %>%
+      filter(cort=="Control", temp=="Cold") %>%
+      data.frame()
+  } else if(t == "CORT-Hot"){
+    df <- learning_df_prob %>%
+      filter(cort=="CORT", temp=="Hot") %>%
+      data.frame()
+  } else if(t == "Control-Hot"){
+    df <- learning_df_prob %>%
+      filter(cort=="Control", temp=="Hot") %>%
+      data.frame()
+  } else {
+    stop("loop wrong")
+  }
+  # Variables selected
+  m <- df$choice_slope__mean
+  u <- df$choice_intercept__mean
+  # Loop per treatment
+  num_individuals <- length(u)
+  for(x in 0:40){
+    for(j in 1:num_individuals){
+      value <- exp(u[j] + m[j] * x) / (1 + exp(u[j] + m[j] * x))
+      treat_matrix[j, x] <- value
+    }
+  }
+  treat_df <- as.data.frame(treat_matrix)
+  colnames(treat_df) <- paste0("X", 1:40)  # Adjust column names
+  treat_df <- gather(treat_df, key = "Trial", value = "Value")  # Reshape data frame
+  treat_df$Treatment <- t
+  fig_df <- rbind(fig_df, treat_df)
+}
+# Transform the database for better use
+fig_choice_df <- fig_df %>%
+  mutate(Trial = gsub("X", "", Trial)) %>%
+  mutate(Trial = as.numeric(Trial)) %>%
+  group_by(Trial, Treatment) %>%
+  summarize(
+    Mean_Predicted_prob = mean(Value),
+    SE_Predicted_prob = sd(Value)/sqrt(length(Value))
+    ) %>%
+  ungroup() %>%
+  mutate(
+    Treatment = factor(Treatment,
+                        levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")),
+    ) %>%
+data.frame()
+#
+## A.2) Make the plot for the probability of choosing the correct feeder first over trials
+fig_prob_choice <- ggplot(df_prob, aes(x = Trial, y = Mean_Predicted_prob, color = Treatment)) +
+  geom_line(linewidth = 1) +
+  scale_color_manual(values = c("CORT-Cold"="#00008B", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +
+  geom_ribbon(aes(ymin = Mean_Predicted_prob - SE_Predicted_prob, ymax = Mean_Predicted_prob + SE_Predicted_prob, fill = Treatment), color = NA, alpha = 0.075) + 
+  scale_fill_manual(values = c("CORT-Cold"="darkblue", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +
+  theme_classic() +
+  labs(y = "Predicted probability of correct choice", x = "Trial") +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "none"
+  )
+#
+## A.3) Make the the plot for the mean slope per treatment using the probability of choosing the correct feeder
+fig_slope_choice <- plot_slopes("choice")
+#
+#
+## A.4) Combine plots of probability and slopes to have the "choice" figure
+fig_results_choice <- plot_grid(fig_prob_choice, fig_slope_choice, nrow = 1)
+ggsave("./output/figures/fig_results_choice.png", plot=fig_results_choice, width = 25, height = 18, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/fig_results.png")
+#
+#
+####### A) Errors figure
+fig_slope_errors <- plot_slopes("errors")
+write.csv(data_fig, file= "./output/Checking/data_fig.csv")
+write.csv(data_plot, file= "./output/Checking/data_plot.csv")
+```
 
@@
 # Suplementary Material
-
-#### Colour preference in the Associative task
-To test if lizards were biased towards the assigned colour as our preliminary analyses suggested, we employed the intercepts  from our posterior distributions.
-We first estimated the predicted probability of choosing the correct feeder first in the first trial, by using the formula:
-$$\text{Probability} = \frac{1}{1 + e^{-Intercept}}$$
-
-Second, we tested the hypothesis that the estimated probability was higher than 0.33 (the probability expected by chance of choosing the correct feeder) using pmcmc.
-If the estimated probability is above 0.33.
-we consider it as an indication that there was a preference towards that colour that could be affecting learning slopes.
-The results per treatment are summarized in @tbl-bias.
-```{r, tbl-bias}
-#| label: tbl-bias
-#| tbl-cap: "Probability of choosing the correct feeder in the first trial when the correct feeder was blue (Prob Blue) or red (Prob Red) for each species and each treatment.
-p-values were obtained using pmcmc and test the hypothesis that the probability is >0.33.
-In bold, those values that are significant (p-value <0.05)"
-source(here("R", "func.R"))
-# First we estimate the probability of choosing right in the first trial using the intercepts from the posteriors
-## 1) L. delicata
-### Group = red
-probright_drCORTCold <- 1/(1+exp(-deli_asso_red$b_Intercept))
-probright_drControlCold <- 1/(1+exp(-(deli_asso_red$b_cortControl + deli_asso_red$b_Intercept)))
-probright_drCORTHot <- 1/(1+exp(-(deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
-probright_drControlHot <- 1/(1+exp(-(deli_asso_red$'b_cortControl:tempHot' + deli_asso_red$b_cortControl + deli_asso_red$b_tempHot + deli_asso_red$b_Intercept)))
-### Group = blue
-probright_dbCORTCold <- 1/(1+exp(-(deli_asso_blue$b_Intercept)))
-probright_dbControlCold <- 1/(1+exp(-(deli_asso_blue$b_cortControl + deli_asso_blue$b_Intercept)))
-probright_dbCORTHot <- 1/(1+exp(-(deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
-probright_dbControlHot <- 1/(1+exp(-(deli_asso_blue$'b_cortControl:tempHot' + deli_asso_blue$b_cortControl + deli_asso_blue$b_tempHot + deli_asso_blue$b_Intercept)))
-## 2) L. guichenoti
-### Group = red
-probright_grCORTCold <- 1/(1+exp(-guich_asso_red$b_Intercept))
-probright_grControlCold <- 1/(1+exp(-(guich_asso_red$b_cortControl + guich_asso_red$b_Intercept)))
-probright_grCORTHot <- 1/(1+exp(-(guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
-probright_grControlHot <- 1/(1+exp(-(guich_asso_red$'b_cortControl:tempHot' + guich_asso_red$b_cortControl + guich_asso_red$b_tempHot + guich_asso_red$b_Intercept)))
-### Group = blue
-probright_gbCORTCold <- 1/(1+exp(-guich_asso_blue$b_Intercept))
-probright_gbControlCold <- 1/(1+exp(-(guich_asso_blue$b_cortControl + guich_asso_blue$b_Intercept)))
-probright_gbCORTHot <- 1/(1+exp(-(guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
-probright_gbControlHot <- 1/(1+exp(-(guich_asso_blue$'b_cortControl:tempHot' + guich_asso_blue$b_cortControl + guich_asso_blue$b_tempHot + guich_asso_blue$b_Intercept)))
-# Second we build a df with all the mean probabilities of choosing right in the first trial for each treatment and species
-prob_df <- data.frame(
-  probright_drCORTCold_t = format_dec(mean(probright_drCORTCold), 3),
-  probright_drControlCold_t = format_dec(mean(probright_drControlCold), 3),
-  probright_drCORTHot_t = format_dec(mean(probright_drCORTHot), 3),
-  probright_drControlHot_t = format_dec(mean(probright_drControlHot), 3),
-  probright_dbCORTCold_t = format_dec(mean(probright_dbCORTCold), 3),
-  probright_dbControlCold_t = format_dec(mean(probright_dbControlCold), 3),
-  probright_dbCORTHot_t = format_dec(mean(probright_dbCORTHot), 3),
-  probright_dbControlHot_t = format_dec(mean(probright_dbControlHot), 3),
-  probright_grCORTCold_t = format_dec(mean(probright_grCORTCold), 3),
-  probright_grControlCold_t = format_dec(mean(probright_grControlCold), 3),
-  probright_grCORTHot_t = format_dec(mean(probright_grCORTHot), 3),
-  probright_grControlHot_t = format_dec(mean(probright_grControlHot), 3),
-  probright_gbCORTCold_t = format_dec(mean(probright_gbCORTCold), 3),
-  probright_gbControlCold_t = format_dec(mean(probright_gbControlCold), 3),
-  probright_gbCORTHot_t = format_dec(mean(probright_gbCORTHot), 3),
-  probright_gbControlHot_t = format_dec(mean(probright_gbControlHot), 3)) %>%
-gather(key = "column", value = "value", .) %>%
-rename("prob_choice" = "value") %>%
-data.frame()
-write.csv(prob_df, file= "./output/Checking/prob_df.csv")
-# Then we build a df with the statistical test (one-tailed pmcmc) that the probability is >0.33 (the probability of choosing right by chance)
-comp_pval <- data.frame(
-  probright_drCORTCold_t = format_dec(pmcmc(probright_drCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drControlCold_t = format_dec(pmcmc(probright_drControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drCORTHot_t = format_dec(pmcmc(probright_drCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_drControlHot_t = format_dec(pmcmc(probright_drControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbCORTCold_t = format_dec(pmcmc(probright_dbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbControlCold_t = format_dec(pmcmc(probright_dbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbCORTHot_t = format_dec(pmcmc(probright_dbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_dbControlHot_t = format_dec(pmcmc(probright_dbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grCORTCold_t = format_dec(pmcmc(probright_grCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grControlCold_t = format_dec(pmcmc(probright_grControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grCORTHot_t = format_dec(pmcmc(probright_grCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_grControlHot_t = format_dec(pmcmc(probright_grControlHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbCORTCold_t = format_dec(pmcmc(probright_gbCORTCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbControlCold_t = format_dec(pmcmc(probright_gbControlCold, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbCORTHot_t = format_dec(pmcmc(probright_gbCORTHot, null = 0.33, twotail = FALSE, dir=">"), 3),
-  probright_gbControlHot_t = format_dec(pmcmc(probright_gbControlHot, null = 0.33, twotail = FALSE, dir=">"), 3)) %>%
-gather(key = "column", value = "value", .) %>%
-rename("p_value" = "value") %>%
-data.frame()
-write.csv(comp_pval, file= "./output/Checking/comp_pval.csv")
-# Now we merge them and organise them to have the final dt for the table
-table_df <- merge(prob_df, comp_pval, by ="column") %>%
-  mutate(Specie = gsub(".*_(d|g).*", "\\1",column)) %>%
-  mutate(Group = gsub(".*_(dr|db|gr|gb).*", "\\1",column)) %>%
-  mutate(Treatment = gsub(".*(CORTCold|ControlCold|CORTHot|ControlHot).*", "\\1",column)) %>% #Last three commands used to get the species, group and treatment from the column names
-  mutate(Treatment = factor(Treatment, 
-                    levels = c("CORTCold", "ControlCold", "CORTHot", "ControlHot"),
-                    labels = c("CORTCold" = "CORT-Cold", "ControlCold" = "Control-Cold", "CORTHot" = "CORT-Hot", "ControlHot" = "Control-Hot"))) %>%
-  mutate(Specie = factor(Specie,
-                levels = c("d", "g"),
-                labels = c("d" = "L. delicata", "g" = "L. guichenoti"))) %>%
-  mutate(Group = recode(Group, "dr" = "Red", "db" = "Blue", "gr" = "Red", "gb" = "Blue")) %>%
-  select(-column) %>%
-  pivot_wider(names_from = Group, values_from = c(prob_choice, p_value)) %>% #To split the df in Red and Blue
-  select(Specie, Treatment, prob_choice_Blue, p_value_Blue, prob_choice_Red, p_value_Red) %>%
-  arrange(Specie, Treatment) %>% # To arrange the rows the way I wantdata.frame()
-data.frame()
-write.csv(table_df, file= "./output/Checking/table_df.csv")
-#################
-colour_table <- flextable(table_df) %>%
-  bold(~ `p_value_Blue` < 0.05, ~ `p_value_Blue` + prob_choice_Blue) %>%
-  bold(~ `p_value_Red` < 0.05, ~ `p_value_Red` + prob_choice_Red) %>%
-  set_table_properties(width = 1) %>%
-  align(align="center", part="all") %>% 
-  set_header_labels(prob_choice_Blue = "Prob Blue", 
-                    p_value_Blue = "p-value Blue", 
-                    prob_choice_Red = "Prob Red", 
-                    p_value_Red = "p-value Red") %>%
-  italic(j = 1, italic = TRUE, part = "body") %>% # To have names of species in italics
-  flextable::compose(i = c(2:4,6:8), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  hline(i = c(4), j = c(1:6), part = "body") %>% # To make some horizontal lines
-  vline(j = c(2,4), part = "all") %>% # To make some vertical lines on body
-autofit()
-colour_table
-```
-
-On average, we found that, for both species, the proportion of correct choices in the first trial was significantly above chance when the correct feeder was blue (_L.
-delicata_: mean Prob choice = `r format_dec(mean((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_dbCORTCold + probright_dbControlCold + probright_dbCORTHot + probright_dbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L.
-guichenoti_: mean Prob choice = `r format_dec(mean((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_gbCORTCold + probright_gbControlCold + probright_gbCORTHot + probright_gbControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`), but not when it was red (_L.
-delicata_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`; _L.
-guichenoti_: mean Prob choice = `r format_dec(mean((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), 3)`, p-value = `r format_dec(pmcmc(((probright_drCORTCold + probright_drControlCold + probright_drCORTHot + probright_drControlHot)/4), null = 0.33, twotail = FALSE, dir=">"), 3)`).
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-#### Husbandry  
-*Breeding colony* -- We tested juveniles coming from a breeding colony established in the lab since 2019.
-There is a total of 270 and 180 adults of *L.
-delicata* and *L.
-guichenoti* respectively, housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-Enclosures are provided with non-stick matting, shelter, and several small water dishes.
-Water is given daily, and they are fed approx.
-40 mid-size crickets (*Acheta domestica*) per enclosure three days a week.
-Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
-Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.
-*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above).
-We checked for the presence of eggs in the boxes three days a week.
-After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.
-*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings.
-After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish.
-During this period, they were sprayed water every day and received 3-6 small *A.
-domestica* crickets three times a week.
-All care otherwise follows similar protocols to adults (see above).
-Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other.
-These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below).
-The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
-We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
-The rooms temperature was set to between 22-24 Celsius.
-
-#### CORT and Temperature manipulation  
-To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle.
-CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data).
-Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching.
-The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
-
-### Learning task
-The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A.
-domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig.
-@fig-Methods B, Stage 3).
-Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away.
-We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials.
-This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.
-In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig.
-@fig-Methods B).
-As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp.
-The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids.
-The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision].
-However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results.
-One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal.
-In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association.
-Lizards were tested in this task once a day for 35 days.
-After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B).
-Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward.
-This test was done once a day for 40 days.
-We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active.
-Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual.
-Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not.
-We considered a choice was made if the head of the lizard was inside the well of one of the ramps.
-We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses.
-We excluded from our analyses those individuals with more than 15 trials failed (i.e.
-they did not make a choice), and we considered the first trial to be the first one where the individual made a choice.
-We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
 

### 2024-07-22 .vdoc.r

@@ -1,513 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-
-```{r, fig-Methods}
-#| label: fig-Methods
-#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
-
-knitr::include_graphics("./Others/SPAL_METH.svg")
-
-#
-#
-#
-#
-#
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: data_processing
-# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
-refit = FALSE
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
[…]

### 2024-07-22 R/data_process.R

@@ -2,7 +2,8 @@
 # 1_data_process
 ####################################
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
-# A) Processing the learning df
+#
+# A) PROCESSING THE LEARNING DF
 data <- read.csv(here("./data/Spatial_learn.csv"))
 #### A.1) Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
 data_spal <- data %>%
@@ -30,7 +31,8 @@ if(refit==TRUE){
                 data = data_spal,
                 family = bernoulli(link = "logit"),
                 chains = 4, cores = 4, iter = 8000, warmup = 2000, 
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
+                control = list(adapt_delta = 0.99, max_treedepth = 12),
+                prior = custom_priors)
   # Errors
   model_errors <- brm(errors ~ 1 + (1 + day|lizard_id) + (1|clutch),
                 data = data_spal,
@@ -53,53 +55,15 @@ posteriors_errors <- as_draws_df(model_errors)
 #
 #### A.3) Modify posteriors df to get the mean slope and intercept of each individual (the latter would be useful for the figures)
 #
+# Getting slopes
+source(here("R", "func.R"))
 ## Choice
-  post_choice <- posteriors_choice %>%
-    select(starts_with("r_lizard_id")) %>%
-    summarise(across(everything(), list(
-      mean = ~ mean(.x, na.rm = TRUE),
-      sd = ~ sd(.x, na.rm = TRUE),
-      se = ~ sd(.x, na.rm = TRUE)/sqrt(length(.x))
-    ))) %>%
-  data.frame()
-  post_choice_b <- post_choice %>%
-    pivot_longer(cols = everything(),
-                names_to = "lizard_id", 
-                values_to = "value") %>% # Extract the relevant columns and reshape them
-    mutate(lizard_id = gsub("r_lizard_id\\.", "", lizard_id),  # Remove prefix
-           lizard_id = gsub("\\.", "_", lizard_id),            # Replace all remaining dots with underscores
-           lizard_id = gsub("_Intercept_", "_Intercept", lizard_id),  
-           lizard_id = gsub("_day_", "_day", lizard_id)) %>% # Remove extra underscore
[…]

### 2024-07-22 R/func.r

@@ -1,22 +1,71 @@
 #################### 
 ####################
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x)
 #
+####################
+####################
+# Function to make the first cleaning for the initial posteriors to estimate individual learning slopes or intercepts
+#' @title tidy_post_df
+#' @param df to select the posteriors to use. Options: choice or errors
+#' @param effect to select whether we want the intercept or the slopes. Options: slope or intercept
+tidy_post_df <- function(df, effect) {
+  if(df == "choice") {
+    
+    data <- posteriors_choice
+    label_df <- "choice_"
+  } else if(df == "errors") {
+    data <- posteriors_errors
+    label_df <- "errors_"
+  } else {
+    stop("df not valid")
+  }
+  if(effect == "slope") {
+    label_effect <- "day"
+    label_df <- paste0(label_df, "slope_")
+    pattern <- "^r_lizard_id\\[\\d+,day\\]$"
+  } else if(effect == "intercept") {
+    label_effect <- "Intercept"
+    label_df <- paste0(label_df, "intercept_")
+    pattern <- "^r_lizard_id\\.\\d+\\.Intercept\\."
+    data <- data %>%
+      select(b_Intercept, matches("^r_lizard.*Intercept]$")) %>%
+      mutate(across(-b_Intercept, ~ . + b_Intercept)) %>%
+    data.frame()
+  } else {
+    stop("effect not valid")
+  }
+  tidy_post_1 <- data %>%
+    select(matches(pattern)) %>%
+    summarise(across(everything(), list(
+      mean = ~ mean(.x, na.rm = TRUE),
+      sd = ~ sd(.x, na.rm = TRUE),
+      se = ~ sd(.x, na.rm = TRUE)/sqrt(length(.x))
+    ))) %>%
+  data.frame()
+  tidy_post_2 <- tidy_post_1 %>%
+    pivot_longer(cols = everything(),
+              names_to = "lizard_id", 
+              values_to = "value") %>% # Extract the relevant columns and reshape them
+    mutate(lizard_id = gsub("r_lizard_id\\.", "", lizard_id),  # Remove prefix
[…]

### 2024-07-23 .vdoc.r
Updates

@@ -1,47 +1,47 @@
----
-docx: 
-bibliography: "./bib/refs.bib"
-csl: "./bib/biology-letters.csl"
-reference-doc: "./bib/tmpl.docx"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-link-citations: true
-crossref:  
-  fig-title: Fig    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Table  # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
-```
-
-## Introduction
-
-
-## Methods  
-
-#### Husbandry  
-*Breeding colony* -- The study was conducted with juveniles of the Delicate skink (_Lamprophpolis delicata_) from a breeding colony kept in the lab since 2019. The colony is composed by a total of 270 adults housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. Enclosures are provided with non-stick matting, shelter, and several small water dishes. Water is given daily, and they are fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle. Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.  
-
[…]

### 2024-07-23 R/data_process.R

@@ -46,7 +46,7 @@ if(refit==TRUE){
     # Read the model from a file
     model_choice <- readRDS(file = paste0(here("output/models/model_choice.rds")))
     model_errors <- readRDS(file = paste0(here("output/models/model_errors.rds")))
-  } 
+} 
 #
 ## Extract posteriors
 posteriors_choice <- as_draws_df(model_choice)
@@ -67,13 +67,42 @@ learning_posteriors <- merge(post_choice_slope, post_errors_slope, by = "lizard_
 #
 #
 #### A.5) Merging initial df with posteriors to get the final learning df with individual learning slopes and treatments
-  mod_spal <- data_spal %>%
-    group_by(lizard_id) %>%
-      filter(day == 1) %>%
-    select(lizard_id, clutch, age, temp, cort) %>% # We modify here the spal df to merge it with the estiamates per invidual
-  data.frame()
+mod_spal <- data_spal %>%
+  group_by(lizard_id) %>%
+    filter(day == 1) %>%
+  select(lizard_id, clutch, age, temp, cort) %>% # We modify here the spal df to merge it with the estiamates per invidual
+data.frame()
 learning_df <- merge(mod_spal, learning_posteriors, by = "lizard_id")
 #
 #
 #
-# B) 
+# B) PROCESSING THE PHYSIOLOGY DF
+#### B.1) Modify the mit_df ("./data/mit_data.csv") and merge with the ()
+mit_data <- read.csv(here("./data/mit_data.csv"))
+mit_df <- mit_data %>%
+  filter(experiment == "medial cortex") %>%
+data.frame()
+euth <- read.csv(here("./data/Euthanasia.csv"))
+physio_df <- merge(mit_df, euth, by = c("plate", "tube"))
+#
+#### B.2) Modify the data base to merge it with learning_df
+mod_physio_df <- physio_df %>%
+  filter(type != "control", ch == "percp.cy5.5"| ch == "apc") %>%
+  mutate(ch = recode(ch, "percp.cy5.5" = "ROS", "apc" = "Mit_density")) %>%
+  select(tube, lizard_id, ch, geo.mean, arith.mean) %>%
+  group_by(lizard_id, ch) %>%
+  summarize(geo.mean = mean(geo.mean, na.rm = TRUE), 
+            arith.mean = mean(arith.mean, na.rm = TRUE),
+            .groups = 'drop') %>%
+  pivot_wider(
[…]

### 2024-07-23 R/func.r

@@ -10,7 +10,6 @@ pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4,
 #' @param effect to select whether we want the intercept or the slopes. Options: slope or intercept
 tidy_post_df <- function(df, effect) {
   if(df == "choice") {
-    
     data <- posteriors_choice
     label_df <- "choice_"
   } else if(df == "errors") {
@@ -187,57 +186,39 @@ format_p <- function(x, n) {
                 paste0("= ", as.character(z))))
   return(tmp)
 }
-###################
-###################
-# Function to create each df for the slopes figure
-#' @title df_plot_slopes
-#' @param type to select the response variable: "choice" or "errors"
-df_plot_slopes <- function(type){
-  data <- learning_df %>%
-    mutate(Treatment = paste(cort, temp, sep = "-")) %>%
-    mutate(Treatment = factor(Treatment,
-      levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot"),
-      labels = c("CORT-Cold" = "CORT-Cold (n=20)",
-                "Control-Cold" = "Control-Cold (n=20)",
-                "CORT-Hot" = "CORT-Hot (n=19)",
-                "Control-Hot" = "Control-Hot (n=20)")
-    )) %>%
-  data.frame()
-  if(type == "choice"){
-    data_fig <- data %>%
-      select(lizard_id, Treatment, choice_slope__mean, choice_slope__sd, choice_slope__se) %>%
-      rename(slope_mean = choice_slope__mean,
-            slope_sd = choice_slope__sd,
-            slope_se = choice_slope__se) %>%
-    data.frame()
-  } else if(type == "errors") {
-    data_fig <- data %>%
-      select(lizard_id, Treatment, errors_slope__mean, errors_slope__sd, errors_slope__se) %>%
-      rename(slope_mean = errors_slope__mean,
-            slope_sd = errors_slope__sd,
-            slope_se = errors_slope__se) %>%
-    data.frame()
-  } else {
-    stop("type not valid")
-  }
-  data_plot <- data_fig %>%
-    group_by(Treatment) %>%
-    summarize(
-      Mean = mean(slope_mean),
-      SD = mean(slope_sd),
-      SE = mean(slope_se)
-    ) %>%
-    ungroup() %>%
-  data.frame()
-  return(data_plot)
-}
[…]

### 2024-07-30 ms.qmd
Updates

@@
 source(here("R", "func.R"))
-# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
-## A.1) L. delicata
-deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
-write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
-deli_asso_blue <- fit_m("asso", "deli", "blue", refit = FALSE)
-write.csv(deli_asso_blue, file= "./output/Checking/deli_asso_blue.csv")
-## A.2) L. guichenoti
-guich_asso_red <- fit_m("asso", "guich", "red", refit = FALSE)
-write.csv(guich_asso_red, file= "./output/Checking/guich_asso_red.csv")
-guich_asso_blue <- fit_m("asso", "guich", "blue", refit = FALSE)
-write.csv(guich_asso_blue, file= "./output/Checking/guich_asso_blue.csv")
-# B) Reversal task. Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
-## B.1) L. delicata
-deli_rev_red <- fit_m("rev", "deli", "red", refit = FALSE)
-write.csv(deli_rev_red, file= "./output/Checking/deli_rev_red.csv")
-deli_rev_blue <- fit_m("rev", "deli", "blue", refit = FALSE)
-write.csv(deli_rev_blue, file= "./output/Checking/deli_rev_blue.csv")
-## B.2) L. guichenoti
-guich_rev_red <- fit_m("rev", "guich", "red", refit = FALSE)
-write.csv(guich_rev_red, file= "./output/Checking/guich_rev_red.csv")
-guich_rev_blue <- fit_m("rev", "guich", "blue", refit = FALSE)
-write.csv(guich_rev_blue, file= "./output/Checking/guich_rev_blue.csv")
+# 
+# df got from data_process.R
+#
+fchoice <- bf(choice_slope_mean + choice_slope_se ~ cort*temp)
+ferrors <- bf(errors_slope_mean + errors_slope_se ~ cort*temp)
+fROS <- bf(ROS ~ Mit_density + cort*temp)
+#
+refit <- FALSE
+if(refit==TRUE){
+  # Choice
+  model_choice_2 <- brm(fchoice + fROS,
+                data = clean_df,
+                family = gaussian,
+                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
+  # Errors
+  model_errors_2 <- brm(ferrors + fROS,
+                data = clean_df,
+                family = gaussian,
+                chains = 4, cores = 4, iter = 8000, warmup = 2000,
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
+  # Write the models to two files
+  saveRDS(model_choice_2, file = paste0(here("output/models/model_choice_2.rds")))
+  saveRDS(model_errors_2, file = paste0(here("output/models/model_errors_2.rds")))
+  } else {
+    # Read the model from a file
+    model_choice_2 <- readRDS(file = paste0(here("output/models/model_choice_2.rds")))
+    model_errors_2 <- readRDS(file = paste0(here("output/models/model_errors_2.rds")))
+} 
+post_gen <- as_draws_df(model_errors_2)
+write.csv(post_gen, file= "./output/Checking/post_gen.csv")
 ```
@@
 ## A.2) Make the plot for the probability of choosing the correct feeder first over trials
-fig_prob_choice <- ggplot(df_prob, aes(x = Trial, y = Mean_Predicted_prob, color = Treatment)) +
+fig_prob_choice <- ggplot(fig_choice_df, aes(x = Trial, y = Mean_Predicted_prob, color = Treatment)) +
   geom_line(linewidth = 1) +
@@
 fig_results_choice <- plot_grid(fig_prob_choice, fig_slope_choice, nrow = 1) 
-ggsave("./output/figures/fig_results_choice.png", plot=fig_results_choice, width = 25, height = 18, units = "cm", dpi = 600)
-knitr::include_graphics("./output/figures/fig_results.png")
+ggsave("./output/figures/fig_results_choice.png", plot=fig_results_choice, width = 25, height = 15, units = "cm", dpi = 600)
 #
 #
-####### A) Errors figure
+#
+####### B) Errors figure
+## B.1) Plot the errors using data_spal
+errors_df <- data_spal %>%
+  mutate(Treatment = paste(cort, temp, sep = "-")) %>%
+    mutate(Treatment = factor(Treatment,
+      levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
+    )) %>%
+  group_by(day, Treatment) %>%
+  summarise(
+    mean_errors = mean(errors, na.rm = TRUE),
+    sd_errors = sd(errors, na.rm = TRUE)
+  ) %>%
+  ungroup() %>%
+  data.frame()
+fig_errors <- ggplot(errors_df, aes(x = day, y = mean_errors, color = Treatment, fill = Treatment)) +
+  geom_smooth(method = "lm", se = TRUE, linewidth = 1, alpha = 0.075) +
+  scale_color_manual(values = c("CORT-Cold" = "#00008B", "Control-Cold" = "#68bde1", "CORT-Hot" = "#b50101", "Control-Hot" = "#fa927d")) +
+  scale_fill_manual(values = c("CORT-Cold" = "darkblue", "Control-Cold" = "#68bde1", "CORT-Hot" = "#b50101", "Control-Hot" = "#fa927d")) +
+  theme_classic() +
+  labs(y = "Mean amount of errors made", x = "Trial") +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "none"
+  )
+## B.2) Make the the plot for the mean slope per treatment using the probability of choosing the correct feeder
 fig_slope_errors <- plot_slopes("errors")
-write.csv(data_fig, file= "./output/Checking/data_fig.csv")
-write.csv(data_plot, file= "./output/Checking/data_plot.csv")
+#
+#
+## B.3) Combine plots of probability and slopes to have the "choice" figure
+fig_results_errors <- plot_grid(fig_errors, fig_slope_errors, nrow = 1) 
+ggsave("./output/figures/fig_results_errors.png", plot=fig_results_errors, width = 25, height = 15, units = "cm", dpi = 600)
+```
+
+```{r, ROS-learning}
+#| label: ROS-learning
+tidy_post_fig_ROS <- post_gen %>%
+  mutate(CORT_Cold = b_ROS_Intercept,
+      Control_Cold = b_ROS_Intercept + b_ROS_cortControl,
+      CORT_Hot = b_ROS_Intercept + b_ROS_tempHot,
+      Control_Hot = b_ROS_Intercept + b_ROS_cortControl + b_ROS_tempHot + `b_ROS_cortControl:tempHot`) %>%
+  select(CORT_Cold, Control_Cold, CORT_Hot, Control_Hot) %>%
+  pivot_longer(cols = c(CORT_Cold, Control_Cold, CORT_Hot, Control_Hot), names_to = "Treatment", values_to = "ROS") %>%
+  mutate(Treatment = factor(Treatment,
+    levels = c("CORT_Cold", "Control_Cold", "CORT_Hot", "Control_Hot"),
+    labels = c("CORT_Cold" = "CORT-Cold (n=20)",
+              "Control_Cold" = "Control-Cold (n=20)",
+              "CORT_Hot" = "CORT-Hot (n=18)",
+              "Control_Hot" = "Control-Hot (n=20)")
+  )) %>%
+data.frame()
+#
+data_plot_ROS <- tidy_post_fig_ROS %>%
+  group_by(Treatment) %>%
+  summarize(
+    Mean = mean(ROS),
+    SD = sd(ROS),
+    SE = sd(ROS)/sqrt(length(ROS))
+  ) %>%
+  ungroup() %>%
+data.frame()
+write.csv(data_plot_ROS, here("output/Checking/data_plot_ROS.csv"))
+#
+fig_ROS <- ggplot(tidy_post_fig_ROS, aes(x = Treatment, y = ROS, fill = Treatment)) +
+  geom_flat_violin(alpha = 0.5) +
+  scale_fill_manual(values = c("CORT-Cold (n=20)"="#00008B", "Control-Cold (n=20)"="#68bde1", 
+                    "CORT-Hot (n=18)"="#b50101", "Control-Hot (n=20)"="#fa927d")) +
+  geom_point(data = data_plot_ROS, aes(y = Mean, x = Treatment), position = position_dodge(width = 0.75), color = "black", fill = "black", size = 3) +
+  geom_segment(data = data_plot_ROS, aes(y = Mean - SD, yend = Mean + SD, x = Treatment, xend = Treatment), size = 1.5, color = "black") +
+  geom_hline(yintercept = mean(data_plot_ROS$Mean), linetype = "dashed", color = "black") +
+  ylim(min(tidy_post_fig_ROS$ROS), max(tidy_post_fig_ROS$ROS)) +
+  coord_flip() +
+  theme_classic() +
+  labs(y = "Slope estimates", x = "Treatments") +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
+    axis.text.y = element_blank(),   # Remove y-axis labels
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "right",
+    legend.title = element_text(size = 12, family = "Times"),
+    legend.text = element_text(size = 11, family = "Times")
+    )
+ggsave("./output/figures/fig_ROS.png", plot=fig_ROS, width = 15, height = 15, units = "cm", dpi = 600)
+#
+#
+r_choice <- cor.test(clean_df$ROS, clean_df$choice_slope_mean, method = "pearson")
+r_choice_text <- paste("r = ", round(r_choice$estimate, 2),
+              " [", round(r_choice$conf.int[1], 2), ", ", round(r_choice$conf.int[2], 2), "]",
+              sep = "")
+#
+r_errors <- cor.test(clean_df$ROS, clean_df$errors_slope_mean, method = "pearson")
+r_errors_text <- paste("r = ", round(r_errors$estimate, 2),
+              " [", round(r_errors$conf.int[1], 2), ", ", round(r_errors$conf.int[2], 2), "]",
+              sep = "")
+#
+plot_ROS_choice <- ggplot(data = clean_df, aes(x = ROS, y = choice_slope_mean)) +
+  geom_smooth(method = "lm", color = "#8D0001", fill = "#FFC3C3") +
+  theme_classic() +
+  labs(y = "Estimated learning slope (Choice)", x = "ROS in medial cortex") +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "none"
+  ) +
+annotate("text", x = Inf, y = Inf, label = r_choice_text, 
+           hjust = 1.05, vjust = 1.95, size = 4.5, color = "#8D0001")
+#
+plot_ROS_errors <- ggplot(data = clean_df, aes(x = ROS, y = errors_slope_mean)) +
+  geom_smooth(method = "lm", color = "#8D0001", fill = "#FFC3C3") +
+  theme_classic() +
+  labs(y = "Estimated learning slope (Errors)", x = "ROS in medial cortex") +
+  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "none"
+  ) +
+annotate("text", x = Inf, y = Inf, label = r_errors_text, 
+           hjust = 1.05, vjust = 1.95, size = 4.5, color = "#8D0001")
+#
+fig_ROS_learning <- plot_grid(plot_ROS_choice, plot_ROS_errors, nrow = 1) 
+ggsave("./output/figures/fig_ROS_learning.png", plot=fig_ROS_learning, width = 25, height = 15, units = "cm", dpi = 600)
+
 ```

### 2024-07-30 .vdoc.r

@@ -121,9 +121,34 @@ source(here("R", "func.R"))
 # 
 # df got from data_process.R
 #
-fchoice <- (choice_slope__mean + choice_slope__sd ~ cort*temp)
-ferrors <- (errors_slope__mean + errors_slope__sd ~ cort*temp)
-fROS <- 
+fchoice <- bf(choice_slope_mean + choice_slope_se ~ cort*temp)
+ferrors <- bf(errors_slope_mean + errors_slope_se ~ cort*temp)
+fROS <- bf(ROS ~ Mit_density + cort*temp)
+#
+refit <- FALSE
+if(refit==TRUE){
+  # Choice
+  model_choice_2 <- brm(fchoice + fROS,
+                data = clean_df,
+                family = gaussian,
+                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
+  # Errors
+  model_errors_2 <- brm(ferrors + fROS,
+                data = clean_df,
+                family = gaussian,
+                chains = 4, cores = 4, iter = 8000, warmup = 2000,
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
+  # Write the models to two files
+  saveRDS(model_choice_2, file = paste0(here("output/models/model_choice_2.rds")))
+  saveRDS(model_errors_2, file = paste0(here("output/models/model_errors_2.rds")))
+  } else {
+    # Read the model from a file
+    model_choice_2 <- readRDS(file = paste0(here("output/models/model_choice_2.rds")))
+    model_errors_2 <- readRDS(file = paste0(here("output/models/model_errors_2.rds")))
+} 
+post_gen <- as_draws_df(model_errors_2)
+write.csv(post_gen, file= "./output/Checking/post_gen.csv")
 #
 #
 #
@@ -277,7 +302,6 @@ real_table
[…]

### 2024-07-30 R/data_process.R

@@ -31,8 +31,7 @@ if(refit==TRUE){
                 data = data_spal,
                 family = bernoulli(link = "logit"),
                 chains = 4, cores = 4, iter = 8000, warmup = 2000, 
-                control = list(adapt_delta = 0.99, max_treedepth = 12),
-                prior = custom_priors)
+                control = list(adapt_delta = 0.99, max_treedepth = 12))
   # Errors
   model_errors <- brm(errors ~ 1 + (1 + day|lizard_id) + (1|clutch),
                 data = data_spal,
@@ -89,20 +88,20 @@ physio_df <- merge(mit_df, euth, by = c("plate", "tube"))
 mod_physio_df <- physio_df %>%
   filter(type != "control", ch == "percp.cy5.5"| ch == "apc") %>%
   mutate(ch = recode(ch, "percp.cy5.5" = "ROS", "apc" = "Mit_density")) %>%
-  select(tube, lizard_id, ch, geo.mean, arith.mean) %>%
-  group_by(lizard_id, ch) %>%
-  summarize(geo.mean = mean(geo.mean, na.rm = TRUE), 
-            arith.mean = mean(arith.mean, na.rm = TRUE),
+  select(tube, lizard_id, sex, ch, arith.mean) %>%
+  group_by(lizard_id, ch, sex) %>%
+  summarize(arith.mean = mean(arith.mean, na.rm = TRUE),
             .groups = 'drop') %>%
   pivot_wider(
     names_from = ch,
-    values_from = c(geo.mean, arith.mean)
+    values_from = arith.mean
   ) %>%
 data.frame()
-write.csv(mod_physio_df, file = "./output/Checking/physio_df.csv")
 #
 #
 #
 # C) MERGING THE DF TO GET THE FINAL ONE
-clean_df <- merge(learning_df, mod_physio_df, by = c("lizard_id"))
+clean_df <- merge(learning_df, mod_physio_df, by = c("lizard_id")) %>%
+  rename("choice_slope_mean" = "choice_slope__mean", "choice_slope_se" = "choice_slope__se",
+        "errors_slope_mean" = "errors_slope__mean", "errors_slope_se" = "errors_slope__se")
 write.csv(clean_df, file = "./output/databases_clean/clean_df.csv")

### 2024-07-30 R/func.r

@@ -38,7 +38,8 @@ tidy_post_df <- function(df, effect) {
     summarise(across(everything(), list(
       mean = ~ mean(.x, na.rm = TRUE),
       sd = ~ sd(.x, na.rm = TRUE),
-      se = ~ sd(.x, na.rm = TRUE)/sqrt(length(.x))
+      se = ~ sd(.x, na.rm = TRUE)/sqrt(length(.x)),
+      pmcmc = ~ pmcmc(mean(.x), null = 0)
     ))) %>%
   data.frame()
   tidy_post_2 <- tidy_post_1 %>%

### 2024-09-23 .vdoc.r
Updates

@@ -1,606 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-
-```{r, fig-Methods}
-#| label: fig-Methods
-#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
-
-knitr::include_graphics("./Others/SPAL_METH.svg")
-
-#
-#
-#
-#
-#
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: data_processing
-# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
-refit = FALSE
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: sampleSize
[…]

### 2024-10-09 ms.qmd
Updates

@@
 #
-specie <- c("delicata", "guichenoti")
-groups <- c("Red", "Blue")
 hormone <- c("CORT", "Control")
@@
 #
-for(i in 1:length(specie)){
-  for(j in 1:length(groups)){
 for(k in 1:length(hormone)){
   for(l in 1:length(temperature)){
-        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
-      }
-    }
+    n_list[[paste(hormone[k], temperature[l], sep = "_")]] <- sample(hormone[k], temperature[l])
   }
@@
   theme_classic() +
-  labs(y = "Slope estimates", x = "Treatments") +
+  labs(y = "ROS", x = "Treatments") +
   theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
@@
   theme(
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
+    axis.title = element_text(size = 14, family = "Times"),
+    axis.text = element_text(size = 12, family = "Times"),
     legend.position = "none"
@@
 annotate("text", x = Inf, y = Inf, label = r_errors_text, 
-           hjust = 1.05, vjust = 1.95, size = 4.5, color = "#8D0001")
+           hjust = 1.28, vjust = 1.95, size = 4.5, color = "#8D0001")
 #
@@
 ggsave("./output/figures/fig_ROS_learning.png", plot=fig_ROS_learning, width = 25, height = 15, units = "cm", dpi = 600)
-
+ggsave("./output/figures/fig_ROS_learning_2.png", plot=plot_ROS_errors, width = 15, height = 10, units = "cm", dpi = 600)
 ```

### 2024-12-28 ms.qmd
Update ms.qmd

@@
 ## Introduction
+**Corticosterone is the major GC produced by
+adrenal steroidogenic cells in reptiles, and an increase in
+its plasma concentration is a frequent indicator of stress
+(Anderson et al. 2014 General and Comparative Endocrinology). In nature,
+GCs in the developing embryos seem to be largely of
+maternal origin (Uller et al. 2009, The Journal of Experimental Biology) and are probably transferred to the growing oocytes through vitellogenesis (Elf et al.,
+2002, 2003 General and Comparative Endocrinology (both)).**
+
+Mechanistically, a change in body size requires
+a change in cell number and/or cell size, and there is evidence that
+body size and cell size undergo coordinated evolutionary (Arendt,
+2007; Starostová et al., 2005; Stevenson et al., 1995) or
+phenotypically plastic changes (Arendt, 2007; Czarnoleski et al.,
+2013; Hessen et al., 2013).
 
@@
 #### Learning
-
-```{r, fig-Methods}
 

### 2025-01-12 ms.qmd
Update ms.qmd

@@
 
-#### Husbandry  
-*Breeding colony* -- The study was conducted with juveniles of the Delicate skink (_Lamprophpolis delicata_) from a breeding colony kept in the lab since 2019.
-The colony is composed by a total of 270 adults housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure.
-Enclosures are provided with non-stick matting, shelter, and several small water dishes.
-Water is given daily, and they are fed approx.
-40 mid-size crickets (*Acheta domestica*) per enclosure three days a week.
-Crickets are dusted with calcium weekly and multivitamin and calcium biweekly.
-To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle.
-Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.
+#### Animal husbandry  
+*Breeding colony* - We tested _L.
+delicata_ coming from a breeding colony established in the laboratory since 2019.
+This colony consisted of 270 adults housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure.
+Enclosures were provided with nonstick matting, shelter, and several small water dishes.
+Water was given daily, and the lizards were fed approx.
+40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
+Crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
+Room temperatures were set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
 
-*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above).
-We checked for the presence of eggs in the boxes three days a week.
-After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.
+*Eggs collection and incubation* - Between mid-November 2023 and mid-January 2024, we provide females with a place to lay the eggs by placing a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures.
+These boxes were checked three days a week.
+After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighted the eggs with a digital scale ± 0.001g error.
+Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.
 
-*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings.
-After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error.
-We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish.
-During this period, they were sprayed water every day and received 3-6 small *A.
-domestica* crickets three times a week.
-All care otherwise follows similar protocols to adults (see above).
-Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation.
-The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other.
-These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below).
-The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack.
-During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*.
-We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle.
-The rooms temperature was set to between 22-24 Celsius.
+*Hatchlings* - Incubators were checked three times a week for hatchlings.
+Lizards were measured and weighed immediately after hatching.
+Snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a digital scale with an accuracy of ± 0.001 g.
+Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with nonstick matting and a small water dish.
+Watering, feeding, and temperature conditions were maintained as for adults (see above).
 
 #### CORT and Temperature manipulation
-To test empirically the effect of early environment we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) in a 2x2 factorial design (@fig-Methods A).
-We first allocated eggs to one of two different treatments: CORT treatment, where eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO (vehicle) at a final (10 pg CORT/mL) concentration (CORT treatment); and a Control treatment, where eggs received an equal volume of the vehicle.
-CORT concentration employed in the CORT treatment represents 2 standard deviations above the mean natural concentration obtained in eggs from both species (non-published data).
-Then, eggs were incubated in one of the two previously mentioned temperature regimes ('Cold' or 'Hot') until hatching.
-The number of eggs per clutch assigned to each hormone and temperature treatment were counterbalanced in both species.
+To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig).
+Eggs were assigned to one fo the two hormone treatments: CORT or Control.
+CORT-treated lizards were topically supplied with 5 µL of crystalline corticosterone (Sigma, Cat.
+No.
+C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), while Control lizards received an equal volume of 100% Ethanol (Control treatment).
+We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [**ONDI'S PAPER**].
+Then, eggs were incubated in one of the two previously mentioned temperature regimes.
+These temperatures are within the natural limits in _L.
+delicata_ [@cheetham2011embryonic].
+The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design.
+At least one egg per clutch was assigned to each treatment, while in cluthces bigger than four the remaining eggs were randomly assigned to one of the treatments.
+When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
 
 #### Learning  
+The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact.
+The maze was constructed from white polylactic acid (PLA) and featured six arms extending from a hexagonal central starting point (see @fig).
+
+In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes  (see @fig).
+During acclimatation, the central are was surrounded by a yellow 3D-printed PLA device mounted on a pulley system.
+At the start of each trial, this device was lifted to startle the lizard.
+If the lizard did not immediately choose an arm, it was gently prodded with a brush at the end of the tail.
+Once the lizard made a choice, the brush was used to encourage movement but without guiding it towards any specific arm.
+If the lizard did not choose correctly after 20 errors — at which point it typically ceased responding — the lizard was gently guided to the correct arm.
+
+Extra-maze cues (but not intra-maze cues) were available to assist with navigation.
+This design aimed to assess the lizards' ability to learn spatial relationships based on external cues rather than relying on familiar objects within the maze.
+To avoid confounding effects from chemical cues, the maze was cleaned with 70% ethanol between trials.
+To control for visual cues related to the correct arm, the maze was rotated every three trials while maintaining the orientation of the correct arm and the maze's position within the room, which remained consistent for each individual.
+
+To control for potential side biases, the correct arm was randomly assigned to one of the six arms for each trial.
+We employed four different maze orientations (North-West, South-West, South, and South-East), and the number of lizards assigned to each orientation was counterbalanced across treatments.
+
+The task was repeated once daily for 40 consecutive days, and the number of mistakes made by each lizard was recorded.
+A mistake was defined as the lizard inserting its head into one of the incorrect arms.
+The primary measure of spatial learning was the total number of errors made before the lizard chose the correct arm.
 

### 2025-01-24 ms.qmd
Update ms.qmd

@@
 ```
+
+#### Brain extraction
+
+#### Brain mitochondrial activity
+Two months after the completion of the tests, we euthanises lizards usin an injectable anaesthetic followed by decapitation.
+We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution (**REF?**).
+After some minutes, we evaluated the lizards response by turning it upside down, and by testing the pinching relfex in the hand.
+Lizards were decapitated once they were not responsive.
+
+**After decapitation, the head was opened and the brain dissected.
+We focused on two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with lizards' chemical and visual perception, respectively (Wynecken, 2007).**
 

### 2025-03-02 ms.qmd
updates

@@
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
+```
+
+```{r, data_processing}
+#| label: data_processing
+# The result will be the final df with the data for the analysis.
+To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder).
+The final df will be saved in (here("output/databases_clean/data_complete.csv")
+refit <- FALSE
+source(here("R", "data_process.R"))
+```
+
+```{r, data_metrics}
+#| label: data_metrics
+# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
+data <- read.csv(here("./data/Spatial_learn.csv"))
+#
+# Total number of clutches
+total_clutches <- data %>% 
+  distinct(clutch) %>% 
+  count() %>% 
+  pull(n)
+# Calculate total number of individuals
+total_individuals <- data %>% 
+  distinct(lizard_id) %>% 
+  count() %>% 
+  pull(n)
+# Calculate number of individuals per clutch
+individuals_per_clutch <- data %>%
+  distinct(lizard_id, clutch) %>%
+  group_by(clutch) %>%
+  summarize(individual_count = n())
 ```
@@
 
-```{r, data_metrics}
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-```
-
 We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
@@
 ## Results
-
-```{r, data_processing}
-#| label: data_processing
-# The result will be the final df with the data for the analysis.
-To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder).
-The final df will be saved in (here("output/databases_clean/data_complete.csv")
-refit = FALSE
-source(here("R", "data_process.R"))
-```
-
 ```{r, sampleSize}

### 2025-03-02 R/data_process.R

@@ -5,8 +5,8 @@ pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4,
 #
 # A) PROCESSING THE LEARNING DF
 data <- read.csv(here("./data/Spatial_learn.csv"))
-#### A.1) Remove individuals who did not participate (more than 15 NAs), remove trials [36-40] (Only in associative) and split treatment into Temp and Cort
-data_spal <- data %>%
+#### A.1) Modify the learning df
+learning_df <- 
   group_by(lizard_id) %>%
     filter(sum(is.na(choice)) <= 15) %>%
   ungroup()  %>%
@@ -18,37 +18,27 @@ data_spal <- data %>%
     mutate(cort = factor(cort,
       levels = c("B", "A"),
       labels = c("B" = "CORT", "A" = "Control"))) %>%
-  ungroup() %>% 
+  ungroup() %>%
 data.frame()
-#
+write.csv(learning_df, here("./output/databases_clean/learning_df.csv"))
 #
 #### A.2) Extract the slope of each individual
 #
-## Fit the models only if they have not been fit yet (if refit = TRUE)
-if(refit==TRUE){
-  # Choice
-  model_choice <- brm(choice ~ 1 + (1 + day|lizard_id) + (1|clutch),
-                data = data_spal,
-                family = bernoulli(link = "logit"),
-                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Errors
-  model_errors <- brm(errors ~ 1 + (1 + day|lizard_id) + (1|clutch),
+## Fit the model only if it has not been fit yet (if refit = TRUE)
+if(refit){
+  model_errors <- brm(errors ~ 1 + (1 + day|lizard_id),
                 data = data_spal,
                 negbinomial(link = "log"),
                 chains = 4, cores = 4, iter = 8000, warmup = 2000,
                 control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Write the models to two files
-  saveRDS(model_choice, file = paste0(here("output/models/model_choice.rds")))
+  # Write the model to rds files
   saveRDS(model_errors, file = paste0(here("output/models/model_errors.rds")))
   } else {
     # Read the model from a file
[…]

### 2025-03-04 ms.qmd
updates

@@
 #| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, PupillometryR, cowplot, png, grid)
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
 ```
@@
 
-```{r, ROS-learning}
-#| label: ROS-learning
-tidy_post_fig_ROS <- post_gen %>%
-  mutate(CORT_Cold = b_ROS_Intercept,
-      Control_Cold = b_ROS_Intercept + b_ROS_cortControl,
-      CORT_Hot = b_ROS_Intercept + b_ROS_tempHot,
-      Control_Hot = b_ROS_Intercept + b_ROS_cortControl + b_ROS_tempHot + `b_ROS_cortControl:tempHot`) %>%
-  select(CORT_Cold, Control_Cold, CORT_Hot, Control_Hot) %>%
-  pivot_longer(cols = c(CORT_Cold, Control_Cold, CORT_Hot, Control_Hot), names_to = "Treatment", values_to = "ROS") %>%
-  mutate(Treatment = factor(Treatment,
-    levels = c("CORT_Cold", "Control_Cold", "CORT_Hot", "Control_Hot"),
-    labels = c("CORT_Cold" = "CORT-Cold (n=20)",
-              "Control_Cold" = "Control-Cold (n=20)",
-              "CORT_Hot" = "CORT-Hot (n=18)",
-              "Control_Hot" = "Control-Hot (n=20)")
-  )) %>%
-data.frame()
-#
-data_plot_ROS <- tidy_post_fig_ROS %>%
-  group_by(Treatment) %>%
-  summarize(
-    Mean = mean(ROS),
-    SD = sd(ROS),
-    SE = sd(ROS)/sqrt(length(ROS))
-  ) %>%
-  ungroup() %>%
-data.frame()
-write.csv(data_plot_ROS, here("output/Checking/data_plot_ROS.csv"))
-#
-fig_ROS <- ggplot(tidy_post_fig_ROS, aes(x = Treatment, y = ROS, fill = Treatment)) +
-  geom_flat_violin(alpha = 0.5) +
-  scale_fill_manual(values = c("CORT-Cold (n=20)"="#00008B", "Control-Cold (n=20)"="#68bde1", 
-                    "CORT-Hot (n=18)"="#b50101", "Control-Hot (n=20)"="#fa927d")) +
-  geom_point(data = data_plot_ROS, aes(y = Mean, x = Treatment), position = position_dodge(width = 0.75), color = "black", fill = "black", size = 3) +
-  geom_segment(data = data_plot_ROS, aes(y = Mean - SD, yend = Mean + SD, x = Treatment, xend = Treatment), size = 1.5, color = "black") +
-  geom_hline(yintercept = mean(data_plot_ROS$Mean), linetype = "dashed", color = "black") +
-  ylim(min(tidy_post_fig_ROS$ROS), max(tidy_post_fig_ROS$ROS)) +
-  coord_flip() +
-  theme_classic() +
-  labs(y = "ROS", x = "Treatments") +
-  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm"),
-    axis.text.y = element_blank(),   # Remove y-axis labels
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
-    legend.position = "right",
-    legend.title = element_text(size = 12, family = "Times"),
-    legend.text = element_text(size = 11, family = "Times")
-    )
-ggsave("./output/figures/fig_ROS.png", plot=fig_ROS, width = 15, height = 15, units = "cm", dpi = 600)
-#
-#
-r_choice <- cor.test(clean_df$ROS, clean_df$choice_slope_mean, method = "pearson")
-r_choice_text <- paste("r = ", round(r_choice$estimate, 2),
-              " [", round(r_choice$conf.int[1], 2), ", ", round(r_choice$conf.int[2], 2), "]",
-              sep = "")
-#
-r_errors <- cor.test(clean_df$ROS, clean_df$errors_slope_mean, method = "pearson")
-r_errors_text <- paste("r = ", round(r_errors$estimate, 2),
-              " [", round(r_errors$conf.int[1], 2), ", ", round(r_errors$conf.int[2], 2), "]",
-              sep = "")
-#
-plot_ROS_choice <- ggplot(data = clean_df, aes(x = ROS, y = choice_slope_mean)) +
-  geom_smooth(method = "lm", color = "#8D0001", fill = "#FFC3C3") +
-  theme_classic() +
-  labs(y = "Estimated learning slope (Choice)", x = "ROS in medial cortex") +
-  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
-  theme(
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
-    legend.position = "none"
-  ) +
-annotate("text", x = Inf, y = Inf, label = r_choice_text, 
-           hjust = 1.05, vjust = 1.95, size = 4.5, color = "#8D0001")
-#
-plot_ROS_errors <- ggplot(data = clean_df, aes(x = ROS, y = errors_slope_mean)) +
-  geom_smooth(method = "lm", color = "#8D0001", fill = "#FFC3C3") +
-  theme_classic() +
-  labs(y = "Estimated learning slope (Errors)", x = "ROS in medial cortex") +
-  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
-  theme(
-    axis.title = element_text(size = 14, family = "Times"),
-    axis.text = element_text(size = 12, family = "Times"),
-    legend.position = "none"
-  ) +
-annotate("text", x = Inf, y = Inf, label = r_errors_text, 
-           hjust = 1.28, vjust = 1.95, size = 4.5, color = "#8D0001")
-#
-fig_ROS_learning <- plot_grid(plot_ROS_choice, plot_ROS_errors, nrow = 1) 
-ggsave("./output/figures/fig_ROS_learning.png", plot=fig_ROS_learning, width = 25, height = 15, units = "cm", dpi = 600)
-ggsave("./output/figures/fig_ROS_learning_2.png", plot=plot_ROS_errors, width = 15, height = 10, units = "cm", dpi = 600)
-```
-
 ## Discussion
@@
 # Suplementary Material
+#### Results of preliminary models
 
-#### Checking the models plots  
-```{r, residuals}
-# Chunk for calling all the models and making the residuals
-# Associative task
-## L. delicata
-### Red
-mod_dar <- readRDS(here("output/models/asso_deli_red.rds"))
-resid_dar <- residuals(mod_dar)
-### Blue
-mod_dab <- readRDS(here("output/models/asso_deli_blue.rds"))
-resid_dab <- residuals(mod_dab)
-## L. guichenoti
-### Red
-mod_gar <- readRDS(here("output/models/asso_guich_red.rds"))
-resid_gar <- residuals(mod_gar)
-### Blue
-mod_gab <- readRDS(here("output/models/asso_guich_blue.rds"))
-resid_gab <- residuals(mod_gab)
-# Reversal task
-## L. delicata
-### Red
-mod_drr <- readRDS(here("output/models/rev_deli_red.rds"))
-resid_drr <- residuals(mod_drr)
-### Blue
-mod_drb <- readRDS(here("output/models/rev_deli_blue.rds"))
-resid_drb <- residuals(mod_drb)
-## L. guichenoti
-### Red
-mod_grr <- readRDS(here("output/models/rev_guich_red.rds"))
-resid_grr <- residuals(mod_grr)
-### Blue
-mod_grb <- readRDS(here("output/models/rev_guich_blue.rds"))
-resid_grb <- residuals(mod_grb)
+```{r, models_preliminary}
+#| label: models_preliminary
+# Fitting preliminary models to see if sex and age are relevant for our models
+source(here("R", "func.R"))
+#
+#
+var_m <- c("mit_density", "mit_potential", "ROS", "DNAdamage", "peroxidation")
+formula_list_ <- list()
+for (p in var_m){
+  formula_list_[[p]] <- paste0(p, "~ cort*temp + age + sex + (1|clutch)")
+  pmodel_name <- paste0("m_prel_", p)
+  assign(pmodel_name, fit_m(df = clean_df,
+                             cat = "prel",
+                             var = p,
+                             formula = formula_list_[[p]],
+                             fam = gaussian(),
+                             refit = FALSE),
+          envir = .GlobalEnv)  # Assign to the global environment
+}
+#
+#
+# Run model behaviour (stimuli separated)
+formula_learn <- slope_mean + slope_sd ~ cort*temp + age + sex + (1|clutch)
+m_prel_learn <- fit_m(df = clean_df,
+                      cat = "prel",
+                      var = "slope_mean",
+                      formula = formula_learn,
+                      fam = gaussian(),
+                      refit = FALSE)
 ```
 
-Model formula for the reversal task is:  
-Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)  
-Plots for the different models of the associative task:  
-1.- _L. delicata_  
-  1.a.- Red  
-```{r , plotmod_drr, out.width="70%", fig.align="center"}
-plot(mod_drr)
+Table S1. Preliminary results of the models testing for Mitochondrial Density.
+
+```{r, results_preliminary_mitdensity}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density"
+#| label: results_preliminary_mitdensity
+#
+sum_mitdensity_prel <- m_prel_mit_density %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws()  %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_mitdensity_prel)
+```
+
+Model formula: mit_density ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S2.  Preliminary results of the models testing for Mitochondrial Potential.
+```{r, results_preliminary_potential}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential"
+#| label: results_preliminary_potential
+#
+sum_potential_prel <- m_prel_mit_potential %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_potential_prel)
+```
+
+Model formula: mit_potential ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S4. Preliminary results of the models testing for ROS.
+```{r, results_preliminary_ros}
+#| tbl-cap: "Preliminary results of the models testing for ROS"
+#| label: results_preliminary_ros
+#
+sum_m_ros_prel <- m_prel_ROS %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_ros_prel)
+```
+
+Model formula: ROS ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S6. Preliminary results of the models testing for DNA damage.
+```{r, results_preliminary_dnadamage}
+#| tbl-cap: "Preliminary results of the models testing for DNA damage"
+#| label: results_preliminary_dnadamage
+#
+sum_m_dnadamage_prel <- m_prel_DNAdamage %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_dnadamage_prel)
+```
+
+Model formula: DNAdamage ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates an effect of sex and age, so they were included in the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S8. Preliminary results of the models testing for lipid peroxidation.
+```{r, results_preliminary_peroxidation}
+#| tbl-cap: "Preliminary results of the models testing for lipid peroxidation"
+#| label: results_preliminary_peroxidation
+#
+sum_m_peroxidation_prel <- m_prel_peroxidation %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_peroxidation_prel)
+```
+
+Model formula: peroxidation ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of sex, so it was discarded from the final models.
+However, we saw an effect of age and we included it in our final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Results of the final models
+
+Table S**13**. R^2^ values of the final models.
+```{r, table_bayesR2}
+#| label: table_bayesR2
+#| tbl-cap: "BayesR2 values of the final models"
+#
+data_bayes <- data.frame(
+  Model = character(0),
+  Mean = numeric(0),
+  Error = numeric(0),
+  Q2_5 = numeric(0),
+  Q97_5 = numeric(0)
+)
+models <- c("mean_mitodensity_def_OB",
+            "mean_potential_def_OB",
+            "mean_ros_def_OB",
+            "mean_dnadamage_def_OB",
+            "mean_peroxidation_def_OB",
+            "t_D_def_Chemical",
+            "mean_mitodensity_def_OT",
+            "mean_potential_def_OT",
+            "mean_ros_def_OT",
+            "mean_dnadamage_def_OT",
+            "mean_peroxidation_def_OT",
+            "t_D_def_Visual")
+#
+for (m in models){
+  mod <- readRDS(here("output/models/", paste0(m, ".rds")))
+  bayes <- bayes_R2(mod)
+  data_bayes <- rbind(data_bayes, data.frame(
+    Model = m,
+    Mean = format_dec(bayes[1], 3),
+    Error = format_dec(bayes[2], 3),
+    Q2_5 = format_dec(bayes[3], 3),
+    Q97_5 = format_dec(bayes[4], 3)
+  ))
+}
+#
+bayes_table_df <- data_bayes %>%
+  mutate(Region = gsub(".*_", "", Model)    # Extract everything after the last "_"
+  ) %>%
+  mutate(Model = factor(Model,
+                        levels = c("mean_mitodensity_def_OB",
+                                  "mean_potential_def_OB",
+                                  "mean_ros_def_OB",
+                                  "mean_dnadamage_def_OB",
+                                  "mean_peroxidation_def_OB",
+                                  "t_D_def_Chemical",
+                                  "mean_mitodensity_def_OT",
+                                  "mean_potential_def_OT",
+                                  "mean_ros_def_OT",
+                                  "mean_dnadamage_def_OT",
+                                  "mean_peroxidation_def_OT",
+                                  "t_D_def_Visual"),
+                        labels = c("m_def_mean_mitodensity_OB" = "Mit density",
+                                  "m_def_mean_potential_OB" = "Mit potential",
+                                  "m_def_mean_ros_OB" = "ROS",
+                                  "m_def_mean_dnadamage_OB" = "DNA damage",
+                                  "m_def_mean_peroxidation_OB" = "Peroxidation",
+                                  "m_def_t_D_Chemical" = "Detection lat",
+                                  "m_def_mean_mitodensity_OT" = "Mit density",
+                                  "m_def_mean_potential_OT" = "Mit potential",
+                                  "m_def_mean_ros_OT" = "ROS",
+                                  "m_def_mean_dnadamage_OT" = "DNA damage",
+                                  "m_def_mean_peroxidation_OT" = "Peroxidation",
+                                  "m_def_t_D_Visual" = "Detection lat")),
+      Region = factor(Region, levels = c("OB", "Chemical", "OT", "Visual"),
+                      labels = c("OB" = "Olfactory bulbs",
+                                "Chemical" = "Chemical",
+                                "OT" = "Optic tecta",
+                                "Visual" = "Visual"))) %>%
+  dplyr::select(Region, Model, Mean, Error, Q2_5, Q97_5) %>%
+  arrange(Region, Model)
+#
+#
+# Create the table
+#
+## Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+bayes_table <- flextable(bayes_table_df) %>%
+  set_header_labels(
+    Region = "Region/Stimulus",
+    Mean = "Mean",
+    Error = "Error",
+    Q2_5 = "2.5%",
+    Q97_5 = "97.5%") %>%
+  align(align = "center", j = c(3:5), part = "body") %>%
+  align(align = "center", j = c(1:5), part = "header") %>%
+  flextable::compose(i = c(2:5,8:11), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = 6, part = "body") %>% 
+  autofit()
+#
+bayes_table
 ```
@@
 
-  1.b.- Blue  
-```{r , plotmod_drb, out.width="70%", fig.align="center"}  
-plot(mod_drb)
+Table S**14**. Results of the final models.
+```{r, tbl-results_OB_table}
+#| label: results_OB_table
+#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
+#| tbl-name: "results_OB"
+#| tbl-label: "results_OB"
+source(here("R", "func.R"))
+# 
+# A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB)
+post_OB_refined <- refine_post(post_OB) %>%
+  arrange(Variable, Predictors)
+#
+# B) Create table
+## Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+OB_table <- flextable(post_OB_refined) %>%
+  align(align = "center", j = c(3:5), part = "body") %>%
+  align(align = "center", j = c(1:5), part = "header") %>%
+  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2:4,6:8,10:12,14:18,20:23,25:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  autofit()
+#
+OB_table
 ```
@@
 
-2.- _L. guichenoti_  
-  2.a.- Red  
-```{r , plotmod_grr, out.width="70%", fig.align="center"}
-plot(mod_grr)
+#### Final models diagnostics (plots)
+
+```{r , plotmod_mitdensity, out.width="70%", fig.align="center"}
+#| label: plotmod_mitdensity
+#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mit_density_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
 ```
+
+Fig S1. Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs.
+Formula: mean_mitodensity ~ cort * temp + (1|clutch)
 
@@
 
-  2.b.- Blue  
-```{r , plotmod_grb, out.width="70%", fig.align="center"}
-plot(mod_grb)
+```{r , plotmod_mitdensity_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_mitdensity_OT
+#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_mitodensity_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
 ```
+
+Fig S2. Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta.
+Formula: mean_mitodensity ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_potential_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_potential_OB
+#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_potential_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
+```
+
+Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_potential_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_potential_OT
+#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_potential_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
+```
+
+Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
+Formula: mean_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_ros_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_ros_OB
+#| caption: "Posterior predictive checks for the model of ROS Production in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_ros_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
+```
+
+Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_ros_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_ros_OT
+#| caption: "Posterior predictive checks for the model of ROS Production in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_ros_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
+```
+
+Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
+Formula: mean_ros ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_dnadamage_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_dnadamage_OB
+#| caption: "Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_dnadamage_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
+```
+
+Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
+Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_dnadamage_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_dnadamage_OT
+#| caption: "Posterior predictive checks for the model of DNA Damage in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_dnadamage_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
+```
+
+Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
+Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_peroxidation_OB, out.width="70%", fig.align="center"}
+#| label: plotmod_peroxidation_OB
+#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mean_peroxidation_def_OB.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
+```
+
+Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
+Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_peroxidation_OT, out.width="70%", fig.align="center"}
+#| label: plotmod_peroxidation_OT
+#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta."
+#
+mod <- readRDS(here("output/models/mean_peroxidation_def_OT.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
+```
+
+Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
+Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_tD_Chemical, out.width="70%", fig.align="center"}
+#| label: plotmod_tD_Chemical
+#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials."
+#
+mod <- readRDS(here("output/models/t_D_def_Chemical.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
+```
+
+Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
+Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_tD_Visual, out.width="70%", fig.align="center"}
+#| label: plotmod_tD_Visual
+#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials."
+#
+mod <- readRDS(here("output/models/t_D_def_Visual.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
+```
+
+Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
+Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-03-04 .vdoc.r

@@ -1,0 +1,1018 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: data_processing
+# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
+refit <- FALSE
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: data_metrics
+# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
+data <- read.csv(here("./data/Spatial_learn.csv"))
+#
+# Total number of clutches
+total_clutches <- data %>% 
+  distinct(clutch) %>% 
+  count() %>% 
+  pull(n)
+# Calculate total number of individuals
+total_individuals <- data %>% 
+  distinct(lizard_id) %>% 
+  count() %>% 
+  pull(n)
+# Calculate number of individuals per clutch
+individuals_per_clutch <- data %>%
+  distinct(lizard_id, clutch) %>%
+  group_by(clutch) %>%
+  summarize(individual_count = n())
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: fig-Methods
+#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
+
[…]

### 2025-03-04 R/data_process.R

@@ -2,11 +2,11 @@
 # 1_data_process
 ####################################
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, forcats)
+source(here("R", "func.R"))
 #
-# A) PROCESSING THE LEARNING DF
-data <- read.csv(here("./data/Spatial_learn.csv"))
+# A) PROCESSING THE LEARNING DF 
 #### A.1) Modify the learning df
-learning_df <- 
+learning_df <- read.csv(here("./data/Spatial_learn.csv")) %>%
   group_by(lizard_id) %>%
     filter(sum(is.na(choice)) <= 15) %>%
   ungroup()  %>%
@@ -27,7 +27,7 @@ write.csv(learning_df, here("./output/databases_clean/learning_df.csv"))
 ## Fit the model only if it has not been fit yet (if refit = TRUE)
 if(refit){
   model_errors <- brm(errors ~ 1 + (1 + day|lizard_id),
-                data = data_spal,
+                data = learning_df,
                 negbinomial(link = "log"),
                 chains = 4, cores = 4, iter = 8000, warmup = 2000,
                 control = list(adapt_delta = 0.99, max_treedepth = 12))
@@ -38,60 +38,83 @@ if(refit){
     model_errors <- readRDS(file = paste0(here("output/models/model_errors.rds")))
 } 
 #
-## Extract posteriors
-posteriors_errors <- as_draws_df(model_errors)
-#
-#
 #### A.3) Modify posteriors df to get the mean slope and intercept of each individual (the latter would be useful for the figures)
+posteriors_errors <- as_draws_df(model_errors) %>%
+  dplyr::select(starts_with("r_")) %>%
+  rename_with(~ str_replace(.x, "r_lizard_id\\[(\\d+),\\s*(Intercept|day)\\]", "\\1_\\2")) %>%
+  summarise(across(everything(), list(
+    mean = ~ mean(.x, na.rm = TRUE),
+    sd = ~ sd(.x, na.rm = TRUE)
+    ))) %>%
+  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
+  mutate(
+    lizard_id = gsub("_.*", "", variable),
+    type = gsub("^\\d+_", "", variable)
+  ) %>%
+  distinct(lizard_id, type, value) %>%
+  pivot_wider(
[…]

### 2025-03-04 R/exploratory.R

@@ -1,0 +1,117 @@
+####################################
+# Exploratory analyses
+####################################
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes,
+ggh4x, cowplot, fitdistrplus, MASS, goftest)
+#
+source(here("R", "func.R"))
+#
+# 
+data_expl <- clean_df %>%
+  mutate(treatment = factor(str_c(cort, temp, sep = "-"),
+                          levels = c("Control-Cold",
+                                    "CORT-Cold",
+                                    "Control-Hot",
+                                    "CORT-Hot")))
+#
+# A) Make histograms for the main variables and figures factorizing main predictors
+var <- c("slope_mean", "mit_density", "mit_potential", "ROS", "DNAdamage", "peroxidation")
+
+hist_list <- list()
+for (i in var) {
+  bin_width <- (max(data_expl[[i]], na.rm = TRUE) - min(data_expl[[i]], na.rm = TRUE)) / sqrt(length(data_expl[[i]]))
+  # Labels
+  if (i == "slope_mean"){
+     lab <- "Learning"
+  } else if (i == "mit_density") {
+     lab <- "Mitochondrial density"
+  } else if (i == "mit_potential") {
+     lab <- "Mitochondrial potential"
+  } else if (i == "ROS") {
+     lab <- "ROS production"
+  } else if (i == "DNAdamage") {
+     lab <- "DNA damage"
+  } else if (i == "peroxidation") {
+     lab <- "Lipid peroxidation"
+  }
+  # Histograms
+  hist_plot <- ggplot(data_expl, aes(x = .data[[i]])) +
+    geom_histogram(binwidth = bin_width, fill = "#062d00", color = "black", alpha = 0.5) +
+    theme_classic() +
+    labs(x = lab, y = "Counts") +
+    theme(axis.title = element_text(size = 12, family = "Times"),
+          axis.text = element_text(size = 10, family = "Times"))
+  hist_list[[i]] <- hist_plot
+}
+fig_hist <- plot_grid(plotlist = hist_list, ncol = 3)
+#
+for (i in var) {
+  if (i == "slope_mean"){
+     lab <- "Learning"
+  } else if (i == "mit_density") {
+     lab <- "Mitochondrial density"
+  } else if (i == "mit_potential") {
[…]

### 2025-03-04 R/func.r

@@ -4,58 +4,6 @@ pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4,
 #
 ####################
 ####################
-# Function to make the first cleaning for the initial posteriors to estimate individual learning slopes or intercepts
-#' @title tidy_post_df
-#' @param df to select the posteriors to use. Options: choice or errors
-#' @param effect to select whether we want the intercept or the slopes. Options: slope or intercept
-tidy_post_df <- function(df, effect) {
-  if(df == "choice") {
-    data <- posteriors_choice
-    label_df <- "choice_"
-  } else if(df == "errors") {
-    data <- posteriors_errors
-    label_df <- "errors_"
-  } else {
-    stop("df not valid")
-  }
-  if(effect == "slope") {
-    label_effect <- "day"
-    label_df <- paste0(label_df, "slope_")
-    pattern <- "^r_lizard_id\\[\\d+,day\\]$"
-  } else if(effect == "intercept") {
-    label_effect <- "Intercept"
-    label_df <- paste0(label_df, "intercept_")
-    pattern <- "^r_lizard_id\\.\\d+\\.Intercept\\."
-    data <- data %>%
-      select(b_Intercept, matches("^r_lizard.*Intercept]$")) %>%
-      mutate(across(-b_Intercept, ~ . + b_Intercept)) %>%
-    data.frame()
-  } else {
-    stop("effect not valid")
-  }
-  tidy_post_1 <- data %>%
-    select(matches(pattern)) %>%
-    summarise(across(everything(), list(
-      mean = ~ mean(.x, na.rm = TRUE),
-      sd = ~ sd(.x, na.rm = TRUE),
-      se = ~ sd(.x, na.rm = TRUE)/sqrt(length(.x)),
-      pmcmc = ~ pmcmc(mean(.x), null = 0)
-    ))) %>%
-  data.frame()
-  tidy_post_2 <- tidy_post_1 %>%
-    pivot_longer(cols = everything(),
-              names_to = "lizard_id", 
-              values_to = "value") %>% # Extract the relevant columns and reshape them
-    mutate(lizard_id = gsub("r_lizard_id\\.", "", lizard_id),  # Remove prefix
-           lizard_id = gsub(paste0(label_effect, "\\."), "", lizard_id)) %>%      # Replace all remaining dots with underscores
-    separate(lizard_id, into = c("lizard_id", "stat"),
[…]

### 2025-03-06 ms.qmd
updates

@@
 ## Results
-```{r, sampleSize}
-#| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_complete.csv") as the sample size per species and group is the same on each task.
-We used function sample (see func.R) to estimate the sample size per treatment and species.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    n_list[[paste(hormone[k], temperature[l], sep = "_")]] <- sample(hormone[k], temperature[l])
-  }
-}
-```
 
@@
 
-```{r, models}
-#| label: models
+```{r, models_mitochondrial}
+#| label: models_mitochondrial
 # Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder).
@@
 # 
-# df got from data_process.R
-#
-fchoice <- bf(choice_slope_mean + choice_slope_se ~ cort*temp)
-ferrors <- bf(errors_slope_mean + errors_slope_se ~ cort*temp)
-fROS <- bf(ROS ~ Mit_density + cort*temp)
-#
-refit <- FALSE
-if(refit==TRUE){
-  # Choice
-  model_choice_2 <- brm(fchoice + fROS,
-                data = clean_df,
-                family = gaussian,
-                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Errors
-  model_errors_2 <- brm(ferrors + fROS,
-                data = clean_df,
-                family = gaussian,
-                chains = 4, cores = 4, iter = 8000, warmup = 2000,
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Write the models to two files
-  saveRDS(model_choice_2, file = paste0(here("output/models/model_choice_2.rds")))
-  saveRDS(model_errors_2, file = paste0(here("output/models/model_errors_2.rds")))
-  } else {
-    # Read the model from a file
-    model_choice_2 <- readRDS(file = paste0(here("output/models/model_choice_2.rds")))
-    model_errors_2 <- readRDS(file = paste0(here("output/models/model_errors_2.rds")))
+#
+# Run models mitochondrial physiology 
+var_m <- c("mit_density", "mit_potential", "ROS", "DNAdamage", "peroxidation")
+for (p in var_m){ 
+  if (p %in% c("mit_density", "mit_potential", "ROS", "DNAdamage")){
+    formula <- paste0(p, "~ cort*temp + (1|clutch)")
+  } else if (p == "peroxidation"){
+    formula <- paste0(p, "~ cort*temp + age + (1|clutch)")
   } 
-post_gen <- as_draws_df(model_errors_2)
-write.csv(post_gen, file= "./output/Checking/post_gen.csv")
-```
-
-```{r, resultsRev}
-# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
-## 1) L. delicata
-### Group = red
-drr_CORTCold <- deli_rev_red$b_trial_reversal
-drr_ControlCold <- (deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$b_trial_reversal)
-drr_CORTHot <- (deli_rev_red$'b_trial_reversal:tempHot' + deli_rev_red$b_trial_reversal)
-drr_ControlHot <- (deli_rev_red$'b_trial_reversal:cortControl:tempHot' + deli_rev_red$b_trial_reversal + deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$'b_trial_reversal:tempHot')
-### Group = blue
-drb_CORTCold <- deli_rev_blue$b_trial_reversal
-drb_ControlCold <- (deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$b_trial_reversal)
-drb_CORTHot <- (deli_rev_blue$'b_trial_reversal:tempHot' + deli_rev_blue$b_trial_reversal)
-drb_ControlHot <- (deli_rev_blue$'b_trial_reversal:cortControl:tempHot' + deli_rev_blue$b_trial_reversal + deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$'b_trial_reversal:tempHot')
-## 2) L. guichenoti
-### Group = red
-grr_CORTCold <- guich_rev_red$b_trial_reversal
-grr_ControlCold <- (guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$b_trial_reversal)
-grr_CORTHot <- (guich_rev_red$'b_trial_reversal:tempHot' + guich_rev_red$b_trial_reversal)
-grr_ControlHot <- (guich_rev_red$'b_trial_reversal:cortControl:tempHot' + guich_rev_red$b_trial_reversal + guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$'b_trial_reversal:tempHot')
-### Group = blue
-grb_CORTCold <- guich_rev_blue$b_trial_reversal
-grb_ControlCold <- (guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$b_trial_reversal)
-grb_CORTHot <- (guich_rev_blue$'b_trial_reversal:tempHot' + guich_rev_blue$b_trial_reversal)
-grb_ControlHot <- (guich_rev_blue$'b_trial_reversal:cortControl:tempHot' + guich_rev_blue$b_trial_reversal + guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$'b_trial_reversal:tempHot')
-```
-
-```{r, tbl-data}
-#| label: tbl-data
-#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group.
-Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean.
-All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero.
-In bold, those values that are significant (p-value <0.05)"
-source(here("R", "func.R"))
-#
-############################## CREATING BIG DF FOR TABLE ##############################
-# Building the vectors for titles of rows and columns
-specie <- c("L. delicata", "L. guichenoti")
-groups <- c("Red", "Blue")
-test <- c("Associative", "Reversal")
-treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
-values <- c("Mean", "95% CI", "p-value")
-# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors.
-p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
-#First get estimates for both tasks
-estimates_asso <- list(
-  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
-  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
-  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
-  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
-)
-#
-estimates_rev <- list(
-  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
-  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
-  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
-  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
-)
-# Then get the mean, co.intervals(95%), and p-values
-asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
-asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
-asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
-asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
-asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
-#
-rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
-rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
-rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
-rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
-rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
-#
-# Building the df for the associative
-asso_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(asso_mean, each = 1),
-  CI = rep(asso_intervals, each = 1),
-  PValue = rep(asso_pvalue, each = 1),
-  Task = rep("Associative", length(asso_mean))
-)
-# Building the df for the reversal
-rev_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(rev_mean, each = 1),
-  CI = rep(rev_intervals, each = 1),
-  PValue = rep(rev_pvalue, each = 1),
-  Task = rep("Reversal", length(rev_mean))
-)
-# Joining both dfs
-table_data <- rbind(asso_df, rev_df)
-table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
-#
-############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
-# Make n_list into a df
-n_df <- as.data.frame(do.call(rbind, n_list)) %>%
-  rename("n" = V1) %>%
-  rownames_to_column("model") %>%
-  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
-  unite("Treatment", c("cort", "temp"), sep = "-") %>%
-  mutate(Specie = factor(Specie,
-                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
-        Treatment = factor(Treatment,
-                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
-# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
-new_table_data <- merge(table_data, n_df) %>%
-  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
-  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
-  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
-  mutate(Specie = factor(Specie,
-                  levels = c("L. delicata", "L. guichenoti")),
-        Group = factor(Group,
-                  levels = c("Red", "Blue")),
-        Treatment = factor(Treatment, 
-                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
-  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
-  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
-  mutate(Treatment = paste0(Treatment, ")"))
-write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
-#
-############################## MAKING THE TABLE ##############################
-## Table format
-set_flextable_defaults(
- font.family = "Times New Roman",
- fint.size = 10)
-# Split the table_data df by task
-real_table <- flextable(new_table_data) %>%
-    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
-    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
-    set_table_properties(width = 1) %>%
-    align(align="center", part="all") %>% 
-    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
-    set_header_labels(Mean_Associative = "Mean",
-                      `95% CI_Associative` = "95% CI",
-                      `p-value_Associative` = "p-value",
-                      Mean_Reversal = "Mean",
-                      `95% CI_Reversal` = "95% CI",
-                      `p-value_Reversal` = "p-value") %>%
-    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
-    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
-    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
-    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
-    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
-    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
-    autofit() 
-real_table
+  pmodel_name <- paste0("m_def_", p)
+  assign(pmodel_name, fit_m(df = clean_df,
+                          cat = "def",
+                          var = p,
+                          formula = formula,
+                          fam = gaussian(),
+                          refit = FALSE),
+        envir = .GlobalEnv)  # Assign to the global environment
+} 
+```
+
+```{r, models_learning}
+#| label: models_learning
+#
+formula_learn <- errors ~ day*cort*temp + (1 + day|lizard_id) + (1|clutch)
+m_def_learn <- fit_m(df = learning_df,
+                      cat = "def",
+                      var = "learning",
+                      formula = formula_learn,
+                      fam = negbinomial(link = "log"),
+                      refit = FALSE)
+```
+
+```{r, learning_results_names}
+#| label: learning_results_names
+# Rename some of the posteriors and make new estimates for the learning rate 
+#
+### Slopes
+slope_ControlCold <- m_def_learn$b_day
+slope_CORTCold <- m_def_learn$b_day + m_def_learn$`b_day:cortCORT`
+slope_ControlHot <- m_def_learn$b_day + m_def_learn$`b_day:tempHot`
+slope_CORTHot <- m_def_learn$b_day + m_def_learn$`b_day:cortCORT:tempHot` + m_def_learn$`b_day:cortCORT` + m_def_learn$`b_day:tempHot`
+### Intercepts (for figure)
+int_ControlCold <- m_def_learn$b_Intercept
+int_CORTCold <- m_def_learn$b_Intercept + m_def_learn$`b_cortCORT`
+int_ControlHot <- m_def_learn$b_Intercept + m_def_learn$`b_tempHot`
+int_CORTHot <- m_def_learn$b_Intercept + m_def_learn$`b_cortCORT:tempHot` + m_def_learn$`b_cortCORT` + m_def_learn$`b_tempHot`
 ```
@@
 #| label: fig-learning
-#| fig.cap: "Results for L.
-delicata (A,B) and L.
-guichenoti (C, D).
-Panels A and C show the predicted probability of choosing the correct feeder first over trials.
-The lines represent the mean predicted probability of choosing the correct feeder first on each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained by using the slope and intercept estimates from the posterior distributions.
-The different colours indicate the different treatments.
-Panels B and D show the distribution of the estimates of slopes per each treatment.
-The x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
-The different colours indicate the different treatments.
-Points and bars represent the mean and standard deviation of the mean of the estimates, respectively."
-#
-#
-####### A) Probability of choosing correct figure
-## A.1) Get the df with the estimated probability of choosing the correct feeder first over trials (using learning_df from data_process.R)
-# Get the df for the intercepts
+#| fig.cap: "Learning results"
 source(here("R", "func.R"))
-post_choice_intercept <- tidy_post_df("choice", "intercept")
-# Merge slope and intercept dfs for choice 
-learning_df_prob <- merge(learning_df, post_choice_intercept, by = "lizard_id", all = TRUE)
-# Create a vector with treatments
-treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
-# Create new matrix and new dfs
-treat_matrix <- matrix(NA, nrow = 20, ncol = 40)
-fig_df <- data.frame()
-# Loop through treatments
-for(t in treatments){
-  if(t == "CORT-Cold"){
-    df <- learning_df_prob %>%
-      filter(cort=="CORT", temp=="Cold") %>%
+#
+#
+####### A) Create df
+# Slopes df (employed in fig-learning B)
+slope_list <- list(`Control-Cold` = slope_ControlCold,
+                 `CORT-Cold` = slope_CORTCold,
+                 `Control-Hot` = slope_ControlHot,
+                 `CORT-Hot` = slope_CORTHot)
+data_fig_learning_slopes <- do.call(rbind, lapply(names(slope_list), function(x) {
+  data.frame(treatment = x, slopes = slope_list[[x]])
+}))
+# Intercepst df (only for fig-learning A)
+intercept_list <- list(`Control-Cold` = int_ControlCold,
+                 `CORT-Cold` = int_CORTCold,
+                 `Control-Hot` = int_ControlHot,
+                 `CORT-Hot` = int_CORTHot)
+data_fig_learning_intercepts <- do.call(rbind, lapply(names(intercept_list), function(x) {
+  data.frame(treatment = x, intercepts = intercept_list[[x]])
+})) %>% 
+  group_by(treatment) %>%
+  summarize(se_int = sd(intercepts)/sqrt(length(intercepts)),
+            intercepts = mean(intercepts))
+# df fig-learning A
+data_fig_learning_slopesA <- data_fig_learning_slopes %>%
+  group_by(treatment) %>%
+  summarize(se_slope = sd(slopes)/sqrt(length(slopes)),
+            slopes = mean(slopes),)
+data_fig_learningB <- merge(data_fig_learning_slopesB, data_fig_learning_intercepts, by = "treatment")
+treatment <- unique(data_fig_learningB$treatment)
+fig_A_df <- data.frame()
+for(t in treatment){ # Loop per treatment
+  df <- data_fig_learningB %>%
+    filter(treatment == t) %>%
     data.frame()
-  } else if(t == "Control-Cold"){
-    df <- learning_df_prob %>%
-      filter(cort=="Control", temp=="Cold") %>%
-      data.frame()
-  } else if(t == "CORT-Hot"){
-    df <- learning_df_prob %>%
-      filter(cort=="CORT", temp=="Hot") %>%
-      data.frame()
-  } else if(t == "Control-Hot"){
-    df <- learning_df_prob %>%
-      filter(cort=="Control", temp=="Hot") %>%
-      data.frame()
-  } else {
-    stop("loop wrong")
-  }
   # Variables selected
-  m <- df$choice_slope__mean
-  u <- df$choice_intercept__mean
+  m <- df$slopes
+  u <- df$intercepts
   # Loop per treatment
@@
   for(x in 0:40){
-    for(j in 1:num_individuals){
-      value <- exp(u[j] + m[j] * x) / (1 + exp(u[j] + m[j] * x))
-      treat_matrix[j, x] <- value
+    if (x == 0){
+      standard <- df$se_int
+    } else {
+      standard <- df$se_slope
+    }
+    value <- exp(u + m * x)
+    temp_df <- data.frame(trial = rep(x, length(value)),
+                          treatment = rep(t, length(value)),
+                          errors = value,
+                          se = standard)
+    fig_A_df <- rbind(fig_A_df, temp_df)
   }
 }
-  treat_df <- as.data.frame(treat_matrix)
-  colnames(treat_df) <- paste0("X", 1:40)  # Adjust column names
-  treat_df <- gather(treat_df, key = "Trial", value = "Value")  # Reshape data frame
-  treat_df$Treatment <- t
-  fig_df <- rbind(fig_df, treat_df)
-}
-# Transform the database for better use
-fig_choice_df <- fig_df %>%
-  mutate(Trial = gsub("X", "", Trial)) %>%
-  mutate(Trial = as.numeric(Trial)) %>%
-  group_by(Trial, Treatment) %>%
-  summarize(
-    Mean_Predicted_prob = mean(Value),
-    SE_Predicted_prob = sd(Value)/sqrt(length(Value))
-    ) %>%
-  ungroup() %>%
-  mutate(
-    Treatment = factor(Treatment,
-                        levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")),
-    ) %>%
-data.frame()
-#
-## A.2) Make the plot for the probability of choosing the correct feeder first over trials
-fig_prob_choice <- ggplot(fig_choice_df, aes(x = Trial, y = Mean_Predicted_prob, color = Treatment)) +
-  geom_line(linewidth = 1) +
-  scale_color_manual(values = c("CORT-Cold"="#00008B", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +
-  geom_ribbon(aes(ymin = Mean_Predicted_prob - SE_Predicted_prob, ymax = Mean_Predicted_prob + SE_Predicted_prob, fill = Treatment), color = NA, alpha = 0.075) + 
-  scale_fill_manual(values = c("CORT-Cold"="darkblue", "Control-Cold"="#68bde1", "CORT-Hot"="#b50101", "Control-Hot"="#fa927d")) +
-  theme_classic() +
-  labs(y = "Predicted probability of correct choice", x = "Trial") +
-  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
-  theme(
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
-    legend.position = "none"
-  )
-#
-## A.3) Make the the plot for the mean slope per treatment using the probability of choosing the correct feeder
-fig_slope_choice <- plot_slopes("choice")
-#
-#
-## A.4) Combine plots of probability and slopes to have the "choice" figure
-fig_results_choice <- plot_grid(fig_prob_choice, fig_slope_choice, nrow = 1) 
-ggsave("./output/figures/fig_results_choice.png", plot=fig_results_choice, width = 25, height = 15, units = "cm", dpi = 600)
-#
-#
-#
-####### B) Errors figure
-## B.1) Plot the errors using data_spal
+# 
+#### C) Plot the fig-learning A and fig-learning B separately 
+fig_learningA <- plot_errors(fig_A_df)
+fig_learningB <- plot_slopes(data_fig_learning_slopes)
+#
+#
+
 errors_df <- data_spal %>%
@@
   data.frame()
-fig_errors <- ggplot(errors_df, aes(x = day, y = mean_errors, color = Treatment, fill = Treatment)) +
-  geom_smooth(method = "lm", se = TRUE, linewidth = 1, alpha = 0.075) +
+fig_errors <- ggplot(learn_df, aes(x = Trial, y = Value, color = Treatment, fill = Treatment)) +
+  geom_smooth(se = TRUE, linewidth = 1, alpha = 0.075) +
   scale_color_manual(values = c("CORT-Cold" = "#00008B", "Control-Cold" = "#68bde1", "CORT-Hot" = "#b50101", "Control-Hot" = "#fa927d")) +
@@
 # Suplementary Material
-#### Results of preliminary models
-
-```{r, models_preliminary}
-#| label: models_preliminary
-# Fitting preliminary models to see if sex and age are relevant for our models
-source(here("R", "func.R"))
-#
-#
-var_m <- c("mit_density", "mit_potential", "ROS", "DNAdamage", "peroxidation")
-formula_list_ <- list()
-for (p in var_m){
-  formula_list_[[p]] <- paste0(p, "~ cort*temp + age + sex + (1|clutch)")
-  pmodel_name <- paste0("m_prel_", p)
-  assign(pmodel_name, fit_m(df = clean_df,
-                             cat = "prel",
-                             var = p,
-                             formula = formula_list_[[p]],
-                             fam = gaussian(),
-                             refit = FALSE),
-          envir = .GlobalEnv)  # Assign to the global environment
-}
-#
-#
-# Run model behaviour (stimuli separated)
-formula_learn <- slope_mean + slope_sd ~ cort*temp + age + sex + (1|clutch)
-m_prel_learn <- fit_m(df = clean_df,
-                      cat = "prel",
-                      var = "slope_mean",
-                      formula = formula_learn,
-                      fam = gaussian(),
-                      refit = FALSE)
-```
-
-Table S1. Preliminary results of the models testing for Mitochondrial Density.
-
-```{r, results_preliminary_mitdensity}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density"
-#| label: results_preliminary_mitdensity
-#
-sum_mitdensity_prel <- m_prel_mit_density %>%
-  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
-  summarise_draws()  %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_mitdensity_prel)
-```
-
-Model formula: mit_density ~ cort * temp + age + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S2.  Preliminary results of the models testing for Mitochondrial Potential.
-```{r, results_preliminary_potential}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential"
-#| label: results_preliminary_potential
-#
-sum_potential_prel <- m_prel_mit_potential %>%
-  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_potential_prel)
-```
-
-Model formula: mit_potential ~ cort * temp + age + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S4. Preliminary results of the models testing for ROS.
-```{r, results_preliminary_ros}
-#| tbl-cap: "Preliminary results of the models testing for ROS"
-#| label: results_preliminary_ros
-#
-sum_m_ros_prel <- m_prel_ROS %>%
-  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_ros_prel)
-```
-
-Model formula: ROS ~ cort * temp + age + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S6. Preliminary results of the models testing for DNA damage.
-```{r, results_preliminary_dnadamage}
-#| tbl-cap: "Preliminary results of the models testing for DNA damage"
-#| label: results_preliminary_dnadamage
-#
-sum_m_dnadamage_prel <- m_prel_DNAdamage %>%
-  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_dnadamage_prel)
-```
-
-Model formula: DNAdamage ~ cort * temp + age + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values. Summary indicates an effect of sex and age, so they were included in the final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S8. Preliminary results of the models testing for lipid peroxidation.
-```{r, results_preliminary_peroxidation}
-#| tbl-cap: "Preliminary results of the models testing for lipid peroxidation"
-#| label: results_preliminary_peroxidation
-#
-sum_m_peroxidation_prel <- m_prel_peroxidation %>%
-  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
-  summarise_draws() %>%
-  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
-#
-flextable(sum_m_peroxidation_prel)
-```
-
-Model formula: peroxidation ~ cort * temp + age + sex + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values.
-Summary indicates no effect of sex, so it was discarded from the final models.
-However, we saw an effect of age and we included it in our final models.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
 
@@
 
+```{r, tbl-contrasts}
+#| label: tbl-contrasts
+#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group.
+Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean.
+All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero.
+In bold, those values that are significant (p-value <0.05)"
+source(here("R", "func.R"))
+#
+############################## CREATING BIG DF FOR TABLE ##############################
+# Building the vectors for titles of rows and columns
+specie <- c("L. delicata", "L. guichenoti")
+groups <- c("Red", "Blue")
+test <- c("Associative", "Reversal")
+treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
+values <- c("Mean", "95% CI", "p-value")
+# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors.
+p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
+#First get estimates for both tasks
+estimates_asso <- list(
+  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
+  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
+  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
+  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
+)
+#
+estimates_rev <- list(
+  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
+  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
+  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
+  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
+)
+# Then get the mean, co.intervals(95%), and p-values
+asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
+asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
+asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
+asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
+asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
+#
+rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
+rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
+rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
+rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
+rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
+#
+# Building the df for the associative
+asso_df <- data.frame(
+  Specie = rep(specie, each = length(groups) * length(treatments)),
+  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
+  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
+  Mean = rep(asso_mean, each = 1),
+  CI = rep(asso_intervals, each = 1),
+  PValue = rep(asso_pvalue, each = 1),
+  Task = rep("Associative", length(asso_mean))
+)
+# Building the df for the reversal
+rev_df <- data.frame(
+  Specie = rep(specie, each = length(groups) * length(treatments)),
+  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
+  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
+  Mean = rep(rev_mean, each = 1),
+  CI = rep(rev_intervals, each = 1),
+  PValue = rep(rev_pvalue, each = 1),
+  Task = rep("Reversal", length(rev_mean))
+)
+# Joining both dfs
+table_data <- rbind(asso_df, rev_df)
+table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
+#
+############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
+# Make n_list into a df
+n_df <- as.data.frame(do.call(rbind, n_list)) %>%
+  rename("n" = V1) %>%
+  rownames_to_column("model") %>%
+  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
+  unite("Treatment", c("cort", "temp"), sep = "-") %>%
+  mutate(Specie = factor(Specie,
+                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
+        Treatment = factor(Treatment,
+                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
+# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
+new_table_data <- merge(table_data, n_df) %>%
+  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
+  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
+  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
+  mutate(Specie = factor(Specie,
+                  levels = c("L. delicata", "L. guichenoti")),
+        Group = factor(Group,
+                  levels = c("Red", "Blue")),
+        Treatment = factor(Treatment, 
+                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
+  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
+  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
+  mutate(Treatment = paste0(Treatment, ")"))
+write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
+#
+############################## MAKING THE TABLE ##############################
+## Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ fint.size = 10)
+# Split the table_data df by task
+real_table <- flextable(new_table_data) %>%
+    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
+    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
+    set_table_properties(width = 1) %>%
+    align(align="center", part="all") %>% 
+    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
+    set_header_labels(Mean_Associative = "Mean",
+                      `95% CI_Associative` = "95% CI",
+                      `p-value_Associative` = "p-value",
+                      Mean_Reversal = "Mean",
+                      `95% CI_Reversal` = "95% CI",
+                      `p-value_Reversal` = "p-value") %>%
+    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
+    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
+    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
+    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
+    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
+    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
+    autofit() 
+real_table
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Results of preliminary models
+
+```{r, models_preliminary}
+#| label: models_preliminary
+# Fitting preliminary models to see if sex and age are relevant for our models
+source(here("R", "func.R"))
+#
+#
+var_m <- c("mit_density", "mit_potential", "ROS", "DNAdamage", "peroxidation")
+formula_list_ <- list()
+for (p in var_m){
+  formula_list_[[p]] <- paste0(p, "~ cort*temp + age + sex + (1|clutch)")
+  pmodel_name <- paste0("m_prel_", p)
+  assign(pmodel_name, fit_m(df = clean_df,
+                             cat = "prel",
+                             var = p,
+                             formula = formula_list_[[p]],
+                             fam = gaussian(),
+                             refit = FALSE),
+          envir = .GlobalEnv)  # Assign to the global environment
+}
+#
+#
+# Run model learning including the effect of sex and interaction
+formula_learn <- errors ~ day*cort*temp + sex + age + (1 + day|lizard_id) + (1|clutch)
+m_prel_learn <- fit_m(df = learning_df,
+                      cat = "prel",
+                      var = "learning",
+                      formula = formula_learn,
+                      fam = negbinomial(link = "log"),
+                      refit = FALSE)
+```
+
+Table S1. Preliminary results of the models testing for Mitochondrial Density.
+
+```{r, results_preliminary_mitdensity}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Density"
+#| label: results_preliminary_mitdensity
+#
+sum_mitdensity_prel <- m_prel_mit_density %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws()  %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_mitdensity_prel)
+```
+
+Model formula: mit_density ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S2.  Preliminary results of the models testing for Mitochondrial Potential.
+```{r, results_preliminary_potential}
+#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential"
+#| label: results_preliminary_potential
+#
+sum_potential_prel <- m_prel_mit_potential %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_potential_prel)
+```
+
+Model formula: mit_potential ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S3. Preliminary results of the models testing for ROS.
+```{r, results_preliminary_ros}
+#| tbl-cap: "Preliminary results of the models testing for ROS"
+#| label: results_preliminary_ros
+#
+sum_m_ros_prel <- m_prel_ROS %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_ros_prel)
+```
+
+Model formula: ROS ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S4. Preliminary results of the models testing for DNA damage.
+```{r, results_preliminary_dnadamage}
+#| tbl-cap: "Preliminary results of the models testing for DNA damage"
+#| label: results_preliminary_dnadamage
+#
+sum_m_dnadamage_prel <- m_prel_DNAdamage %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_dnadamage_prel)
+```
+
+Model formula: DNAdamage ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S5. Preliminary results of the models testing for lipid peroxidation.
+```{r, results_preliminary_peroxidation}
+#| tbl-cap: "Preliminary results of the models testing for lipid peroxidation"
+#| label: results_preliminary_peroxidation
+#
+sum_m_peroxidation_prel <- m_prel_peroxidation %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_peroxidation_prel)
+```
+
+Model formula: peroxidation ~ cort * temp + age + sex + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+Summary indicates no effect of sex, so it was discarded from the final models.
+However, we saw an effect of age and we included it in our final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S6. Preliminary results of the models testing for learning.
+```{r, results_preliminary_learning}
+#| tbl-cap: "Preliminary results of the models testing for lipid peroxidation"
+#| label: results_preliminary_peroxidation
+#
+sum_m_learn_prel <- m_prel_learn %>%
+  dplyr::select(starts_with("b_")) %>%  # Removes random effect terms
+  summarise_draws() %>%
+  mutate(across(where(is.numeric), ~ as.numeric(format_dec(.x, 3))))
+#
+flextable(sum_m_learn_prel)
+```
+
+Model formula: errors ~ day * cort * temp + sex + age+ (1 + day | lizard_id) + (1|clutch).
+Model convergence was checked through rhat and ess_bulk values.
+**Summary indicates no effect of sex, so it was discarded from the final models.
+However, we saw an effect of age and we included it in our final models.**
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Final models diagnostics (plots)

### 2025-03-06 .vdoc.r

@@ -117,262 +117,119 @@ knitr::include_graphics("./Others/SPAL_METH.svg")
 #
 #
 #
-#| label: sampleSize
-# List with the sample sizes from the database (here("output/databases_clean/data_complete.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    n_list[[paste(hormone[k], temperature[l], sep = "_")]] <- sample(hormone[k], temperature[l])
-  }
-}
 #
 #
 #
-#
-#
-#| label: models
+#| label: models_mitochondrial
 # Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
 source(here("R", "func.R"))
 # 
-# df got from data_process.R
 #
-fchoice <- bf(choice_slope_mean + choice_slope_se ~ cort*temp)
-ferrors <- bf(errors_slope_mean + errors_slope_se ~ cort*temp)
-fROS <- bf(ROS ~ Mit_density + cort*temp)
-#
-refit <- FALSE
-if(refit==TRUE){
-  # Choice
-  model_choice_2 <- brm(fchoice + fROS,
-                data = clean_df,
-                family = gaussian,
-                chains = 4, cores = 4, iter = 8000, warmup = 2000, 
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Errors
-  model_errors_2 <- brm(ferrors + fROS,
-                data = clean_df,
-                family = gaussian,
-                chains = 4, cores = 4, iter = 8000, warmup = 2000,
-                control = list(adapt_delta = 0.99, max_treedepth = 12))
-  # Write the models to two files
-  saveRDS(model_choice_2, file = paste0(here("output/models/model_choice_2.rds")))
[…]

### 2025-03-06 R/data_process.R

@@ -16,8 +16,8 @@ learning_df <- read.csv(here("./data/Spatial_learn.csv")) %>%
       levels = c("23", "28"),
       labels = c("23" = "Cold", "28" = "Hot"))) %>%
     mutate(cort = factor(cort,
-      levels = c("B", "A"),
-      labels = c("B" = "CORT", "A" = "Control"))) %>%
+      levels = c("A", "B"),
+      labels = c("A" = "Control", "B" = "CORT"))) %>%
   ungroup() %>%
 data.frame()
 write.csv(learning_df, here("./output/databases_clean/learning_df.csv"))
@@ -26,7 +26,7 @@ write.csv(learning_df, here("./output/databases_clean/learning_df.csv"))
 #
 ## Fit the model only if it has not been fit yet (if refit = TRUE)
 if(refit){
-  model_errors <- brm(errors ~ 1 + (1 + day|lizard_id),
+  model_errors <- brm(errors ~ day + (1 + day|lizard_id),
                 data = learning_df,
                 negbinomial(link = "log"),
                 chains = 4, cores = 4, iter = 8000, warmup = 2000,

### 2025-03-06 R/exploratory.R

@@ -115,3 +115,18 @@ ggsave(here("./output/figures/exploratory/fig_qqplots.png"), final_qq_plot, widt
 #
 # Log-Normal: DNA Damage, Lipid Peroxidation
 # Normal: Mitochondrial Density, Mitochondrial Potential, ROS Production, Learning slopes 
+#
+#
+# C) Check overdisperssion for errors
+mean_errors <- mean(learning_df$errors, na.rm = TRUE)
+var_errors <- var(learning_df$errors, na.rm = TRUE)
+overdispersion_ratio <- var_errors / mean_errors # if this is > 1, there is overdispersion, so we need a negative binomial
+#
+# Overdispersion ratio is 4.152071, so we can't use a Poisson model
+# Can we fit neg_binomial?
+zeroes <- sum(learning_df$errors == 0, na.rm = TRUE)
+n <- length(learning_df$errors)
+prop_zeroes <- zeroes / n
+#
+# Proportion of zeroes is 0.2414557 so we should use a zero-inflated model ---> fam = zero_inflated_negbinomial()
+#

### 2025-03-06 R/func.r

@@ -89,27 +89,6 @@ qq_plots_single <- function(df, vle, label) {
     
   return(fig_qq)
 }
-###################
-###################
-# Tidy estimates from fit_m
-#' @title tidy_post
-#' @description The df obtained by fit_m contains the estimates of the Intercept and fixed effects of "Choice ~ trial (1 + lizard)"
-#' for each treatment, species, and group,but it extracts the estimates of the four chains, each chain in a column, here we want to
-#' tidy up to two columns (Intercept and trial), plus the ones indicating species, group, tand treatment
-#' @param df Dataframe used
-#' @return Same df but tidy
-tidy_post <- function(df) {
-  # Select data
-  data <- df
-  # Split original df into four by chain, and give the columns a common name
-  res1 <- data%>%select(matches("^X1.b_"),treatment_level, species_level,bias_level)%>%rename_all(~sub("^X1.b_", "", .))
-  res2 <- data%>%select(matches("^X2.b_"),treatment_level, species_level,bias_level)%>%rename_all(~sub("^X2.b_", "", .))
-  res3 <- data%>%select(matches("^X3.b_"),treatment_level, species_level,bias_level)%>%rename_all(~sub("^X3.b_", "", .))
-  res4 <- data%>%select(matches("^X4.b_"),treatment_level, species_level,bias_level)%>%rename_all(~sub("^X4.b_", "", .))
-  # Bind data again
-  new_df <- bind_rows(res1, res2,res3,res4)
-return(new_df) 
-}
 ####################
 ####################
 # Estimate pmcm
@@ -160,37 +139,10 @@ format_p <- function(x, n) {
 ####################
 # Function to create the plot for the slopes
 #' @title plot_slopes
-#' @param type to select the df for the violin plot
-plot_slopes <- function(type){
-  if(type == "choice") {
-    data <- posteriors_choice
-    label_df <- "choice_"
-  } else if(type == "errors") {
-    data <- posteriors_errors
-    label_df <- "errors_"
-  } else {
-    stop("df not valid")
-  }
-  label_effect <- "day"
-  pattern <- "^r_lizard_id\\[\\d+,day\\]$"
-  tidy_post <- data %>%
-    select(matches(pattern)) %>%
-    pivot_longer(cols = everything(),
-              names_to = "lizard_id", 
[…]

### 2025-03-10 ms.qmd
updates

@@
 
+```{r, sampleSize}
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+df_damage <- clean_df %>%
+  filter(DNAdamage != "NA")
+n_damage <- list()
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_damage[[list_name]] <- sample(df = df_damage, corti = hormone[k], therm = temperature[l])
+  }
+}
+```
+
 ## Introduction
-**Corticosterone is the major GC produced by
-adrenal steroidogenic cells in reptiles, and an increase in
-its plasma concentration is a frequent indicator of stress
-(Anderson et al. 2014 General and Comparative Endocrinology). In nature,
-GCs in the developing embryos seem to be largely of
-maternal origin (Uller et al. 2009, The Journal of Experimental Biology) and are probably transferred to the growing oocytes through vitellogenesis (Elf et al.,
-2002, 2003 General and Comparative Endocrinology (both)).**
-
-Mechanistically, a change in body size requires
-a change in cell number and/or cell size, and there is evidence that
-body size and cell size undergo coordinated evolutionary (Arendt,
-2007; Starostová et al., 2005; Stevenson et al., 1995) or
-phenotypically plastic changes (Arendt, 2007; Czarnoleski et al.,
-2013; Hessen et al., 2013).
+
+Cognition encompasses processes such as perception, learning, memory, and decision-making, which allow animals to gather, retain, and use environmental information [@shettleworth].
+These processes are fundamental for survival and reproduction, influencing behaviours like foraging, mate selection, or antipredatory strategies [@dukas_evolutionary_2004].
+Learning - the acquisition and consolidation of new information [@dukas_evolutionary_2004] - enables individuals to create new associations between events, which can be essential for coping with environmental changes [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
+However, the capacity to form these associations varies among individuals, potentially affecting their responses to environmental challenges [@ward2016ecological; @welklin2024spatial].
+Animals that learn faster may exploit novel resources better or may start responding to new threats more efficiently than slow learners [@ward2016ecological].
+Consequently, individual differences in learning abilities can have profound effects on survival and reproductive output, ultimately influencing population growth rates and stability [@ward2016ecological; @welklin2024spatial].
+Understanding the mechanisms that drive these differences is therefore crucial for predicting how populations will respond to novel conditions.
+
+Factors like age, sex, or early- life conditions can have important effects on learning abilities [@szuran_water_1994; @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @carazo_sex_2014; @noble_age-dependent_2014].
+Early developmental environment is particularly influential, as the brain is highly sensitive to environmental inputs during this period [@zhu_prenatal_2004].
+Early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
+Among the physiological mechanisms underlying learning, mitochondrial activity has emerged as a key factor influencing brain function [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+The neural processes involved learning impose substantial energetic demands [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], making efficient mitochondrial respiration essential [@du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+Additionally, learning depends on a dense network of functional neurons [@lefebvre_taxonomic_2011; @amiel_effects_2017], which can be compromised by excessive oxidative stress byproduct of mitochondrial function [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+Studies on domestic rats (_Rattus norvegicus_) have shown that the effects of early environment on mitochondrial physiology can have profound consequences for cognitive abilities [@zhu_prenatal_2004; @cao2014ampk].
+However, the extent to which these effects are generalizable to other taxa, particularly ectotherms, remains largely unexplored.
+
+Since mitochondria are maternally inherited, maternal condition plays a fundamental role in shaping offspring mitochondrial activity [@picard_mitochondria_2014].
+In particular, maternal stress can strongly influence how mitochondria operate in the offspring [@zhu_prenatal_2004].
+Stressful situations elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryo [@uller_sex-specific_2009], altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) [@zhu_prenatal_2004; @haussmann_embryonic_2012].
+
+Temperature can be a significant source of maternal stress in ectotherms.
+Thermal environments outside of the optimal range can elevate GCs in the mother, which can be passed to the offspring [see @Crino_2023].
+In addition, temperature can also directly influence offspring development, particularly during early life [@noble_developmental_2018; @crino2024eggs].
+The prenatal thermal environment plays a crucial role in shaping mitochondrial function, affecting energy metabolism and oxidative stress [@stier2022experimental; @crino2024eggs].
+For instance, hatchlings of the red-eared slider turtle (_Trachemys scripta elegans_) incubated at 26.5 ºC exhibited greater oxidative damage than those incubated at temperatures 0.7 to 1.2 ºC warmer [@treidel2016temperature].
+Thus, the combined effects of prenatal GCs and temperature may profoundly influence mitochondrial function, with important consequences for brain development and cognition.
+However, the extent to which prenatal GCs and temperature interact to shape cognitive abilities via mitochondrial physiology remains largely unknown.
+
+Here, we examined how prenatal temperature and corticosterone (CORT) — the primary GC in reptiles — affect brain mitochondrial physiology and spatial learning in the delicate skink (_Lampropholis delicata_).
+We hypothesized that prenatal CORT and temperature would influence learning abilities through their effects on mitochondrial activity in the brain.
+Specifically, we predicted that prenatal CORT would decrease energy production while increasing oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
+Similarly, high temperatures would decrease mitochondrial efficienty [@zavorka_climate_2021; @crino2024eggs], but with the potential to reduce oxidative stress [@treidel2016temperature].
+We further predicted that the combined effects of CORT and temperature would lead to complex interactions, with both factors negatively impacting mitochondrial efficiency but having opposite effects on oxidative stress.
+These effects would lead to differences in learning abilities, that could be affected by the balance between energy production and oxidative stress [@du_dynamic_2009; @picard_energetic_2018; @alexandrov_neuronal_2022].
+By examining these interactions, we aim to clarify how prenatal environmental factors shape learning abilities through mitochondrial function, shedding light on the role of early-life conditions in cognitive development and its mechanismsm.
 
@@
 #### Animal husbandry  
-*Breeding colony* - We tested _L.
-delicata_ coming from a breeding colony established in the laboratory since 2019.
+*Breeding colony* - Juveniles of _L.
+delicata_ came from a breeding colony established in the laboratory since 2019.
 This colony consisted of 270 adults housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure.
-Enclosures were provided with nonstick matting, shelter, and several small water dishes.
-Water was given daily, and the lizards were fed approx.
-40 mid-size crickets (_Acheta domestica_) per enclosure three days a week.
-Crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
+Enclosures were provided with shelter, nonstick matting, and several small water dishes.
+The lizards were fed approx.
+40 mid-size crickets (_Acheta domestica_) per enclosure three days a week, and water was given daily.
+The crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
 Room temperatures were set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
 
-*Eggs collection and incubation* - Between mid-November 2023 and mid-January 2024, we provide females with a place to lay the eggs by placing a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures.
+*Eggs collection and incubation* - Between mid-November 2023 to mid-January 2024, we placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures to provide females with a place to lay the eggs.
 These boxes were checked three days a week.
@@
 Lizards were measured and weighed immediately after hatching.
-Snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a digital scale with an accuracy of ± 0.001 g.
+snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a digital scale with an accuracy of ± 0.001 g.
 Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with nonstick matting and a small water dish.
@@
 #### CORT and Temperature manipulation
-To test the interactive effects of CORT and incubation temperature, we manipulated CORT concentrations in eggs and incubated them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig).
-Eggs were assigned to one fo the two hormone treatments: CORT or Control.
-CORT-treated lizards were topically supplied with 5 µL of crystalline corticosterone (Sigma, Cat.
+We tested the combined effects of prenatal CORT and temperature by elevating CORT concentrations in eggs and then incubating them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods **LTTER**).
+CORT-treated eggs were topically supplied with 5 µL of crystalline corticosterone (Sigma, Cat.
 No.
-C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), while Control lizards received an equal volume of 100% Ethanol (Control treatment).
-We selected doses based on previous studies where CORT treatment increased mean yolk CORT levels ~3.7x higher than control eggs [**ONDI'S PAPER**].
-Then, eggs were incubated in one of the two previously mentioned temperature regimes.
-These temperatures are within the natural limits in _L.
+C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), while Control eggs received an equal volume of 100% Ethanol.
+CORT dose increased mean yolk CORT levels ~3.7x higher than control eggs in previous studies [@crino2024eggs].
+Eggs were then incubated in one of the two previously mentioned temperature regimes that are within the natural limits in _L.
 delicata_ [@cheetham2011embryonic].
-The number of eggs per clutch assigned to each hormone and temperature treatment was counterbalanced in a partial split-clutch design.
-At least one egg per clutch was assigned to each treatment, while in cluthces bigger than four the remaining eggs were randomly assigned to one of the treatments.
+
+We counterbalanced the number of eggs per clutch assigned to each hormone and temperature using a partial split-clutch design where at least one egg per clutch was assigned to each treatment.
+For clutches bigger than four, we assigned the remaining eggs randomly to one of the treatments.
 When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
@@
 #### Learning  
-The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact.
-The maze was constructed from white polylactic acid (PLA) and featured six arms extending from a hexagonal central starting point (see @fig).
-
-In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes  (see @fig).
-During acclimatation, the central are was surrounded by a yellow 3D-printed PLA device mounted on a pulley system.
+The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact (see @fig-Methods **LTTER**).
+In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes (see @fig-Methods **LTTER**).
+During acclimatation, the central are was surrounded by a yellow device mounted on a pulley system.
 At the start of each trial, this device was lifted to startle the lizard.
@@
 
-Extra-maze cues (but not intra-maze cues) were available to assist with navigation.
+Extra-maze cues (but not intra-maze cues) were available to assist with navigation (see @fig-Methods **LTTER**).
 This design aimed to assess the lizards' ability to learn spatial relationships based on external cues rather than relying on familiar objects within the maze.
-To avoid confounding effects from chemical cues, the maze was cleaned with 70% ethanol between trials.
-To control for visual cues related to the correct arm, the maze was rotated every three trials while maintaining the orientation of the correct arm and the maze's position within the room, which remained consistent for each individual.
-
+Additionally, to prevent lizards from relying on subtle intra-maze cues, the maze was replaced every three trials with one of four identical mazes.
+Each replacement preserved the correct arm’s orientation and the maze’s position within the room, ensuring consistency for each individual.
+The maze was also cleaned with 70% ethanol between trials to avoid the influence of chemical cues.
 To control for potential side biases, the correct arm was randomly assigned to one of the six arms for each trial.
@@
 
-#### Brain extraction
-
 #### Brain mitochondrial activity
-Two months after the completion of the tests, we euthanises lizards usin an injectable anaesthetic followed by decapitation.
-We injected peritoneally 10 mg/kg of a 10 mg/mL alfaxan solution (**REF?**).
-After some minutes, we evaluated the lizards response by turning it upside down, and by testing the pinching relfex in the hand.
-Lizards were decapitated once they were not responsive.
-
-**After decapitation, the head was opened and the brain dissected.
-We focused on two main regions of the brain, the olfactory bulbs and the optic tecta, as they are associated with lizards' chemical and visual perception, respectively (Wynecken, 2007).**
+Immediately following the completion of the tests, we asessed mitochondrial physiology in the brain of all remaining lizards (see below).
+We euthanized lizards via peritoneal injection 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic), followed by decapitation.
+Before decapitation we evaluated the lizard’s response by turning it upside down and by testing the pinching relfex in one of the front limbs.
+This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
+
+After decapitation, we extracted manually the medial cortext in the telencephalon (@fig-Methods **LETTER**).
+We selected the medial cortex because is considered homologous to the mammalian hippocampus, where spatial cognition is encoded [@rodríguez2002conservation; @naumann2015reptilian].
+The tissue was transferred to 1xPBS solutioin and then homogenized mechanically using a 100 µm mesh filter (pluriStrainer).
+The resultant homogenate was divided into two aliquotes: one was used fresh for mitochondrial density, membrane potential, ROS assays; and the other was cryopreserved for later measurements of oxidative damage (DNA damage, lipid peroxidation).
+
+Fresh homogenate suspensions were stained with 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
+We used these fluorescent probes to assess mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
+We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample to differentiate viable cells from debris.
+After adding all the dyes, the samples were incubated at 32 °C for 30 minutes and then immediately transferred to flow cytometry facilities for data collection, within 2 hours.
+
+The aliquotes reserved for oxidative damage assays, were stained with 20 µL of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
+These dyes were used to measure cell viability and lipid peroxidation, respectively.
+The sampled were then fixed in 1% Neutral-Buffered Formalin, washed, and then preserved in a 1X Tris-EDTA solution with 10% DMSO at -20 °C.
+On the day of the oxidative damage assays, the samples were thawed, and DMSO was removed.
+Then, cell membranes were permeabilized by incubating the samples in 200 µL of 1X PBS containing 5mM (6.14 mg/mL) digitonin for 20 minutes at 32°C.
+Following membrane permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody, and allowed them to incubate overnight (~12 hours).
+8-OHdG Polyclonal Antibody was used to assessed 8-OHdG, a marker of oxidative damage on DNA.
+The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32°C for 20 minutes and performed the flow cytometry.
+Oxidative damage assays were performed within 6 months of the initial analysis of fresh samples.
+
+Flow cytometry assays were performed on a Becton Dickson LSRFortessa X-20 flow cytometer using the default wavelength filters.
+The detectors and voltage settings for each assay were determined during pilot trials and and remained consistent throughout the experiment to ensure comparability across different plates and samples.
+Data was processed using FlowJo (v.
+10.1) software.
+We used the geometric mean as our main response variables in analyses.
+For further details on the homogenization, staining, or flow cytometry assays, see the Supplementary Material.
 
@@
 
-We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
-**As such, we run a total of different 8 models employing Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022).
-Each model consisted of four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations.'Choice', i.e.
-whether the individual chose correct (1) or not (0) was used as the response variable.
-The fixed effects of the model included a triple interaction between: trial ('Associative trial' or 'Reversal trial') as a numeric variable, and hormone treatmet ('CORT' versus 'Control') and the temperature at which eggs were incubated ('Cold' versus 'Hot') as factors.
-For the random effects, we employed lizard identity as a random intercept, and as a random slope we included the variable trial ('Associative' or 'Reversal') within each level of lizard identity.
-We used the resulting posterior of these models to evaluate learning differences between treatments within and between species and colour assigned.
-More specifically, we calculated learning slopes by using the estimates of the trial variable per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not.
-We used the pmcmc method to test whether those slopes or the comparissons between 'Treatments' (e.g.
-slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) were different from zero (two-tailed tests).
-We considered statistical significance if p-value < 0.05.**
-
-There are a total of  `r total_individuals`, and `r total_clutches` clutches in the database.
-The average number of individuals per clutch is `r mean(individuals_per_clutch$individual_count)`, with a minimum of `r min(individuals_per_clutch$individual_count)` individuals per clutch (`r sum(individuals_per_clutch$individual_count == min(individuals_per_clutch$individual_count))` clutches) and a maximum of `r max(individuals_per_clutch$individual_count)` (`r sum(individuals_per_clutch$individual_count == max(individuals_per_clutch$individual_count))` clutches).
-Clutches with only one individual do not contribute to the estimation of within-clutch variability.
-Including these clutches might lead to misleading variance estimates and convergence issues in the hierarchical model.
-Therefore, we initially fit the model without the clutch effect to obtain stable estimates at the individual level.
+We performed all the analyses using brm models from stan [@stan] on R version 4.4.0 [@R].
+We run a series of univariate and miltivariate models to test the effects of early envirnoment on each variable separately plus the relationships between physiology and learning.
+All models consisted of four MCMC chains of 8000 iterations, with a warm up interval of 2000 iterations.
+
+Univariate models were used to test the effects of early environment on each of the variables recorded: mitochondrial density, mitochondrial potential, ROS, DNA damage, lipid peroxidation, and the number of errors as a measure of learning.
+We included in all the models the hormone treatment (CORT vs Control), incubation temperature (Cold vs Hot) and their interaction.
+Sex and age were included in preliminary models, and excluded from the final models when they were not significant (see Supplementary Material).
+For all univariate models, clutch identity was included as a random factor.
+Mitochondrial related variables (density, potential, ROS, DNA damage, and lipid peroxidation) where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [**REF DAN**].
+These variables were considered to follow a normal distribution.
+The number of errors was modeled as a function of trial, cort, temperature, and the triple interaction.
+We included in this model the trial within each level of lizard identity as a random slope.
+The error structure was modelled using a negative binomial distribution with a logit link function [negbinomial(link = "log")].
+Otherwise, the procedure was as in the other models.
+
+We used the posterior distributions of parameters from these models to test for differences between treatments.
+Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments.
+Slope estimates >0 were considered as evidence of learning.
+pMCMC were used to test the hypothesis that slopes and slope contrasts are different from zero.
+We considered an effect statistically significant when pMCMC <0.05.
+
+To explore links between mitochondrial function and detection latency, we fitted a multivariate model on brm.
+We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-SEM).
+Individual measurements for learning performance were obtained by modelling errors as a function of trial.
+We included the trial within each level of lizard identity as a random slope, which was used to obtain the learning slopes for each individual.
+These learning slopes, as well as the rest of variables, were standardized as before (see above).
+We assumed a gaussian family for all the variables.
+Factors found as non-significant in the univariate models were excluded.
+Clutch identity remained a random factor.
+Direct, indirect, and total effects were derived from posterior estimates following a Structural Equations Modelling approach [@kline2005principles] (see **TABLE** in Supplementary Material).
 
@@
 
-Originally, we started with 96 lizards, 48 per species and 12 per treatment per species.
+**Originally, we started with 96 lizards, 48 per species and 12 per treatment per species.
 However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards.
-Final sample sizes per treatment and species are disclosed on @tbl-data.
+Final sample sizes per treatment and species are disclosed on @tbl-data.**
 
@@
 
-```{r, learning_results_names}
-#| label: learning_results_names
+```{r, organise_posteriors_mit}
+#| label: organise_posteriors_mit
+#
+# Organising the posteriors of the previous models for tables and figures
+#
+source(here("R", "func.R"))
+#
+#
+post_mit <- data.frame()
+# Names posteriors:
+names <- c("m_def_mit_density", "m_def_mit_potential", "m_def_ROS", "m_def_DNAdamage", "m_def_peroxidation")
+#
+# Organising the results
+for (pos in names) {
+  model <- get(pos)      # Get the model from the global environment
+  post_result <- tidy_post(model)        # Apply tidy_post to each model
+  # Add a new column to the rmodel
+  post_result$Model <- pos
+  post_mit <- bind_rows(post_mit, post_result)
+}
+#
+#
+#
+# Extracting the posteriors for the models with mitochondrial variables and the values of interest.
+Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
+#
+MD <- post_values(m_def_mit_density, "none")
+MP <- post_values(m_def_mit_potential, "none")
+ROS <- post_values(m_def_ROS, "none")
+DNA <- post_values(m_def_DNAdamage, "none")
+LP <- post_values(m_def_peroxidation, "none")
+```
+
+```{r, fig-results_mit}
+#| label: fig-results_mit
+#| fig-cap: "Estimates of mitochondial density (A), mitochondrial potential (B), ROS (C), DNA damage (D), and lipid peroxidation (E) in the medial cortex of L.
+delicata hatchlings as a function of the different prenatal conditions.
+Black dots indicate the posterior mean, and the bars represent the SD of the estimates.
+The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions.
+Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
+#| fig-name: "fig-results_oxidative"
+#
+source(here("R", "func.R"))
+#
+# A) Plotting the results for all variables
+plot_mit_density <- plotting(MD, "Mit density")
+plot_mit_potential <- plotting(MP, "Mit potential")
+plot_ros <- plotting(ROS, "ROS")
+#
+plot_legend_top <- plotting(ROS, "ROS") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_top))
+legend_mit_top <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+plot_dnadamage <- plotting(DNA, "DNA damage")
+plot_peroxidation <- plotting(LP, "Lipid peroxidation")
+#
+plot_legend_bottom <- plotting(DNA, "DNA damage") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
+legend_mit_bottom <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+# B) Organising the plots
+fig_mit_top <- plot_grid(plot_mit_density, plot_mit_potential,
+                        nrow = 1, rel_widths = c(1, 1))
+fig_mit_mid <- plot_grid(plot_ros, NULL,
+                        nrow = 1, rel_widths = c(1, 1))
+fig_mit_bottom <- plot_grid(plot_dnadamage, plot_peroxidation,
+                        nrow = 1, rel_widths = c(1, 1))
+#
+# C) Merging everything in final figure
+# Create the figure grid with extra space for legend
+fig_mit <- plot_grid(
+  fig_mit_top, fig_mit_mid, NULL, fig_mit_bottom, NULL,
+  nrow = 5, rel_heights = c(0.3, 0.3, 0.05, 0.3, 0.05))
+# Final composition: Merge everything, adding images and legend
+final_plot_mit <- ggdraw(fig_mit) +
+  # Insert legend in the middle for ROS and energy variables
+  draw_grob(legend_mit_top, x = 0.49, y = 0.39, width = 0.01, height = 0.01) +
+  # Insert legend in the bottom for DNA damage and lipid peroxidation
+  draw_grob(legend_mit_bottom, x = 0.49, y = 0.03, width = 0.01, height = 0.01) +
+  # Insert title for each plot
+  annotate("text", x = 0.0325, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.531, y = 0.985, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0325, y = 0.682, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0325, y = 0.335, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.53, y = 0.335, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  # Insert figure brain
+  draw_image(here("./Others/brain_fig.png"), x = 0.5, y = 0.3, width = 0.5, height = 0.5)
+#
+# Print final plot
+ggsave(here("./output/figures/text/results_mit.png"), plot = final_plot_mit, width = 21, height = 21, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/results_mit.png")
+```
+
+```{r, organise_posteriors_learning}
+#| label: organise_posteriors_learning
 # Rename some of the posteriors and make new estimates for the learning rate 
@@
   group_by(treatment) %>%
-  summarize(se_int = sd(intercepts)/sqrt(length(intercepts)),
+  summarize(int_sd = sd(intercepts),
             intercepts = mean(intercepts))
@@
   group_by(treatment) %>%
-  summarize(se_slope = sd(slopes)/sqrt(length(slopes)),
+  summarize(slope_sd = sd(slopes),
             slopes = mean(slopes),)
-data_fig_learningB <- merge(data_fig_learning_slopesB, data_fig_learning_intercepts, by = "treatment")
-treatment <- unique(data_fig_learningB$treatment)
+data_fig_learningA <- merge(data_fig_learning_slopesA, data_fig_learning_intercepts, by = "treatment")
+treatment <- unique(data_fig_learningA$treatment)
 fig_A_df <- data.frame()
 for(t in treatment){ # Loop per treatment
-  df <- data_fig_learningB %>%
+  df <- data_fig_learningA %>%
     filter(treatment == t) %>%
@@
     if (x == 0){
-      standard <- df$se_int
+      sd <- df$int_sd
     } else {
-      standard <- df$se_slope
+      q25 <- df$slope_sd
     }
@@
                           errors = value,
-                          se = standard)
+                          sd = sd * value)
     fig_A_df <- rbind(fig_A_df, temp_df)
@@
 #### C) Plot the fig-learning A and fig-learning B separately 
-fig_learningA <- plot_errors(fig_A_df)
+fig_learningA <- plot_errorsday(fig_A_df)
 fig_learningB <- plot_slopes(data_fig_learning_slopes)
 #
-#
-
-errors_df <- data_spal %>%
-  mutate(Treatment = paste(cort, temp, sep = "-")) %>%
-    mutate(Treatment = factor(Treatment,
-      levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
-    )) %>%
-  group_by(day, Treatment) %>%
-  summarise(
-    mean_errors = mean(errors, na.rm = TRUE),
-    sd_errors = sd(errors, na.rm = TRUE)
-  ) %>%
-  ungroup() %>%
-  data.frame()
-fig_errors <- ggplot(learn_df, aes(x = Trial, y = Value, color = Treatment, fill = Treatment)) +
-  geom_smooth(se = TRUE, linewidth = 1, alpha = 0.075) +
-  scale_color_manual(values = c("CORT-Cold" = "#00008B", "Control-Cold" = "#68bde1", "CORT-Hot" = "#b50101", "Control-Hot" = "#fa927d")) +
-  scale_fill_manual(values = c("CORT-Cold" = "darkblue", "Control-Cold" = "#68bde1", "CORT-Hot" = "#b50101", "Control-Hot" = "#fa927d")) +
-  theme_classic() +
-  labs(y = "Mean amount of errors made", x = "Trial") +
-  theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) +
-  theme(
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
-    legend.position = "none"
-  )
-## B.2) Make the the plot for the mean slope per treatment using the probability of choosing the correct feeder
-fig_slope_errors <- plot_slopes("errors")
-#
-#
-## B.3) Combine plots of probability and slopes to have the "choice" figure
-fig_results_errors <- plot_grid(fig_errors, fig_slope_errors, nrow = 1) 
-ggsave("./output/figures/fig_results_errors.png", plot=fig_results_errors, width = 25, height = 15, units = "cm", dpi = 600)
+#### D) Combine plots A and B to have fig-learning
+fig_results_learning <- plot_grid(fig_learningA, fig_learningB, nrow = 1) +
+  # Insert title for each plot
+  annotate("text", x = 0.045, y = 0.935, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.545, y = 0.935, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") 
+#
+ggsave("./output/figures/text/fig_results_errors.png", plot=fig_results_learning , width = 21, height = 7, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/text/fig_results_errors.png")
+```
+
+#### SEM
+
+```{r, model_sem}
+#| label: model_sem
+# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
+# Learning here was evaluated obtaining the learning slopes for each individual (see data_process.R).
+# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
+#
+source(here("R", "func.R"))
+#
+refit <- FALSE
+#
+if(refit){
+  m_SEM <- brm(
+    bf(slope_mean + slope_sd ~ mit_density + mit_potential + DNAdamage + peroxidation + (1|clutch)) +
+    bf(mean_dnadamage ~ ROS + (1|clutch)) +
+    bf(peroxidation ~ ROS + (1|clutch)) +
+    bf(mean_ros ~ mit_density + mit_potential + (1|clutch)) +
+  set_rescor(TRUE),
+  family = gaussian(),
+  data = clean_df, 
+  chains = 4, cores = 4, iter = 8000, warmup = 2000,
+  control = list(adapt_delta = 0.99, max_treedepth = 11))
+  # Save the model
+  saveRDS(m_SEM, file = here("output/models/m_SEM.rds"))
+} else {
+  m_SEM <- readRDS(here("output/models/m_SEM.rds"))
+}
 ```
@@
 )
-models <- c("mean_mitodensity_def_OB",
-            "mean_potential_def_OB",
-            "mean_ros_def_OB",
-            "mean_dnadamage_def_OB",
-            "mean_peroxidation_def_OB",
-            "t_D_def_Chemical",
-            "mean_mitodensity_def_OT",
-            "mean_potential_def_OT",
-            "mean_ros_def_OT",
-            "mean_dnadamage_def_OT",
-            "mean_peroxidation_def_OT",
-            "t_D_def_Visual")
+models <- c("mit_density_def",
+            "mit_potential_def",
+            "ROS_def",
+            "DNAdamage_def",
+            "peroxidation_def",
+            "learning_def")
 #
@@
 bayes_table_df <- data_bayes %>%
-  mutate(Region = gsub(".*_", "", Model)    # Extract everything after the last "_"
-  ) %>%
   mutate(Model = factor(Model,
-                        levels = c("mean_mitodensity_def_OB",
-                                  "mean_potential_def_OB",
-                                  "mean_ros_def_OB",
-                                  "mean_dnadamage_def_OB",
-                                  "mean_peroxidation_def_OB",
-                                  "t_D_def_Chemical",
-                                  "mean_mitodensity_def_OT",
-                                  "mean_potential_def_OT",
-                                  "mean_ros_def_OT",
-                                  "mean_dnadamage_def_OT",
-                                  "mean_peroxidation_def_OT",
-                                  "t_D_def_Visual"),
-                        labels = c("m_def_mean_mitodensity_OB" = "Mit density",
-                                  "m_def_mean_potential_OB" = "Mit potential",
-                                  "m_def_mean_ros_OB" = "ROS",
-                                  "m_def_mean_dnadamage_OB" = "DNA damage",
-                                  "m_def_mean_peroxidation_OB" = "Peroxidation",
-                                  "m_def_t_D_Chemical" = "Detection lat",
-                                  "m_def_mean_mitodensity_OT" = "Mit density",
-                                  "m_def_mean_potential_OT" = "Mit potential",
-                                  "m_def_mean_ros_OT" = "ROS",
-                                  "m_def_mean_dnadamage_OT" = "DNA damage",
-                                  "m_def_mean_peroxidation_OT" = "Peroxidation",
-                                  "m_def_t_D_Visual" = "Detection lat")),
-      Region = factor(Region, levels = c("OB", "Chemical", "OT", "Visual"),
-                      labels = c("OB" = "Olfactory bulbs",
-                                "Chemical" = "Chemical",
-                                "OT" = "Optic tecta",
-                                "Visual" = "Visual"))) %>%
-  dplyr::select(Region, Model, Mean, Error, Q2_5, Q97_5) %>%
-  arrange(Region, Model)
+                        levels = c("mit_density_def",
+                                  "mit_potential_def",
+                                  "ROS_def",
+                                  "DNAdamage_def",
+                                  "peroxidation_def",
+                                  "learning_def"),
+                        labels = c("mit_density_def" = "Mit density",
+                                  "mit_potential_def" = "Mit potential",
+                                  "ROS_def" = "ROS",
+                                  "DNAdamage_def" = "DNA damage",
+                                  "peroxidation_def" = "Lipid peroxidation",
+                                  "learning_def" = "Learning"))) %>%
+  dplyr::select(Model, Mean, Error, Q2_5, Q97_5) %>%
+  arrange(Model)
 #
@@
     Q97_5 = "97.5%") %>%
-  align(align = "center", j = c(3:5), part = "body") %>%
+  align(align = "center", j = c(2:5), part = "body") %>%
   align(align = "center", j = c(1:5), part = "header") %>%
-  flextable::compose(i = c(2:5,8:11), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  flextable::hline(i = 6, part = "body") %>% 
   autofit()
@@
 
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S**14**. Results of the final models.
-```{r, tbl-results_OB_table}
-#| label: results_OB_table
-#| tbl-cap: "Results of the models testing for Olfactory Bulbs."
-#| tbl-name: "results_OB"
-#| tbl-label: "results_OB"
+Mean indicates the mean of the R^2^ values obtained from the posteriors of the model, error the estimated SE of the mean and 95% CI indicates the 95% confidence interval around the mean.
+Model indicates the name of the variable analyzed.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S**14**. Summary of the final models for all mitochondrial variables.
+```{r, tbl-results_mit_table}
+#| label: results_mit_table
+#| tbl-cap: "Summary of the final models for all mitochondrial variables."
+#| tbl-label: "results_mit_table"
+#
 source(here("R", "func.R"))
@@
 # A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB)
-post_OB_refined <- refine_post(post_OB) %>%
+post_mit_refined <- refine_post(post_mit) %>%
   arrange(Variable, Predictors)
@@
 #
-OB_table <- flextable(post_OB_refined) %>%
+mit_table <- flextable(post_mit_refined) %>%
   align(align = "center", j = c(3:5), part = "body") %>%
@@
   bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:12,14:18,20:23,25:27), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::compose(i = c(2:4,6:8,10:12,14:16,18:21), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   autofit()
 #
-OB_table
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r, tbl-contrasts}
+mit_table
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S **!%**. Contrasts between treatments for all the variables analysed.
+```{r tbl-contrasts}
 #| label: tbl-contrasts
-#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group.
-Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean.
-All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero.
-In bold, those values that are significant (p-value <0.05)"
+#| tbl-cap: "Contrasts between prenatal conditions for mitochondrial physiology and learning."
+#
 source(here("R", "func.R"))
 #
-############################## CREATING BIG DF FOR TABLE ##############################
-# Building the vectors for titles of rows and columns
-specie <- c("L. delicata", "L. guichenoti")
-groups <- c("Red", "Blue")
-test <- c("Associative", "Reversal")
-treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
-values <- c("Mean", "95% CI", "p-value")
-# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors.
-p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
-#First get estimates for both tasks
-estimates_asso <- list(
-  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
-  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
-  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
-  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
+# A) Modify posteriors df for learning where we only get/compare the slopes
+LEARN <- data.frame(Control_Cold = slope_ControlCold, 
+                  CORT_Cold = slope_CORTCold,
+                  Control_Hot = slope_ControlHot,
+                  CORT_Hot = slope_CORTHot)
+
+# B) Organise df for mit physiology & learning
+var <- c("MD", "MP", "ROS", "DNA", "LP", "LEARN")
+data_table <- data.frame()
+for(x in var){
+  df <- get(x)
+  Temperature <- format_dec(mean(c(df$CORT_Hot, df$Control_Hot)) - mean(c(df$CORT_Cold, df$Control_Cold)), 3)
+  pMCMC_temp <- format_p(pmcmc(c(df$CORT_Hot, df$Control_Hot) - c(df$CORT_Cold, df$Control_Cold)), 3, equal = FALSE)
+  CORT <- format_dec(mean(c(df$Control_Hot, df$Control_Cold)) - mean(c(df$CORT_Hot, df$CORT_Cold)), 3)
+  pMCMC_cort <- format_p(pmcmc(c(df$Control_Hot, df$Control_Cold) - c(df$CORT_Hot, df$CORT_Cold)), 3, equal = FALSE)
+  Interaction <- format_dec((mean(df$Control_Hot) - mean(df$CORT_Hot)) - (mean(df$Control_Cold) - mean(df$CORT_Cold)), 3)
+  pMCMC_int <- format_p(pmcmc((df$Control_Hot - df$CORT_Hot) - (df$Control_Cold - df$CORT_Cold)), 3, equal = FALSE)
+  data_temp <- data.frame(Variable = x,
+                          Temperature = as.numeric(Temperature),
+                          pMCMC_temp = as.numeric(pMCMC_temp),
+                          CORT = as.numeric(CORT),
+                          pMCMC_cort = as.numeric(pMCMC_cort),
+                          Interaction = as.numeric(Interaction),
+                          pMCMC_int = as.numeric(pMCMC_int))
+  data_table <- dplyr::bind_rows(data_table, data_temp)
+}
+# Modify the df
+data_table_final <- data_table %>%
+  pivot_longer(cols = c(Temperature, CORT, Interaction), 
+               names_to = "Predictor", 
+               values_to = "Contrast") %>%
+  mutate(
+    # Extract the pMCMC values from the corresponding columns
+    `pMCMC contrast` = case_when(
+      Predictor == "Temperature" ~ pMCMC_temp,
+      Predictor == "CORT" ~ pMCMC_cort,
+      Predictor == "Interaction" ~ pMCMC_int
     )
-#
-estimates_rev <- list(
-  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
-  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
-  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
-  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
+  ) %>%
+  mutate(
+    Variable = case_when(
+      Variable == "MD" ~ "Mit density",
+      Variable == "MP" ~ "Mit potential",
+      Variable == "ROS" ~ "ROS",
+      Variable == "DNA" ~ "DNA damage",
+      Variable == "LP" ~ "Lipid peroxidation",
+      Variable == "LEARN" ~ "Learning slopes",
+      TRUE ~ Variable
       )
-# Then get the mean, co.intervals(95%), and p-values
-asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
-asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
-asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
-asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
-asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
-#
-rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
-rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
-rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
-rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
-rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
-#
-# Building the df for the associative
-asso_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(asso_mean, each = 1),
-  CI = rep(asso_intervals, each = 1),
-  PValue = rep(asso_pvalue, each = 1),
-  Task = rep("Associative", length(asso_mean))
-)
-# Building the df for the reversal
-rev_df <- data.frame(
-  Specie = rep(specie, each = length(groups) * length(treatments)),
-  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
-  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
-  Mean = rep(rev_mean, each = 1),
-  CI = rep(rev_intervals, each = 1),
-  PValue = rep(rev_pvalue, each = 1),
-  Task = rep("Reversal", length(rev_mean))
-)
-# Joining both dfs
-table_data <- rbind(asso_df, rev_df)
-table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
-#
-############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
-# Make n_list into a df
-n_df <- as.data.frame(do.call(rbind, n_list)) %>%
-  rename("n" = V1) %>%
-  rownames_to_column("model") %>%
-  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
-  unite("Treatment", c("cort", "temp"), sep = "-") %>%
-  mutate(Specie = factor(Specie,
-                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
-        Treatment = factor(Treatment,
-                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
-# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
-new_table_data <- merge(table_data, n_df) %>%
-  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
-  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
-  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
-  mutate(Specie = factor(Specie,
-                  levels = c("L. delicata", "L. guichenoti")),
-        Group = factor(Group,
-                  levels = c("Red", "Blue")),
-        Treatment = factor(Treatment, 
-                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
-  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
-  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
-  mutate(Treatment = paste0(Treatment, ")"))
-write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
-#
-############################## MAKING THE TABLE ##############################
-## Table format
+    ) %>%
+  dplyr::select(Variable, Predictor, Contrast, `pMCMC contrast`)
+#
+# C) Make the contrasts table:
+#
 set_flextable_defaults(
  font.family = "Times New Roman",
- fint.size = 10)
-# Split the table_data df by task
-real_table <- flextable(new_table_data) %>%
-    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
-    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
-    set_table_properties(width = 1) %>%
-    align(align="center", part="all") %>% 
-    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
-    set_header_labels(Mean_Associative = "Mean",
-                      `95% CI_Associative` = "95% CI",
-                      `p-value_Associative` = "p-value",
-                      Mean_Reversal = "Mean",
-                      `95% CI_Reversal` = "95% CI",
-                      `p-value_Reversal` = "p-value") %>%
-    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
-    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
-    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
-    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
-    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
-    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
+ font.size = 10)
+#
+contrast_table <- flextable(data_table_final) %>%
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>%
+  bold(~`pMCMC contrast` < 0.05, j = c("pMCMC contrast", "Contrast", "Predictor"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC contrast` <0.001, j = c("pMCMC contrast", "Contrast", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
+  flextable::compose(i = c(2,3,5,6,8,9,11,12,14,15,17,18), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
   autofit()
-real_table
+#
+contrast_table
+```
+
+Contrasts were done by:
+_Temperature_: β~Hot~ - β~Cold~
+_CORT_: β~CORT~ - β~Control~
+_Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r, fig-learning_raw}
+#| label: fig-learning_raw
+#| fig.cap: "Raw data of the learning task."
+#
+source(here("R", "func.R"))
+# Modify the df to plot the raw data
+learning_df_plot <- learning_df %>%
+  mutate(treatment = factor(trt, levels = c("B_23", "A_23", "B_28", "A_28"),
+                      labels = c("B_23" = "CORT-Cold (n=20)",
+                                 "A_23" = "Control-Cold (n=20)",
+                                 "B_28" = "CORT-Hot (n=20)",
+                                 "A_28" = "Control-Hot (n=20)"))) %>%
+  group_by(treatment, day) %>%
+  mutate(mean_raw = mean(na.omit(errors)),
+         sd_raw = sd(na.omit(errors))) %>%
+data.frame()
+# Get the plot and combine it with the modeled data
+fig_raw_learning <- plot_errorsday(fig_A_df) +
+  geom_point(data = learning_df_plot, aes(x = day, y = mean_raw, color = treatment), size = 1.5) +
+  scale_color_manual(values = c("CORT-Cold (n=20)" = "#00008B",
+                                "Control-Cold (n=20)" = "#68bde1",
+                                "CORT-Hot (n=20)" = "#b50101",
+                                "Control-Hot (n=20)" = "#fa927d")) +
+  facet_grid(~treatment) +
+  ylim(c(min(na.omit(learning_df_plot$mean_raw)), max(na.omit(learning_df_plot$mean_raw)))) +  # Adjust y-axis limits
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "bottom",
+    legend.title = element_blank()
+  )
+ggsave("./output/figures/suppl/fig_learning_raw.png", plot=fig_raw_learning , width = 21, height = 15, units = "cm", dpi = 600)
 ```

### 2025-03-10 .vdoc.r

@@ -1,1024 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: data_processing
-# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
-refit <- FALSE
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: fig-Methods
-#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
-
-knitr::include_graphics("./Others/SPAL_METH.svg")
-
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
[…]

### 2025-03-10 R/func.r

@@ -8,12 +8,11 @@ pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4,
 #' @title sample
 #' @description Estract sample size per group
 #' @param df To select the df
-#' @param sp To select the species of interest ("deli"/"guich")
 #' @param corti To select cort treatment ("CORT"/"Control")
 #' @param therm To select temp treatment ("Cold"/"Hot")
-sample <- function(df, sp, corti, therm){
+sample <- function(df, corti, therm){
   sample_size <- df %>%
-                filter(species == sp, cort == corti, temp == therm, trial_reversal == 1) %>%
+                filter(cort == corti, temp == therm) %>%
                 group_by(lizard_id) %>%
                 summarise(n = n()) %>%
                 summarise(total_count = sum(n)) %>%
@@ -128,12 +127,184 @@ format_dec <- function(x, n) {
 #' @title format_p
 #' @param x The object
 #' @param n The number of decimals
-format_p <- function(x, n) {
+#' @param equal To write whether we want the equal or not
+format_p <- function(x, n, equal) {
+  if (equal == TRUE){
+    label <- "="
+  } else {
+    label <- ""
+  }
   z <- sprintf(paste0("%.",n,"f"), x)
   tmp <- ifelse(as.numeric(z) <= 0.001, "< 0.001",
          ifelse(as.numeric(z) <= 0.05 & as.numeric(z) > 0.001, "< 0.05",
-                paste0("= ", as.character(z))))
+                paste0(label, as.character(z))))
   return(tmp)
+}
+###################
+###################
+# Function to make the first cleaning for the posteriors and create a big df that includes
+# all variables and their predictors
+#' @title tidy_post
+#' @param df to select the posteriors to use. Options: choice or errors
+tidy_post <- function(df) {
+  df <- as.data.frame(df)
+  summary_df <- df %>%
+    dplyr::select(starts_with("b_")) %>%  # Select only parameter estimates
+    pivot_longer(everything(), names_to = "Predictor", values_to = "Estimate") %>%
+    group_by(Predictor) %>%
+    summarise(
+      Mean = format_dec(mean(Estimate), 3),
+      CI_lower = format_dec(quantile(Estimate, 0.05, na.rm = TRUE), 3),
[…]

### 2025-03-11 ms.qmd
updates

@@
 #
+data_SEM <- clean_df %>%
+  mutate(obs = 1:nrow(.),
+        vec = rep(1, length(obs)))
 if(refit){
   m_SEM <- brm(
-    bf(slope_mean + slope_sd ~ mit_density + mit_potential + DNAdamage + peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ ROS + (1|clutch)) +
-    bf(peroxidation ~ ROS + (1|clutch)) +
-    bf(mean_ros ~ mit_density + mit_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(slope_mean | se(slope_sd, sigma = FALSE) ~ mit_density + mit_potential + DNAdamage + peroxidation + (1|clutch) + (1|q|obs)) +
+    bf(DNAdamage | se(vec, sigma = FALSE) ~ ROS + (1|clutch)+ (1|p|obs)) +
+    bf(peroxidation | se(vec, sigma = FALSE) ~ ROS + (1|clutch) + (1|p|obs)) +
+    bf(ROS | se(vec, sigma = FALSE) ~ mit_density + mit_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
-  data = clean_df, 
+  data = data_SEM, 
   chains = 4, cores = 4, iter = 8000, warmup = 2000,

### 2025-03-11 R/data_process.R

@@ -114,7 +114,7 @@ clean_df <- merge(learning_slopes, final_mito_df, by = "lizard_id") %>%
   dplyr::select(lizard_id, clutch, age, sex, temp, cort,
                 int_mean, int_sd, slope_mean, slope_sd,
                 mit_density, mit_potential, ROS, DNAdamage, peroxidation) %>%
-  mutate(across(where(~ is.double(.) & !is.integer(.)),
-                ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE))))
+  mutate(across(where(~ is.double(.) & !is.integer(.)) & !any_of("slope_sd"),
+                ~ (. - mean(., na.rm = TRUE)) / (2 * sd(., na.rm = TRUE)))) %>%
 data.frame()
 write.csv(clean_df, file = "./output/databases_clean/clean_df.csv")

### 2025-03-11 .vdoc.r
updates

@@ -1,36 +1,36 @@
----
-docx: 
-bibliography: "./bib/refs.bib"
-csl: "./bib/biology-letters.csl"
-reference-doc: "./bib/tmpl.docx"
-execute:
-  echo: false
-  error: false
-  cache: false
-  warning: false
-link-citations: true
-crossref:  
-  fig-title: Fig    # (default is "Figure")
-  title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
-  tbl-prefix: Table  # (default is "Table")
-editor_options: 
-  chunk_output_type: console
----
-
-```{r}
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-```
-
-```{r, data_processing}
+#
+#
+#
 #| label: data_processing
 # The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
 refit <- FALSE
 source(here("R", "data_process.R"))
-```
-
-```{r, data_metrics}
+#
+#
+#
 #| label: data_metrics
 # Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
 data <- read.csv(here("./data/Spatial_learn.csv"))
@@ -50,9 +50,9 @@ individuals_per_clutch <- data %>%
   distinct(lizard_id, clutch) %>%
   group_by(clutch) %>%
   summarize(individual_count = n())
-```
-
-```{r, sampleSize}
+#
+#
+#
 #| label: sampleSize
 # List with the sample sizes from the main database.
 source(here("R", "func.R"))
@@ -78,75 +78,75 @@ for(k in 1:length(hormone)){
     n_damage[[list_name]] <- sample(df = df_damage, corti = hormone[k], therm = temperature[l])
   }
 }
-```
-
-## Introduction
-
[…]

### 2025-03-16 ms.qmd
Updates

--- 2025-03-10 ms.qmd
@@
 ---
+title: "Spatial learning and brain mitochondria are robust to prenatal conditions in the delicate skink"
 docx: 
@@
 
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Daniel W.A. Noble^1^
+
+^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
+^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
+$\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
+
+ORCID:
+
+Pablo Recio ORCID: 0000-0002-5890-0218  
+Dalton C. Leibold ORCID: 0000-0001-9645-2033  
+Ondi L. Crino ORCID: 0000-0001-5700-1387  
+Christopher R. Friesen ORCID: 0000-0001-5338-7454  
+Daniel W.A. Noble ORCID: 0000-0001-9460-8743
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+## Abstract
+Learning is highly sensitive to environmental conditions, including early thermal environments and prenatal exposure to glucocorticoids.
+These factors influence mitochondrial activity, which plays a crucial role in neural development and synaptic plasticity, ultimately shaping cognitive abilities.
+Here, we investigated the interactive effects of prenatal temperature and corticosterone (CORT) — the primary glucocorticoid in reptiles — on spatial learning and brain mitochondrial activity in the delicate skink (_Lampropholis delicata_).
+We supplemented eggs with either CORT or a control vehicle and incubating them under two thermal regimes: cold (23 ± 3°C) or hot (28 ± 3°C).
+Juveniles were then tested in a spatial learning task over 40 days, after which mitochondrial activity in the medial cortex was assessed.
+Our results revealed that both mitochondrial physiology and spatial learning in _L.
+delicata_ are robust to prenatal temperature and CORT exposure.
+In addition, we found no significant relationship between mitochondrial function and cognitive abilities.
+However, mitochondrial potential increased ROS levels, which could impact oxidative stress under high metabolic demands or low antioxidant capacity.
+These findings highlight the cognitive and physiological resilience of _L.
+delicata_ to early-life environmental challenges.
+Future research should explore whether this robustness extends to other brain regions and life stages.
+
 ## Introduction
 
-Cognition encompasses processes such as perception, learning, memory, and decision-making, which allow animals to gather, retain, and use environmental information [@shettleworth].
-These processes are fundamental for survival and reproduction, influencing behaviours like foraging, mate selection, or antipredatory strategies [@dukas_evolutionary_2004].
 Learning - the acquisition and consolidation of new information [@dukas_evolutionary_2004] - enables individuals to create new associations between events, which can be essential for coping with environmental changes [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
@@
 Early developmental environment is particularly influential, as the brain is highly sensitive to environmental inputs during this period [@zhu_prenatal_2004].
-Early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
-Among the physiological mechanisms underlying learning, mitochondrial activity has emerged as a key factor influencing brain function [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+In fact, early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
+Among the physiological mechanisms underlying learning, mitochondrial activity is considered a key factor [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
 The neural processes involved learning impose substantial energetic demands [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], making efficient mitochondrial respiration essential [@du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
 Additionally, learning depends on a dense network of functional neurons [@lefebvre_taxonomic_2011; @amiel_effects_2017], which can be compromised by excessive oxidative stress byproduct of mitochondrial function [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
-Studies on domestic rats (_Rattus norvegicus_) have shown that the effects of early environment on mitochondrial physiology can have profound consequences for cognitive abilities [@zhu_prenatal_2004; @cao2014ampk].
-However, the extent to which these effects are generalizable to other taxa, particularly ectotherms, remains largely unexplored.
+Studies on mammals have shown the pervassive effects of mitochondrial physiology on cognitive abilities [@zhu_prenatal_2004; @hara_presynaptic_2014; @cao2014ampk].
+However, the extent to which these effects are generalizable to other taxa remains largely unexplored.
 
@@
 In particular, maternal stress can strongly influence how mitochondria operate in the offspring [@zhu_prenatal_2004].
-Stressful situations elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryo [@uller_sex-specific_2009], altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) [@zhu_prenatal_2004; @haussmann_embryonic_2012].
+Stressful situations elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryo [@uller_sex-specific_2009] altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].
+For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) [@zhu_prenatal_2004; @haussmann_embryonic_2012; @cao2014ampk].
 
@@
 Here, we examined how prenatal temperature and corticosterone (CORT) — the primary GC in reptiles — affect brain mitochondrial physiology and spatial learning in the delicate skink (_Lampropholis delicata_).
-We hypothesized that prenatal CORT and temperature would influence learning abilities through their effects on mitochondrial activity in the brain.
+We hypothesized that prenatal CORT and temperature would influence learning abilities through impacts on brain mitochondrial activity.
 Specifically, we predicted that prenatal CORT would decrease energy production while increasing oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
@@
 
-#### Animal husbandry  
-*Breeding colony* - Juveniles of _L.
-delicata_ came from a breeding colony established in the laboratory since 2019.
-This colony consisted of 270 adults housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure.
-Enclosures were provided with shelter, nonstick matting, and several small water dishes.
-The lizards were fed approx.
-40 mid-size crickets (_Acheta domestica_) per enclosure three days a week, and water was given daily.
-The crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
-Room temperatures were set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
-
-*Eggs collection and incubation* - Between mid-November 2023 to mid-January 2024, we placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures to provide females with a place to lay the eggs.
-These boxes were checked three days a week.
-After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighted the eggs with a digital scale ± 0.001g error.
-Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
-The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.
-
-*Hatchlings* - Incubators were checked three times a week for hatchlings.
-Lizards were measured and weighed immediately after hatching.
-snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a digital scale with an accuracy of ± 0.001 g.
-Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with nonstick matting and a small water dish.
-Watering, feeding, and temperature conditions were maintained as for adults (see above).
+#### Animals' origin and collection
+Subjects came from a breeding colony established in the laboratory since 2019.
+It consisted in 270 adults living in groups of two males and four females.
+Eggs were collected from these groups between mid-Novemnber 2023 to mid-January 2024.
+After collecting the eggs, we treated them with CORT or vehicle and incubated them under two different temperature regimes (see below).
+Clutch and egg identity were asigned immediately after collection, and the eggs were incubated in individual cups until hatching.
+Hatchlings were also kept in individual enclosures until the end of the experiment.
+For details on husbandry and breeding conditions, see Methods 1: Animal husbandry in Supplementary Material.
 
 #### CORT and Temperature manipulation
-We tested the combined effects of prenatal CORT and temperature by elevating CORT concentrations in eggs and then incubating them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design (@fig-Methods **LTTER**).
+We tested the combined effects of prenatal CORT and temperature by elevating CORT concentrations in eggs and then incubating them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design.
+We used a partial split clutch design where eggs from a given clutch were distributed equally across the four treatments when clutch sizes were larger than four and randomly across treatments when less than four.
 CORT-treated eggs were topically supplied with 5 µL of crystalline corticosterone (Sigma, Cat.
@@
 
-We counterbalanced the number of eggs per clutch assigned to each hormone and temperature using a partial split-clutch design where at least one egg per clutch was assigned to each treatment.
-For clutches bigger than four, we assigned the remaining eggs randomly to one of the treatments.
-When we found less than four eggs in a clutch, we assigned each egg randomly to one of the treatments.
-
 #### Learning  
-The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact (see @fig-Methods **LTTER**).
-In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes (see @fig-Methods **LTTER**).
-During acclimatation, the central are was surrounded by a yellow device mounted on a pulley system.
+The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact.
+In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes.
+During acclimatation, the central area was surrounded by a yellow device mounted on a pulley system.
 At the start of each trial, this device was lifted to startle the lizard.
@@
 
-Extra-maze cues (but not intra-maze cues) were available to assist with navigation (see @fig-Methods **LTTER**).
-This design aimed to assess the lizards' ability to learn spatial relationships based on external cues rather than relying on familiar objects within the maze.
-Additionally, to prevent lizards from relying on subtle intra-maze cues, the maze was replaced every three trials with one of four identical mazes.
+We assessed lizards' spatial learning using external cues for navigation while avoiding intra-maze cues.
+To avoid subtle intra-maze cues, the maze was replaced every three trials with one of four identical mazes.
 Each replacement preserved the correct arm’s orientation and the maze’s position within the room, ensuring consistency for each individual.
-The maze was also cleaned with 70% ethanol between trials to avoid the influence of chemical cues.
+Additionally, the maze was cleaned with 70% ethanol between trials to avoid the influence of chemical cues.
 To control for potential side biases, the correct arm was randomly assigned to one of the six arms for each trial.
@@
 
-The task was repeated once daily for 40 consecutive days, and the number of mistakes made by each lizard was recorded.
-A mistake was defined as the lizard inserting its head into one of the incorrect arms.
-The primary measure of spatial learning was the total number of errors made before the lizard chose the correct arm.
-
-```{r, fig-Methods}
-#| label: fig-Methods
-#| fig.cap: "Panel **A** shows the experimental design of the study.
-On top, the treatments applied to the eggs.
-On the bottom the learning task and the brain region extracted together with the physiological analyses performed.
-In panel **B** the details and measurements of the spatial learning maze."
-
-knitr::include_graphics("./Others/SPAL_METH.svg")
-
-```
+The task was repeated once daily for 40 consecutive days, and the number of errors made by each lizard was recorded.
+An error was defined as the lizard inserting its head into one of the incorrect arms, choosing the same arm two or more times consecutively was counted as independent errors.
 
@@
 We euthanized lizards via peritoneal injection 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic), followed by decapitation.
-Before decapitation we evaluated the lizard’s response by turning it upside down and by testing the pinching relfex in one of the front limbs.
-This protocol was approved by the Animal Ethics Committee of the Australian National University (Protocol number: A2022_33).
-
-After decapitation, we extracted manually the medial cortext in the telencephalon (@fig-Methods **LETTER**).
+Before decapitation we evaluated lizard's righting response and the pinching reflex in one of the front limbs.
+
+After decapitation, we extracted the medial cortext in the telencephalon.
 We selected the medial cortex because is considered homologous to the mammalian hippocampus, where spatial cognition is encoded [@rodríguez2002conservation; @naumann2015reptilian].
-The tissue was transferred to 1xPBS solutioin and then homogenized mechanically using a 100 µm mesh filter (pluriStrainer).
-The resultant homogenate was divided into two aliquotes: one was used fresh for mitochondrial density, membrane potential, ROS assays; and the other was cryopreserved for later measurements of oxidative damage (DNA damage, lipid peroxidation).
+The tissue was transferred to 1xPBS solution and then homogenized mechanically using a 100 µm mesh filter (pluriStrainer).
+The resultant homogenate was divided into two aliquotes: one was used fresh for mitochondrial density, membrane potential, and ROS assays; and the other was cryopreserved for later measurements of DNA damage and lipid peroxidation.
 
@@
 We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample to differentiate viable cells from debris.
-After adding all the dyes, the samples were incubated at 32 °C for 30 minutes and then immediately transferred to flow cytometry facilities for data collection, within 2 hours.
-
-The aliquotes reserved for oxidative damage assays, were stained with 20 µL of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
+These samples were analyzed in a flow cytometer within 2 hours of brain extraction.
+
+The aliquotes reserved for oxidative damage assays were stained with 20 µL of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
 These dyes were used to measure cell viability and lipid peroxidation, respectively.
-The sampled were then fixed in 1% Neutral-Buffered Formalin, washed, and then preserved in a 1X Tris-EDTA solution with 10% DMSO at -20 °C.
+The samples were then fixed in 1% Neutral-Buffered Formalin, washed, and preserved in a 1X Tris-EDTA solution with 10% DMSO at -20 °C.
 On the day of the oxidative damage assays, the samples were thawed, and DMSO was removed.
-Then, cell membranes were permeabilized by incubating the samples in 200 µL of 1X PBS containing 5mM (6.14 mg/mL) digitonin for 20 minutes at 32°C.
-Following membrane permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody, and allowed them to incubate overnight (~12 hours).
-8-OHdG Polyclonal Antibody was used to assessed 8-OHdG, a marker of oxidative damage on DNA.
-The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32°C for 20 minutes and performed the flow cytometry.
+Then, cell membranes were permeabilized in 200 µL of 1X PBS containing 5mM (6.14 mg/mL) digitonin.
+Following membrane permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody - a marker of DNA damage - and allowed them to incubate overnight (~12 hours).
+The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyzed the samples in the the flow cytometer.
 Oxidative damage assays were performed within 6 months of the initial analysis of fresh samples.
 
-Flow cytometry assays were performed on a Becton Dickson LSRFortessa X-20 flow cytometer using the default wavelength filters.
-The detectors and voltage settings for each assay were determined during pilot trials and and remained consistent throughout the experiment to ensure comparability across different plates and samples.
+Flow cytometry was performed on a Becton Dickson LSRFortessa X-20 flow cytometer using the default wavelength filters.
+The detectors and voltage settings for each assay were determined during pilot trials and and remained consistent throughout the experiment.
 Data was processed using FlowJo (v.
 10.1) software.
-We used the geometric mean as our main response variables in analyses.
-For further details on the homogenization, staining, or flow cytometry assays, see the Supplementary Material.
+We obtained the mean fluorescent intensity for mitochondrial density, mitochondrial potential, ROS production, DNA damage, and lipid peroxidation.
+For further details on the homogenization, staining, or flow cytometry assays, see Methods 2: flow cytometry in Supplementary Material.
 
 #### Statistical analyses
-
 We performed all the analyses using brm models from stan [@stan] on R version 4.4.0 [@R].
-We run a series of univariate and miltivariate models to test the effects of early envirnoment on each variable separately plus the relationships between physiology and learning.
+We run a series of univariate and multivariate models to test the effects of early envirnoment on each variable separately plus the relationships between physiology and learning.
 All models consisted of four MCMC chains of 8000 iterations, with a warm up interval of 2000 iterations.
@@
 We included in all the models the hormone treatment (CORT vs Control), incubation temperature (Cold vs Hot) and their interaction.
-Sex and age were included in preliminary models, and excluded from the final models when they were not significant (see Supplementary Material).
+Sex and age were included in preliminary models and excluded from the final models when they were not significant.
 For all univariate models, clutch identity was included as a random factor.
-Mitochondrial related variables (density, potential, ROS, DNA damage, and lipid peroxidation) where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [**REF DAN**].
+Mitochondrial related variables (density, potential, ROS, DNA damage, and lipid peroxidation) where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [@gelman2008scaling].
 These variables were considered to follow a normal distribution.
+
 The number of errors was modeled as a function of trial, cort, temperature, and the triple interaction.
@@
 Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments.
-Slope estimates >0 were considered as evidence of learning.
+Slope estimates <0 were considered as evidence of learning.
 pMCMC were used to test the hypothesis that slopes and slope contrasts are different from zero.
@@
 To explore links between mitochondrial function and detection latency, we fitted a multivariate model on brm.
-We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-SEM).
+We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem).
 Individual measurements for learning performance were obtained by modelling errors as a function of trial.
@@
 Clutch identity remained a random factor.
-Direct, indirect, and total effects were derived from posterior estimates following a Structural Equations Modelling approach [@kline2005principles] (see **TABLE** in Supplementary Material).
+Direct, indirect, and total effects were derived from posterior estimates following a Structural Equations Modelling approach [@kline2005principles] (see Table S4).
+
+One of the plates used to analyze oxidative damage malfunctioned, compromising data quality.
+This data was excluded from the final analysis and, as a result, the sample size for DNA damage and lipid peroxidation is smaller than for the other variables.
+For the multivariate model, we imputed the missing values.
+The final sample sizes are presented below.
 
@@
 
-**Originally, we started with 96 lizards, 48 per species and 12 per treatment per species.
-However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards.
-Final sample sizes per treatment and species are disclosed on @tbl-data.**
+We started with 80 lizards - 20 per treatment - coming from `r total_clutches` clutches.
+However, due to natural mortality (n = 2), only 79 lizards were included in the learning analyses, and 78 in the mitochondrial analyses.
+Besides, due to malfunctioning of the flow cytometer, we ended up with 40 inividuals (10 per treatment) for DNA damage and lipid peroxidation.
+Final sample sizes per treatment and species are disclosed on @fig-learning and @fig-results_mit.
 
@@
 
-```{r, organise_posteriors_mit}
-#| label: organise_posteriors_mit
-#
-# Organising the posteriors of the previous models for tables and figures
-#
-source(here("R", "func.R"))
-#
-#
-post_mit <- data.frame()
-# Names posteriors:
-names <- c("m_def_mit_density", "m_def_mit_potential", "m_def_ROS", "m_def_DNAdamage", "m_def_peroxidation")
-#
-# Organising the results
-for (pos in names) {
-  model <- get(pos)      # Get the model from the global environment
-  post_result <- tidy_post(model)        # Apply tidy_post to each model
-  # Add a new column to the rmodel
-  post_result$Model <- pos
-  post_mit <- bind_rows(post_mit, post_result)
-}
-#
-#
-#
-# Extracting the posteriors for the models with mitochondrial variables and the values of interest.
-Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
-#
-MD <- post_values(m_def_mit_density, "none")
-MP <- post_values(m_def_mit_potential, "none")
-ROS <- post_values(m_def_ROS, "none")
-DNA <- post_values(m_def_DNAdamage, "none")
-LP <- post_values(m_def_peroxidation, "none")
-```
-
-```{r, fig-results_mit}
-#| label: fig-results_mit
-#| fig-cap: "Estimates of mitochondial density (A), mitochondrial potential (B), ROS (C), DNA damage (D), and lipid peroxidation (E) in the medial cortex of L.
-delicata hatchlings as a function of the different prenatal conditions.
-Black dots indicate the posterior mean, and the bars represent the SD of the estimates.
-The y-axis represents the posterior estimates of the variable of interest, and the x-axis represents the different prenatal conditions.
-Lines with asterisks represent significant differences between groups based on pMCMC values (pMCMC < 0.05), no lines indicate no significant differences."
-#| fig-name: "fig-results_oxidative"
-#
-source(here("R", "func.R"))
-#
-# A) Plotting the results for all variables
-plot_mit_density <- plotting(MD, "Mit density")
-plot_mit_potential <- plotting(MP, "Mit potential")
-plot_ros <- plotting(ROS, "ROS")
-#
-plot_legend_top <- plotting(ROS, "ROS") + theme(legend.position = "bottom", legend.title = element_blank())
-gtable <- ggplot_gtable(ggplot_build(plot_legend_top))
-legend_mit_top <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
-#
-plot_dnadamage <- plotting(DNA, "DNA damage")
-plot_peroxidation <- plotting(LP, "Lipid peroxidation")
-#
-plot_legend_bottom <- plotting(DNA, "DNA damage") + theme(legend.position = "bottom", legend.title = element_blank())
-gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
-legend_mit_bottom <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
-#
-# B) Organising the plots
-fig_mit_top <- plot_grid(plot_mit_density, plot_mit_potential,
-                        nrow = 1, rel_widths = c(1, 1))
-fig_mit_mid <- plot_grid(plot_ros, NULL,
-                        nrow = 1, rel_widths = c(1, 1))
-fig_mit_bottom <- plot_grid(plot_dnadamage, plot_peroxidation,
-                        nrow = 1, rel_widths = c(1, 1))
-#
-# C) Merging everything in final figure
-# Create the figure grid with extra space for legend
-fig_mit <- plot_grid(
-  fig_mit_top, fig_mit_mid, NULL, fig_mit_bottom, NULL,
-  nrow = 5, rel_heights = c(0.3, 0.3, 0.05, 0.3, 0.05))
-# Final composition: Merge everything, adding images and legend
-final_plot_mit <- ggdraw(fig_mit) +
-  # Insert legend in the middle for ROS and energy variables
-  draw_grob(legend_mit_top, x = 0.49, y = 0.39, width = 0.01, height = 0.01) +
-  # Insert legend in the bottom for DNA damage and lipid peroxidation
-  draw_grob(legend_mit_bottom, x = 0.49, y = 0.03, width = 0.01, height = 0.01) +
-  # Insert title for each plot
-  annotate("text", x = 0.0325, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.531, y = 0.985, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0325, y = 0.682, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.0325, y = 0.335, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.53, y = 0.335, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  # Insert figure brain
-  draw_image(here("./Others/brain_fig.png"), x = 0.5, y = 0.3, width = 0.5, height = 0.5)
-#
-# Print final plot
-ggsave(here("./output/figures/text/results_mit.png"), plot = final_plot_mit, width = 21, height = 21, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/text/results_mit.png")
-```
-
 ```{r, organise_posteriors_learning}
@@
 #| label: fig-learning
-#| fig.cap: "Learning results"
+#| fig.cap: "Results for learning analyses.
+Panel A shows the predicted number of errors over trials.
+The lines represent the mean predicted number of errors for each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained using the slope and intercept estimates from the posterior distributions.
+Panel B shows the distribution of the estimates of slopes per each treatment.
+The x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
+Points and bars represent the mean and standard deviation of the estimated slopes, respectively.
+Dashed lines indicate value 0.
+The different colors in both panels indicate the different treatments."
 source(here("R", "func.R"))
@@
 
-#### SEM
+We found that the lizards improved their choices over time, making less mistakes as the trials progressed (see @fig-learning and Table S1).
+However, the learning rate was not influenced by CORT, temperature, or their interaction (see contrasts in Table S2).
+
+Mitochondrial density, mitochondrial potential, ROS, DNA damage, and lipid peroxidation was not affected by prenatal conditions (see @fig-results_mit and Table S2).
+We did not see any significant effects of sex and age on mitochondrial physiology either (see Table S3 and Tables S5-S9).
+
+```{r, organise_posteriors_mit}
+#| label: organise_posteriors_mit
+#
+# Organising the posteriors of the previous models for tables and figures
+#
+source(here("R", "func.R"))
+#
+#
+post_mit <- data.frame()
+# Names posteriors:
+names <- c("m_def_mit_density", "m_def_mit_potential", "m_def_ROS", "m_def_DNAdamage", "m_def_peroxidation")
+#
+# Organising the results
+for (pos in names) {
+  model <- get(pos)      # Get the model from the global environment
+  post_result <- tidy_post(model)        # Apply tidy_post to each model
+  # Add a new column to the rmodel
+  post_result$Model <- pos
+  post_mit <- bind_rows(post_mit, post_result)
+}
+#
+#
+#
+# Extracting the posteriors for the models with mitochondrial variables and the values of interest.
+Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
+#
+MD <- post_values(m_def_mit_density, "none")
+MP <- post_values(m_def_mit_potential, "none")
+ROS <- post_values(m_def_ROS, "none")
+DNA <- post_values(m_def_DNAdamage, "none")
+LP <- post_values(m_def_peroxidation, "none")
+```
+
+```{r, fig-results_mit}
+#| label: fig-results_mit
+#| fig-cap: "Estimates of mitochondial density (A), mitochondrial potential (B), ROS (C), DNA damage (D), and lipid peroxidation (E) in the medial cortex as a function of the different prenatal conditions.
+The x-axis represents the estimated values and in the y-axis is the density of the estimates.
+Points and bars represent the mean and standard deviation of the estimated values, respectively.
+The different colors in both panels indicate the different treatments."
+#| fig-name: "fig-results_oxidative"
+#
+source(here("R", "func.R"))
+#
+# A) Plotting the results for all variables
+plot_mit_density <- plotting(MD, "Mit density")
+plot_mit_potential <- plotting(MP, "Mit potential")
+plot_ros <- plotting(ROS, "ROS")
+#
+plot_legend_top <- plotting(ROS, "ROS") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_top))
+legend_mit_top <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+plot_dnadamage <- plotting(DNA, "DNA damage")
+plot_peroxidation <- plotting(LP, "Lipid peroxidation")
+#
+plot_legend_bottom <- plotting(DNA, "DNA damage") + theme(legend.position = "bottom", legend.title = element_blank())
+gtable <- ggplot_gtable(ggplot_build(plot_legend_bottom))
+legend_mit_bottom <- gtable$grobs[[which(sapply(gtable$grobs, function(x) x$name) == "guide-box")]]
+#
+# B) Organising the plots
+fig_mit_top <- plot_grid(plot_mit_density, plot_mit_potential,
+                        nrow = 1, rel_widths = c(1, 1))
+fig_mit_mid <- plot_grid(plot_ros, NULL,
+                        nrow = 1, rel_widths = c(1, 1))
+fig_mit_bottom <- plot_grid(plot_dnadamage, plot_peroxidation,
+                        nrow = 1, rel_widths = c(1, 1))
+#
+# C) Merging everything in final figure
+# Create the figure grid with extra space for legend
+fig_mit <- plot_grid(
+  fig_mit_top, fig_mit_mid, NULL, fig_mit_bottom, NULL,
+  nrow = 5, rel_heights = c(0.3, 0.3, 0.05, 0.3, 0.05))
+# Final composition: Merge everything, adding images and legend
+final_plot_mit <- ggdraw(fig_mit) +
+  # Insert legend in the middle for ROS and energy variables
+  draw_grob(legend_mit_top, x = 0.49, y = 0.39, width = 0.01, height = 0.01) +
+  # Insert legend in the bottom for DNA damage and lipid peroxidation
+  draw_grob(legend_mit_bottom, x = 0.49, y = 0.03, width = 0.01, height = 0.01) +
+  # Insert title for each plot
+  annotate("text", x = 0.0325, y = 0.985, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.531, y = 0.985, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0325, y = 0.682, label = "C", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.0325, y = 0.335, label = "D", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.53, y = 0.335, label = "E", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  # Insert figure brain
+  draw_image(here("./Others/brain_fig.png"), x = 0.49, y = 0.3, width = 0.5, height = 0.5)
+#
+# Print final plot
+ggsave(here("./output/figures/text/results_mit.png"), plot = final_plot_mit, width = 21, height = 21, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/results_mit.png")
+```
 
@@
 #
+data_SEM <- clean_df %>%
+  mutate(obs = 1:nrow(.),
+        vec = rep(1, length(obs)))
 if(refit){
   m_SEM <- brm(
-    bf(slope_mean + slope_sd ~ mit_density + mit_potential + DNAdamage + peroxidation + (1|clutch)) +
-    bf(mean_dnadamage ~ ROS + (1|clutch)) +
-    bf(peroxidation ~ ROS + (1|clutch)) +
-    bf(mean_ros ~ mit_density + mit_potential + (1|clutch)) +
-  set_rescor(TRUE),
+    bf(slope_mean | se(slope_sd, sigma = FALSE) ~ mit_density + mit_potential + mi(DNAdamage) + mi(peroxidation) + (1|clutch) + (1|q|obs)) +
+    bf(DNAdamage | mi() + se(vec, sigma = FALSE) ~ ROS + (1|clutch)+ (1|p|obs)) +
+    bf(peroxidation | mi() + se(vec, sigma = FALSE) ~ ROS + (1|clutch) + (1|p|obs)) +
+    bf(ROS | se(vec, sigma = FALSE) ~ mit_density + mit_potential + (1|clutch) + (1|t|obs)) +
+  set_rescor(FALSE),
   family = gaussian(),
-  data = clean_df, 
+  data = data_SEM, 
   chains = 4, cores = 4, iter = 8000, warmup = 2000,
-  control = list(adapt_delta = 0.99, max_treedepth = 11))
+  control = list(adapt_delta = 0.99, max_treedepth = 15))
   # Save the model
@@
 
+```{r, sem_tidy}
+#| label: sem_tidy
+source(here("R", "func.R"))
+#
+# I am extracting here all the values for getting the total effects of each of the variables in the model.
+I am using the posterior values for each of the variables to get the total effects assuming that:
+## total effect = direct effect + indirect effect + residual correlation
+# In other words:
+## total effect = 
+# Extract the posteriors for the SEM model
+post_sem <- as_draws_df(m_SEM) 
+#
+#### A) Get the direct paths / regression coefficients per each variable
+# 
+# A.1) Learning
+coeff_mitodensity_learn <- post_sem$b_slopemean_mit_density
+coeff_potential_learn <- post_sem$b_slopemean_mit_potential
+coeff_dna_learn<- post_sem$bsp_slopemean_miDNAdamage
+coeff_perox_learn <- post_sem$bsp_slopemean_miperoxidation
+# 
+# A.2) DNA damage
+coeff_ros_dna <- post_sem$b_DNAdamage_ROS
+#
+# A.3) Lipid peroxidation
+coeff_ros_perox <- post_sem$b_peroxidation_ROS
+#
+# A.4) ROS
+coeff_mitodensity_ros <- post_sem$b_ROS_mit_density
+coeff_potential_ros <- post_sem$b_ROS_mit_potential
+#
+#
+#### B) Get the indirect paths for each variable
+#
+# B.1) Detection
+undir_ros_learn <- coeff_ros_dna * coeff_dna_learn + coeff_ros_perox * coeff_perox_learn
+undir_mitodensity_learn <- coeff_mitodensity_ros * coeff_ros_dna * coeff_dna_learn + coeff_mitodensity_ros * coeff_ros_perox * coeff_perox_learn
+undir_potential_learn <- coeff_potential_ros * coeff_ros_dna * coeff_dna_learn + coeff_potential_ros * coeff_ros_perox * coeff_perox_learn
+#
+# B.2) DNA damage
+undir_mitodensity_dna <- coeff_mitodensity_ros * coeff_ros_dna
+undir_potential_dna <- coeff_potential_ros * coeff_ros_dna
+#
+# B.3) Lipid peroxidation
+undir_mitodensity_perox <- coeff_mitodensity_ros * coeff_ros_perox
+undir_potential_perox <- coeff_potential_ros * coeff_ros_perox
+#
+#
+#### C) Get the total effects for each variable
+#
+# C.1) Detection
+total_mitodensity_learn <- coeff_mitodensity_learn + undir_mitodensity_learn
+total_potential_learn <- coeff_potential_learn + undir_potential_learn
+total_ros_learn <- undir_ros_learn
+total_dna_learn <- coeff_dna_learn
+total_perox_learn <- coeff_perox_learn
+#
+# C.2) DNA damage
+total_ros_dna <- coeff_ros_dna
+total_mitodensity_dna <- undir_mitodensity_dna
+total_potential_dna <- undir_potential_dna
+#
+# C.3) Lipid peroxidation
+total_ros_perox <- coeff_ros_perox
+total_mitodensity_perox <- undir_mitodensity_perox
+total_potential_perox <- undir_potential_perox
+#
+# C.4) ROS
+total_mitodensity_ros <- coeff_mitodensity_ros
+total_potential_ros <- coeff_potential_ros
+#
+#
+# D) Create a df with the values for each variable
+#
+# D.1) Detection (for example)
+learn_sem <- data.frame(
+  reference_variable = rep("learn", 5),
+  predictor_modulator = c("mit_density",
+                          "mit_potential",
+                          "ROS",
+                          "DNAdamage",
+                          "peroxidation"),
+  direct_effects = I(list(coeff_mitodensity_learn,
+                          coeff_potential_learn,
+                          NA,
+                          coeff_dna_learn,
+                          coeff_perox_learn)),
+  indirect_effects = I(list(undir_mitodensity_learn,
+                          undir_potential_learn,
+                          undir_ros_learn,
+                          NA,
+                          NA)),
+  total_effects = I(list(total_mitodensity_learn,
+                        total_potential_learn,
+                        total_ros_learn,
+                        total_dna_learn,
+                        total_perox_learn))
+  ) %>% 
+  group_by(predictor_modulator) %>%
+  summarize(
+    # Summarizing each list column (direct_effects, indirect_effects, etc.)
+    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
+    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
+    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
+    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
+    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
+    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
+    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
+    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
+    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
+  mutate(source = "learn_sem")
+#
+# D.2) DNA damage
+DNA_sem <- data.frame(
+  reference_variable = rep("DNA damage", 3),
+  predictor_modulator = c("mit_density",
+                          "mit_potential",
+                          "ROS"),
+  direct_effects = I(list(NA,
+                          NA,
+                          coeff_ros_dna)),
+  indirect_effects = I(list(undir_mitodensity_dna,
+                          undir_potential_dna,
+                          NA)),
+  total_effects = I(list(total_mitodensity_dna,
+                        total_potential_dna,
+                        total_ros_dna))) %>% 
+  group_by(predictor_modulator) %>%
+  summarize(
+    # Summarizing each list column (direct_effects, indirect_effects, etc.)
+    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
+    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
+    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
+    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
+    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
+    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
+    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
+    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
+    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
+  mutate(source = "DNA_sem")
+#
+# D.3) Lipid peroxidation
+perox_sem <- data.frame(
+  reference_variable = rep("lipid peroxidation", 3),
+  predictor_modulator = c("mit_density",
+                          "mit_potential",
+                          "ROS"),
+  direct_effects = I(list(NA,
+                          NA,
+                          coeff_ros_perox)),
+  indirect_effects = I(list(undir_mitodensity_perox,
+                          undir_potential_perox,
+                          NA)),
+  total_effects = I(list(total_mitodensity_perox,
+                        total_potential_perox,
+                        total_ros_perox))) %>% 
+  group_by(predictor_modulator) %>%
+  summarize(
+    # Summarizing each list column (direct_effects, indirect_effects, etc.)
+    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
+    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
+    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
+    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
+    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
+    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
+    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
+    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
+    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
+  mutate(source = "perox_sem")
+#
+# D.4) ROS
+ROS_sem <- data.frame(
+  reference_variable = rep("ROS", 2),
+  predictor_modulator = c("mit_density",
+                          "mit_potential"),
+  direct_effects = I(list(coeff_mitodensity_ros,
+                          coeff_potential_ros)),
+  indirect_effects = I(list(NA,
+                          NA)),
+  total_effects = I(list(total_mitodensity_ros,
+                        total_potential_ros))) %>% 
+  group_by(predictor_modulator) %>%
+  summarize(
+    # Summarizing each list column (direct_effects, indirect_effects, etc.)
+    mean_direct_effects = format_dec(mean(unlist(direct_effects), na.rm = TRUE), 3),
+    mean_indirect_effects = format_dec(mean(unlist(indirect_effects), na.rm = TRUE), 3),
+    mean_total_effects = format_dec(mean(unlist(total_effects), na.rm = TRUE), 3),
+    q5_direct = format_dec(quantile(unlist(direct_effects), 0.05, na.rm = TRUE), 3),
+    q95_direct = format_dec(quantile(unlist(direct_effects), 0.95, na.rm = TRUE), 3),
+    q5_indirect = format_dec(quantile(unlist(indirect_effects), 0.05, na.rm = TRUE), 3),
+    q95_indirect = format_dec(quantile(unlist(indirect_effects), 0.95, na.rm = TRUE), 3),
+    q5_total = format_dec(quantile(unlist(total_effects), 0.05, na.rm = TRUE), 3),
+    q95_total = format_dec(quantile(unlist(total_effects), 0.95, na.rm = TRUE), 3)) %>%
+  mutate(source = "ROS_sem")
+#
+#
+# E) Merge everything into a single df
+#
+sem_results <- bind_rows(learn_sem, DNA_sem, perox_sem, ROS_sem)
+```
+
+The multivariate model showed that mitochondrial physiology was not related to learning abilities (see @fig-sem and Table S4).
+We did not find significant relationships between ROS and oxidative damage or mitochondrial density and ROS (see @fig-sem and Table S4).
+However, we found that ROS production increased with mitochondrial potential (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], pMCMC `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`).
+
+```{r, fig-sem}
+#| label: fig-sem
+#| fig-cap: "Schematic representation of the multivariate model and the predicted coefficient for each variable.
+Arrows indicate the directionality of the estimate and the values show the mean predicted coefficient and the 95% CI around the mean."
+#| fig-name: "fig-sem"
+# 
+# A) Getting all the direct coefficients
+density_learn <- paste0(format_dec(mean(coeff_mitodensity_learn), 3),
+                  " [", format_dec(quantile(coeff_mitodensity_learn, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_mitodensity_learn, 0.95), 3), "]")
+potential_learn <- paste0(format_dec(mean(coeff_potential_learn), 3),
+                  " [", format_dec(quantile(coeff_potential_learn, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_potential_learn, 0.95), 3), "]")
+dna_learn <- paste0(format_dec(mean(coeff_dna_learn), 3),
+                  " [", format_dec(quantile(coeff_dna_learn, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_dna_learn, 0.95), 3), "]")
+perox_learn <- paste0(format_dec(mean(coeff_perox_learn), 3),
+                  " [", format_dec(quantile(coeff_perox_learn, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_perox_learn, 0.95), 3), "]")
+#
+ros_dna <- paste0(format_dec(mean(coeff_ros_dna), 3),
+                  " [", format_dec(quantile(coeff_ros_dna, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_ros_dna, 0.95), 3), "]")
+#
+ros_perox <- paste0(format_dec(mean(coeff_ros_perox), 3),
+                  " [", format_dec(quantile(coeff_ros_perox, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_ros_perox, 0.95), 3), "]")
+#
+density_ros <- paste0(format_dec(mean(coeff_mitodensity_ros), 3),
+                  " [", format_dec(quantile(coeff_mitodensity_ros, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_mitodensity_ros, 0.95), 3), "]")
+potential_ros <- paste0(format_dec(mean(coeff_potential_ros), 3),
+                  " [", format_dec(quantile(coeff_potential_ros, 0.05), 3),
+                  ", ", format_dec(quantile(coeff_potential_ros, 0.95), 3), "]")
+#
+imgSEM <- readPNG(here("Others", "SEM.png"))
+plot_SEM <- rasterGrob(imgSEM, interpolate = TRUE)
+#
+fig_SEM <- ggdraw(plot_SEM) +
+  annotate("text", x = 0.433, y = 0.19, label = density_learn, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.433, y = 0.85, label = potential_learn, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.877, y = 0.675, label = dna_learn, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.877, y = 0.387, label = perox_learn, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.547, y = 0.745, label = ros_dna, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.547, y = 0.295, label = ros_perox, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.325, y = 0.284, label = density_ros, hjust = 1, vjust = 1, size = 3.5, family = "Times") +
+  annotate("text", x = 0.324, y = 0.757, label = potential_ros, hjust = 1, vjust = 1, size = 3.5, family = "Times")
+ggsave(here("./output/figures/text/SEM.png"), plot = fig_SEM, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/text/SEM.png")
+```
+
 ## Discussion
+
+We found that lizards improved their performance over time, indicating learning.
+However, this improvement was not influenced by prenatal conditions.
+Similarly, mitochondrial physiology was not affected by prenatal conditions, and we found no significant relationships between mitochondrial physiology and learning abilities.
+However, ROS production increased with mitochondrial potential, suggesting that mitochondrial potential may be a good predictor of ROS production in medial cortex of _L.
+delicata_ with potential implications for the balance between energy production and oxidative damage.
+
+#### Spatial learning in _Lampropholis delicata_ is robust to prenatal conditions  
+Lizards exhibited learning, but learning rates were not influenced by prenatal CORT or temperature.
+This contrasts with findings in other taxa, where incubation temperature and prenatal GC elevations affected learning and brain development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017].
+Our results suggest that _L.
+delicata_ buffers cognitive functions against early environmental inputs, consistent with previous findings showing that behavioural flexibility in this species is also unaffected by prenatal CORT or temperature [@recio2025early].
+However, since prenatal CORT and temperature can have region-specific effects on brain development [@coomber_independent_1997; @sakata_neural_2000], other cognitive domains might be more sensitive to prenatal conditions.
+Future research should explore whether different brain regions in _L.
+delicata_ vary in their resilience to prenatal inputs to gain a more comprehensive understanding of cognitive development in this species.
+
+#### Mitochondrial physiology is not affected by prenatal conditions  
+Contrary to our predictions, prenatal temperature and CORT did not significantly influence mitochondrial physiology in the medial cortex of _L.
+delicata_.
+Studies in other taxa have been shown that incubation at high temperatures could decrease energy production and oxidative stress [@treidel2016temperature; @zavorka_climate_2021] while, GCs elevation or maternal stress is related to lower mitochondrial efficiency and higher oxidative stress [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
+
+In _L.
+delicata_, previous research has demonstrated that high incubation temperatures decreased mitochondrial efficiency in the liver [@crino2024eggs].
+Our study shows, however, that mitochondrial physiology in _L.
+delicata_'s brain is robust to both incubation temeprature and prenatal CORT.
+Nevertheless, brain effects of prenatal conditions can be region-dependent, with some areas having higher sensitivity to prenatal stressors than others [@coomber_independent_1997; @sakata_neural_2000].
+As such, further investigations are necessary to determine whether mitochondrial responses to prenatal conditions vary between different brain regions in _L.
+delicata_ and their potential effects on behaviour.
+
+#### Mitochondrial physiology is not related to learning abilities
+Mitochondria play a critical role in cognitive abilities by synthesizing energy [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018], and through effects on cell senescence and death caused by oxidative damage [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022].
+For example, donut-shaped mitochondrial density — associated with higher ROS production — on presynaptic neurons impairs visuospatial working-memory in rhesus monkeys (_Macaca mulatta_) [@hara_presynaptic_2014].
+However, we found no significant relationships between mitochondrial physiology and learning rate in _L.
+delicata_.
+
+Our findings suggest that energy production is not a limiting factor for spatial learning in _L.
+delicata_.
+The lack of association between mitochondrial density, mitochondrial potential, and learning abilities may reflect the species' ability to maintain energy production above the threshold required to prevent cognitive dysfunction.
+Alternatively, _L.
+delicata_ can sustain cognitive function despite fluctuations in energy availability.
+
+However, we found a significant positive relationship between mitochondrial potential and ROS in the medial cortex.
+This aligns with the role of mitochondrial membrane potential in driving ATP synthesis and electron transport, processes that inherently generate ROS as byproducts [@rice2002brain].
+Normally, moderate ROS levels serve essential signaling functions and are controlled by antioxidants [@rice2002brain; @terman2006oxidative].
+However, excessive ROS production can exceed antioxidant action, leading to oxidative stress and cellular dysfunction [@rice2002brain; @terman2006oxidative].
+In our study, we did not find any significant relationship between ROS and oxidative damage, suggesing that antioxidant action was sufficient to counteract ROS effects.
+However, higher energetic demands in the medial cortex may overcome antioxidant capactities and elevate oxidative stress.
+
+Additionally, DNA damage or lipid peroxidation in the medial cortex were not linked to spatial learning.
+This suggests that oxidative damage levels were not high enough to impair cognitive function or that compensatory mechanisms counteracted any negative effects.
+Nevertheless, since oxidative damage accumulate over time [@terman2006oxidative], the relationships between oxidative damage and cognitive dysfunction could become more pronounced in older individuals [@hara_presynaptic_2014].
+Future research should investigate the long-term consequences of mitochondrial activity on brain health and the potential mechanisms that sustain cognitive resilience in _L.
+delicata_.
+
+#### Conclusions
+We found that spatial learning in _L.
+delicata_ was not influenced by prenatal CORT or temperature, nor was mitochondrial physiology in the medial cortex.
+Additionally, there was no relationship between mitochondrial function and learning abilities in a spatial task.
+However, we found that ROS production increased with mitochondrial potential, suggesting that links between energy production, mitochondrial respiration, and oxidative stress may become more pronounced under stressful conditions or with aging, potentially influencing cognitive function over time.
+
+### Ethics
+Both the breeding animals and the experimental lizards were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate levels of humidity (see Methods 1: Animal husbandry in Supplementary Material for details).
+Euthanasia was performed by peritoneal injection of a 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
+We monitored the animals to ensure there was no irritation from the agent as indicated by distressed animals.
+Before disposing of the lizard, we confirmed the absence of righting response and the pinching reflex in one of the front limbs.
+All the protocols complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33).
+
+### Data accessibility  
+All data, data description, and R code are available in public repository [https://github.com/Pablo-Recio/CORT-Temp_Spatial_learning](https://github.com/Pablo-Recio/CORT-Temp_Spatial_learning).
+
+### Declaration of AI use
+We declare Chat GPT was used mainly for questions related to coding and data analyses.
+All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+
+### Authors' contributions
+P.R.: conceptualization, methodology, data collection, data curation, formal analysis, writing—original draft, writing—review and editing; D.C.L.: conceptualization, methodology, data collection, writing—review and editing; O.C.: conceptualization, methodology, writing—review and editing; C.F.: conceptualization, methodology, funding acquisition, writing—review and editing; D.N.: conceptualization, methodology, funding acquisition, project administration, resources, supervision, writing—review and editing.
+All authors gave final approval for publication and agreed to be held accountable for the work performed therein.
+
+### Conflict of interest declaration  
+We declare we have no competing interests.
+
+### Funding
+This work was supported by a National Australian University fellowship (to P.R. and D.C.L.), and the Australian Research Council (grant no. DP210101152) to D.N. and C.R.F.
+
+### Acknowledgements
+We thank Dr.
+Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
+We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards.
+We also thank ANU MakerSpace, where we designed and built the prototypes of the 3D printed feeders.
+Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.
 
@@
 
+#### Methods 1: Animal husbandry
+
+*Breeding colony* - Juveniles of _L.
+delicata_ came from a breeding colony established in the laboratory since 2019.
+This colony consisted of 270 adults housed in plastic containers (41.5 L x 30.5 W x 21 H cm) with six lizards (two males and four females) per enclosure.
+Enclosures were provided with shelter, nonstick matting, and several small water dishes.
+The lizards were fed approx.
+40 mid-size crickets (_Acheta domestica_) per enclosure three days a week, and water was given daily.
+The crickets were dusted with calcium weekly and multivitamin and calcium biweekly.
+Room temperatures were set to 22-24 ºC, but we also provided the enclosures with a heat chord and a heat lamp following a 12 h light:12 h dark cycle keeping warm side of enclosures is usually at 34 ºC.
+
+*Eggs collection and incubation* - Between mid-November 2023 to mid-January 2024, we placed a small box (12.5 L x 8.3 W x 5 H cm) with moist vermiculite in one side of the communal enclosures to provide females with a place to lay the eggs.
+These boxes were checked three days a week.
+After egg collection, we measured length and width with a digital caliper to the nearest 0.1 mm and weighted the eggs with a digital scale ± 0.001g error.
+Then eggs were treated with CORT or vehicle (see CORT and temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite).
+The cups were covered with cling wrap to retain moisture and left in two incubators at two different temperatures (see CORT and temperature manipulation below) until hatching.
+
+*Hatchlings* - Incubators were checked three times a week for hatchlings.
+Lizards were measured and weighed immediately after hatching.
+snout-vent length (SVL) and tail length (TL) were measured to the nearest millimeter, and weight was recorded using a digital scale with an accuracy of ± 0.001 g.
+Hatchlings were then placed in individual enclosures (18.7L x 13.2W x 6.3H cm) with nonstick matting and a small water dish.
+Watering, feeding, and temperature conditions were maintained as for adults (see above).
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Methods 2: flow cytometry
+
+#### Brain mitochondrial activity
+_Homogenates_
+After the medial cortex was extracted, it was transferred immediately to 1.5mL centrifuge tubes containing 100µL of cold 1X PBS and kept on ice until further processing.
+The tissue was mechanically homogenized by placing the tissue in the well of a 100µm mesh filter (pluriStrainer) affixed atop a 1.5mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
+The resulting homogenate was then rinsed through the filter with 1000µL of cold 1X PBS to prepare a homogenate suspension.
+Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500µL 1x PBS.
+This step was to performed to remove cellular debris from homogenates.
+
+From each 500µL suspension of homogenate collected on a given trial day, we first added 100µL of homogenate to a pooled sample of each tissue type to use for single-color controls, and the remaining 400µL of homogenate was split among two 200µL aliquots.
+One aliquot was used fresh to measure mitochondrial function (mitochondrial density, membrane potential, ROS), one aliquot was cryopreserved for later measurements of oxidative damage (8-OHdG, lipid peroxidation), and the third aliquot was cryopreserved for a different experiment.
+
+_Staining fresh samples_
+From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50µL of homogenate in duplicates (2 wells per homogenate).
+To each replicate well, we added 5µL of a fluorescent probe mix containing equal parts 5µM MitoTracker Deep Red FM, 2.5µM MitoTracker Orange CMTMRos, and 50µM MitoSOX Red.
+We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
+We then added 5µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
+We then loaded 6 wells with 50µL of homogenate taken from each pooled homogenate suspension (12 wells total), which were to be negative and single-color controls.
+One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5µL of one of 5µM MitoTracker Deep Red FM, 2.5µM MitoTracker Orange CMTMRos, 50µM MitoSOX Red, or #µM Hoechst 33342 Nuclear Viability Dye.
+Any remaining pooled homogenate was fixed and frozen as previously described.
+We incubated the loaded plate at 32 °C for 30 minutes to stain and then diluted the samples with 50µL cold 1x PBS to halt the staining process.
+Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours.
+Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).
+
+The aliquots destined to the analysis of oxidative damage were stained with 20µL (10µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20µL of 100µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32°C for 20 minutes.
+Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet and continue with cryopreservation.
+
+_Cryopreservation_
+Aliquots were fixed first by adding the samples into a 1mL solution of 1% Neutral-Buffered Formalin (as a fixative agent) and incubating them at 32°C for 20 min.
+Then we centrifuged the samples at 800-1200xg for 10 min to pellet cells, removed the supernatant, and resuspended the cells (from now on this process will be called as 'washing') in 1mL cold 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant).
+Samples were stored at -20°C until oxidative damage assays.
+
+_Staining cryopreserved samples_
+Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in hot water.
+We washed each thawed sample twice, the first time resuspending the pelleted cells in 1000µL warm 1X Tris-EDTA, and the second time in 200µL warm 1X PBS containing 5mM (6.14 mg/mL) digitonin.
+We incubated the samples at 32°C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200µL 1X PBS.
+We added 20µL of 70µM 8-OHdG Polyclonal Antibody to each sample, and we left the homogenate overnight (~12 hours) for the antibody to bind to 8-OHdG, a marker of oxidative damage on DNA.
+The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32°C for 20 minutes.
+After the cells had been tagged with 8-OHdG antibodies and counterstained, we washed the cells once more and resuspended the pellet in 400µL of 1X PBS.
+Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate.
+Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated.
+We then loaded a 96-well plate with 100µL of each single-color control and 100µL in triplicate of each sample.
+We performed all flow cytometry assays on samples within 48-hours of thawing the samples.
+
+_Flow cytometry_
+All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
+Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v.
+8.0.1) and BD CS&T fluorescent beads (Lot No.
+30664) diluted at 2 drops to 150µL 1X PBS.
+During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector.
+These thresholds were chosen to filter small debris or inviable or non-intact cells from our observations.
+The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment.
+Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events.
+For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels.
+For the oxidative damage assay, we recorded data from the following channels: FSC (425; forward scatter), SSC (300; side scatter), Alexa-Fluor 488 (275; 8-OHdG Antibody + Alexa-Fluor 488 conjugate), BUV-496 (525; Hoechst 33342), and PE-Cy5 (850; BODIPY 665/676).
+Fluorescent intensity data was collected via the BD FACSDiva (v.
+8.0.1) software, with no compensation applied during data collection, and all on a linear scale (detectable range of 0-252166).
+We recorded data for both the area and height of the fluorescent signal, but only used the area in downstream analyses, with height being recorded for the sake of quality control.
+Data was exported from BD FACSDiva as individual *.fcs (“flow cytometry standard”) files for each sample, then imported into FlowJo (v.
+10.1) for processing.
+
+_Data processing_
+In FlowJo v.
+10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events.
+We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains.
+We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103).
+For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103).
+For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103).
+When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population.
+We used the backgated population for compensation of fluorescent spillover between different fluorescent probes.
+To account for fluorescent spillover, we used a traditional compensation matrix using the compensation function of FlowJo v.10.1.
+We identified the “positive” population for each channel as the brightest ~2.5% of the distribution of observations in the respective single-color control for that channel and used unstained controls as a universal negative.
+We visually inspected the compensation matrix and its effects on population distributions for under- and over-compensation, whereupon we changed the compensation matrix manually until data was properly compensated.
+We applied the compensation matrix to all samples for downstream processing.
+Following compensation, we again gated the data following the same process as for backgating but using the compensated parameters for each channel.
+Following gating, we exported the geometric mean (mean fluorescent intensity; MFI) and robust confidence-values for each channel for each sample.
+For analysis, we exported summary statistics of only the area of the fluorescent signal.
+Although we exported robust confidence values for checking repeatability between replicate samples, we used the mean and standard deviation of geometric means among replicates for each individual as our main response variables in analyses.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
 #### Results of the final models
 
-Table S**13**. R^2^ values of the final models.
-```{r, table_bayesR2}
-#| label: table_bayesR2
-#| tbl-cap: "BayesR2 values of the final models"
-#
-data_bayes <- data.frame(
-  Model = character(0),
-  Mean = numeric(0),
-  Error = numeric(0),
-  Q2_5 = numeric(0),
-  Q97_5 = numeric(0)
+```{r, fig_learning_raw}
+#| label: fig_learning_raw
+#
+source(here("R", "func.R"))
+# Modify the df to plot the raw data
+learning_df_plot <- learning_df %>%
+  mutate(treatment = factor(trt, levels = c("B_23", "A_23", "B_28", "A_28"),
+                      labels = c("B_23" = "CORT-Cold (n=20)",
+                                 "A_23" = "Control-Cold (n=20)",
+                                 "B_28" = "CORT-Hot (n=19)",
+                                 "A_28" = "Control-Hot (n=20)"))) %>%
+  group_by(treatment, day) %>%
+  mutate(mean_raw = mean(na.omit(errors)),
+         sd_raw = sd(na.omit(errors))) %>%
+data.frame()
+# Get the plot and combine it with the modeled data
+fig_raw_learning <- plot_errorsday(fig_A_df) +
+  geom_point(data = learning_df_plot, aes(x = day, y = mean_raw, color = treatment), size = 1.5) +
+  scale_color_manual(values = c("CORT-Cold (n=20)" = "#00008B",
+                                "Control-Cold (n=20)" = "#68bde1",
+                                "CORT-Hot (n=19)" = "#b50101",
+                                "Control-Hot (n=20)" = "#fa927d")) +
+  facet_grid(~treatment) +
+  ylim(c(min(na.omit(learning_df_plot$mean_raw)), max(na.omit(learning_df_plot$mean_raw)))) +  # Adjust y-axis limits
+  theme(
+    axis.title = element_text(size = 12, family = "Times"),
+    axis.text = element_text(size = 10, family = "Times"),
+    legend.position = "bottom",
+    legend.title = element_blank()
   )
-models <- c("mit_density_def",
-            "mit_potential_def",
-            "ROS_def",
-            "DNAdamage_def",
-            "peroxidation_def",
-            "learning_def")
-#
-for (m in models){
-  mod <- readRDS(here("output/models/", paste0(m, ".rds")))
-  bayes <- bayes_R2(mod)
-  data_bayes <- rbind(data_bayes, data.frame(
-    Model = m,
-    Mean = format_dec(bayes[1], 3),
-    Error = format_dec(bayes[2], 3),
-    Q2_5 = format_dec(bayes[3], 3),
-    Q97_5 = format_dec(bayes[4], 3)
-  ))
-}
-#
-bayes_table_df <- data_bayes %>%
-  mutate(Model = factor(Model,
-                        levels = c("mit_density_def",
-                                  "mit_potential_def",
-                                  "ROS_def",
-                                  "DNAdamage_def",
-                                  "peroxidation_def",
-                                  "learning_def"),
-                        labels = c("mit_density_def" = "Mit density",
-                                  "mit_potential_def" = "Mit potential",
-                                  "ROS_def" = "ROS",
-                                  "DNAdamage_def" = "DNA damage",
-                                  "peroxidation_def" = "Lipid peroxidation",
-                                  "learning_def" = "Learning"))) %>%
-  dplyr::select(Model, Mean, Error, Q2_5, Q97_5) %>%
-  arrange(Model)
-#
-#
-# Create the table
-#
-## Table format
+ggsave("./output/figures/suppl/Figure_S1_learning_raw.png", plot=fig_raw_learning , width = 21, height = 15, units = "cm", dpi = 600)
+knitr::include_graphics("./output/figures/suppl/Figure_S1_learning_raw.png")
+```
+
+Fig S1. Raw data of the learning task. Each point represents the mean number of errors per day. The lines represent the modeled data.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S1. Estimates of the learning slopes per treatment, the 95% CI, and pMCMC values testing the hypothesis that slope is different from 0. In bold, the values that are significant at pMCMC < 0.05.
+```{r, tbl-learning}
+#| label: tbl-learning
+#| tbl-cap: "Learning tasks slopes per treatment"
+source(here("R", "func.R"))
+#
+table_learn_df <- data_fig_learning_slopes %>%
+  group_by(treatment) %>%
+  summarize(mean_slope = format_dec(mean(slopes), 3),
+            q025_slope = format_dec(quantile(slopes, 0.025), 3),
+            q975_slope = format_dec(quantile(slopes, 0.975), 3),
+            pMCMC = format_p(pmcmc(slopes), 3, equal = FALSE)) %>%
+  mutate(`95 CI` = paste0("[", q025_slope, ", ", q975_slope, "]")) %>%
+  dplyr::select(treatment, mean_slope, `95 CI`, pMCMC)
+#
+# Make Table
 set_flextable_defaults(
@@
 #
-bayes_table <- flextable(bayes_table_df) %>%
-  set_header_labels(
-    Region = "Region/Stimulus",
-    Mean = "Mean",
-    Error = "Error",
-    Q2_5 = "2.5%",
-    Q97_5 = "97.5%") %>%
-  align(align = "center", j = c(2:5), part = "body") %>%
-  align(align = "center", j = c(1:5), part = "header") %>%
+table_slopes <- flextable(table_learn_df) %>%
+  align(align = "center", j = c(3,4), part = "body") %>%
+  align(align = "center", j = c(1:4), part = "header") %>%
+  flextable::compose(j = 1, value = as_paragraph("Treatment"), part = "header") %>%
+  flextable::compose(j = 2, value = as_paragraph("Estimated slope"), part = "header") %>%
+  bold(~`pMCMC` == "< 0.05", j = c("pMCMC", "mean_slope", "95 CI", "treatment"),
+       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
+  bold(~`pMCMC` == "< 0.001", j = c("pMCMC", "mean_slope", "95 CI", "treatment")) %>%  # Bold when PMCMC is "<0.001"
   autofit()
 #
-bayes_table
-```
-
-Mean indicates the mean of the R^2^ values obtained from the posteriors of the model, error the estimated SE of the mean and 95% CI indicates the 95% confidence interval around the mean.
-Model indicates the name of the variable analyzed.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S**14**. Summary of the final models for all mitochondrial variables.
-```{r, tbl-results_mit_table}
-#| label: results_mit_table
-#| tbl-cap: "Summary of the final models for all mitochondrial variables."
-#| tbl-label: "results_mit_table"
-#
-source(here("R", "func.R"))
-# 
-# A) Refining the df summarizing the posteriors for OB/Chemical stimulus (post_OB)
-post_mit_refined <- refine_post(post_mit) %>%
-  arrange(Variable, Predictors)
-#
-# B) Create table
-## Table format
-set_flextable_defaults(
- font.family = "Times New Roman",
- font.size = 10)
-#
-mit_table <- flextable(post_mit_refined) %>%
-  align(align = "center", j = c(3:5), part = "body") %>%
-  align(align = "center", j = c(1:5), part = "header") %>%
-  bold(~`PMCMC` < 0.05, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors"),
-       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
-  bold(~`PMCMC` <0.001, j = c("PMCMC", "Estimate Mean", "95% CI", "Predictors")) %>%  # Bold when PMCMC is "<0.001"
-  flextable::compose(i = c(2:4,6:8,10:12,14:16,18:21), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
-  autofit()
-#
-mit_table
-```
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S **!%**. Contrasts between treatments for all the variables analysed.
+table_slopes
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+Table S2.
+Estimate of the contrasts between treatments for all the variables analysed, and pMCMC values testing the hypothesis that contrast is different from 0.
+In bold, the values that are significant at pMCMC < 0.05.
 ```{r tbl-contrasts}
@@
 
-```{r, fig-learning_raw}
-#| label: fig-learning_raw
-#| fig.cap: "Raw data of the learning task."
-#
+#### Results SEM
+
+Table S3. Estimated direct, indirect, and total coefficients from the multivariate model
+```{r, table_sem_results}
+#| label: table_sem_results
+#| tbl-cap: "Structural Equation Models"
+#
+# Table created from the df sem_results_OB (see above)
 source(here("R", "func.R"))
-# Modify the df to plot the raw data
-learning_df_plot <- learning_df %>%
-  mutate(treatment = factor(trt, levels = c("B_23", "A_23", "B_28", "A_28"),
-                      labels = c("B_23" = "CORT-Cold (n=20)",
-                                 "A_23" = "Control-Cold (n=20)",
-                                 "B_28" = "CORT-Hot (n=20)",
-                                 "A_28" = "Control-Hot (n=20)"))) %>%
-  group_by(treatment, day) %>%
-  mutate(mean_raw = mean(na.omit(errors)),
-         sd_raw = sd(na.omit(errors))) %>%
-data.frame()
-# Get the plot and combine it with the modeled data
-fig_raw_learning <- plot_errorsday(fig_A_df) +
-  geom_point(data = learning_df_plot, aes(x = day, y = mean_raw, color = treatment), size = 1.5) +
-  scale_color_manual(values = c("CORT-Cold (n=20)" = "#00008B",
-                                "Control-Cold (n=20)" = "#68bde1",
-                                "CORT-Hot (n=20)" = "#b50101",
-                                "Control-Hot (n=20)" = "#fa927d")) +
-  facet_grid(~treatment) +
-  ylim(c(min(na.omit(learning_df_plot$mean_raw)), max(na.omit(learning_df_plot$mean_raw)))) +  # Adjust y-axis limits
-  theme(
-    axis.title = element_text(size = 12, family = "Times"),
-    axis.text = element_text(size = 10, family = "Times"),
-    legend.position = "bottom",
-    legend.title = element_blank()
-  )
-ggsave("./output/figures/suppl/fig_learning_raw.png", plot=fig_raw_learning , width = 21, height = 15, units = "cm", dpi = 600)
-```
+#
+# Modify database
+table_sem_df <- sem_results %>%
+  mutate(`Direct effects` = paste0(mean_direct_effects, " [", q5_direct, ", ", q95_direct, "]"),
+         `Indirect effects` = paste0(mean_indirect_effects, " [", q5_indirect, ", ", q95_indirect, "]"),
+         `Total effects` = paste0(mean_total_effects, " [", q5_total, ", ", q95_total, "]")) %>%
+  dplyr::select(source, predictor_modulator, `Direct effects`, `Indirect effects`, `Total effects`) %>%
+  mutate(
+      predictor_modulator = case_when(
+        predictor_modulator == "mit_density" ~ "Mitochondrial density",
+        predictor_modulator == "mit_potential" ~ "Mitochondrial potential",
+        predictor_modulator == "DNAdamage" ~ "DNA damage",
+        predictor_modulator == "peroxidation" ~ "Lipid peroxidation",
+        predictor_modulator == "ROS" ~ "ROS",
+        TRUE ~ predictor_modulator
+      ),
+      source = case_when(
+        source == "learn_sem" ~ "Learning",
+        source == "DNA_sem" ~ "DNA damage",
+        source == "perox_sem" ~ "Lipid peroxidation",
+        source == "ROS_sem" ~ "ROS",
+        TRUE ~ source
+      ),
+      `Direct effects` = case_when(`Direct effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Direct effects`),
+      `Indirect effects` = case_when(`Indirect effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Indirect effects`),
+      `Total effects` = case_when(`Total effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Total effects`)) %>%
+  rename(Predictor = predictor_modulator, Response = source) %>%
+  mutate(Response = factor(Response, levels = c("Learning", "DNA damage", "Lipid peroxidation", "ROS")),
+         Predictor = factor(Predictor, levels = c("Mitochondrial density",
+                                                "Mitochondrial potential",
+                                                "ROS",
+                                                "DNA damage",
+                                                "Lipid peroxidation"))) %>%
+  arrange(Response, Predictor)
+#
+# Table format
+set_flextable_defaults(
+ font.family = "Times New Roman",
+ font.size = 10)
+#
+table_sem <- flextable(table_sem_df) %>%
+  align(align = "center", part = "header") %>%
+  flextable::compose(i = c(2:5,7,8,10,11,13), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
+  flextable::hline(i = c(5,8,11), part = "body") %>% 
+  autofit()
+#
+table_sem
+```
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+#### Final univariate models diagnostics (plots)
+
+```{r , plotmod_mitdensity, out.width="70%", fig.align="center"}
+#| label: plotmod_mitdensity
+#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
+#
+mod <- readRDS(here("output/models/mit_density_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S2.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S2.png")
+```
+
+Fig S2. Posterior predictive checks for the model of Mitochondrial Density.
+Formula: mit_density ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_potential, out.width="70%", fig.align="center"}
+#| label: plotmod_potential
+#| caption: "Posterior predictive checks for the model of Mitochondrial Potential."
+#
+mod <- readRDS(here("output/models/mit_potential_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S3.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S3.png")
+```
+
+Fig S3. Posterior predictive checks for the model of Mitochondrial Potential.
+Formula: mit_potential ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_ros, out.width="70%", fig.align="center"}
+#| label: plotmod_ros
+#| caption: "Posterior predictive checks for the model of ROS."
+#
+mod <- readRDS(here("output/models/ROS_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S4.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S4.png")
+```
+
+Fig S4. Posterior predictive checks for the model of ROS.
+Formula: ROS ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_dnadamage, out.width="70%", fig.align="center"}
+#| label: plotmod_dnadamage
+#| caption: "Posterior predictive checks for the model of DNA damage."
+#
+mod <- readRDS(here("output/models/DNAdamage_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S5.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S5.png")
+```
+
+Fig S5. Posterior predictive checks for the model of DNA Damage.
+Formula: DNAdamage ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_peroxidation, out.width="70%", fig.align="center"}
+#| label: plotmod_peroxidation
+#| caption: "Posterior predictive checks for the model of lipid peroxidation."
+#
+mod <- readRDS(here("output/models/peroxidation_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+#
+fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
+ggsave("./output/figures/suppl/Figure_S6.png", plot = fig_mod,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S6.png")
+```
+
+Fig S6. Posterior predictive checks for the model of lipid peroxidation.
+Formula: peroxidation ~ cort * temp + (1|clutch)
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+```{r , plotmod_learn, out.width="70%", fig.align="center"}
+#| label: plotmod_learn
+#| caption: "Posterior predictive checks for the model of learning."
+#
+mod <- readRDS(here("output/models/learning_def.rds"))
+plot_mod <- plot(mod, plot = FALSE)
+#
+plot_mod_1 <- plot_mod[[1]]
+plot_mod_2 <- plot_mod[[2]]
+plot_mod_3 <- plot_mod[[3]]
+#
+ggsave("./output/figures/suppl/Figure_S7A.png", plot = plot_mod_1,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S7A.png")
+#
+ggsave("./output/figures/suppl/Figure_S7B.png", plot = plot_mod_2,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S7B.png")
+#
+ggsave("./output/figures/suppl/Figure_S7C.png", plot = plot_mod_3,
+       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
+knitr::include_graphics("./output/figures/suppl/Figure_S7C.png")
+```
+
+Fig S7. Posterior predictive checks for the model of spatial learning.
+Formula: errors ~ day * cort * temp + (1 + day|lizard_id) + (1|clutch)
 
@@
 
-Table S1. Preliminary results of the models testing for Mitochondrial Density.
-
+Table S4. Preliminary results of the models testing for Mitochondrial Density.
 ```{r, results_preliminary_mitdensity}
@@
 
-Table S2.  Preliminary results of the models testing for Mitochondrial Potential.
+Table S5.  Preliminary results of the models testing for Mitochondrial Potential.
 ```{r, results_preliminary_potential}
@@
 
-Table S3. Preliminary results of the models testing for ROS.
+Table S6. Preliminary results of the models testing for ROS.
 ```{r, results_preliminary_ros}
@@
 
-Table S4. Preliminary results of the models testing for DNA damage.
+Table S7. Preliminary results of the models testing for DNA damage.
 ```{r, results_preliminary_dnadamage}
@@
 
-Table S5. Preliminary results of the models testing for lipid peroxidation.
+Table S8. Preliminary results of the models testing for lipid peroxidation.
 ```{r, results_preliminary_peroxidation}
@@
 
-Table S6. Preliminary results of the models testing for learning.
+Table S9. Preliminary results of the models testing for learning.
 ```{r, results_preliminary_learning}
 #| tbl-cap: "Preliminary results of the models testing for lipid peroxidation"
-#| label: results_preliminary_peroxidation
+#| label: results_preliminary_learning
 #
@@
 Model formula: errors ~ day * cort * temp + sex + age+ (1 + day | lizard_id) + (1|clutch).
-Model convergence was checked through rhat and ess_bulk values.
-**Summary indicates no effect of sex, so it was discarded from the final models.
-However, we saw an effect of age and we included it in our final models.**
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-#### Final models diagnostics (plots)
-
-```{r , plotmod_mitdensity, out.width="70%", fig.align="center"}
-#| label: plotmod_mitdensity
-#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mit_density_def.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX3.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX3.png")
-```
-
-Fig S1. Posterior predictive checks for the model of Mitochondrial Density in Olfactory Bulbs.
-Formula: mean_mitodensity ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_mitdensity_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_mitdensity_OT
-#| caption: "Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_mitodensity_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX4.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX4.png")
-```
-
-Fig S2. Posterior predictive checks for the model of Mitochondrial Density in Optic Tecta.
-Formula: mean_mitodensity ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_potential_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_potential_OB
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_potential_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX5.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX5.png")
-```
-
-Fig S3. Posterior predictive checks for the model of Mitochondrial Potential in Olfactory Bulbs.
-Formula: mean_potential ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_potential_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_potential_OT
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_potential_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX6.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX6.png")
-```
-
-Fig S4. Posterior predictive checks for the model of Mitochondrial Potential in Optic Tecta.
-Formula: mean_potential ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_ros_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_ros_OB
-#| caption: "Posterior predictive checks for the model of ROS Production in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_ros_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX7.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX7.png")
-```
-
-Fig S5. Posterior predictive checks for the model of ROS Production in Olfactory Bulbs.
-Formula: mean_ros ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_ros_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_ros_OT
-#| caption: "Posterior predictive checks for the model of ROS Production in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_ros_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX8.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX8.png")
-```
-
-Fig S6. Posterior predictive checks for the model of ROS Production in Optic Tecta.
-Formula: mean_ros ~ cort * temp + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_dnadamage_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_dnadamage_OB
-#| caption: "Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_dnadamage_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX9.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX9.png")
-```
-
-Fig S7. Posterior predictive checks for the model of DNA Damage in Olfactory Bulbs.
-Formula: mean_dnadamage ~ cort * temp + age_euthanasia + sex + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_dnadamage_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_dnadamage_OT
-#| caption: "Posterior predictive checks for the model of DNA Damage in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_dnadamage_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX10.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX10.png")
-```
-
-Fig S8. Posterior predictive checks for the model of DNA Damage in Optic Tecta.
-Formula: mean_dnadamage ~ cort * temp + sex + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_peroxidation_OB, out.width="70%", fig.align="center"}
-#| label: plotmod_peroxidation_OB
-#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs."
-#
-mod <- readRDS(here("output/models/mean_peroxidation_def_OB.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX11.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX11.png")
-```
-
-Fig S9. Posterior predictive checks for the model of Lipid Peroxidation in Olfactory Bulbs.
-Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_peroxidation_OT, out.width="70%", fig.align="center"}
-#| label: plotmod_peroxidation_OT
-#| caption: "Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta."
-#
-mod <- readRDS(here("output/models/mean_peroxidation_def_OT.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX12.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX12.png")
-```
-
-Fig S10. Posterior predictive checks for the model of Lipid Peroxidation in Optic Tecta.
-Formula: mean_peroxidation ~ cort * temp + age_euthanasia + (1|clutch)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_tD_Chemical, out.width="70%", fig.align="center"}
-#| label: plotmod_tD_Chemical
-#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials."
-#
-mod <- readRDS(here("output/models/t_D_def_Chemical.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX1.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX1.png")
-```
-
-Fig S11. Posterior predictive checks for the model of Detection Latency (t_D) in Chemical trials.
-Formula: t_D~ cort * temp + motivation + cort:motivation + (1|clutch) + (1|lizard_id)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-```{r , plotmod_tD_Visual, out.width="70%", fig.align="center"}
-#| label: plotmod_tD_Visual
-#| caption: "Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials."
-#
-mod <- readRDS(here("output/models/t_D_def_Visual.rds"))
-plot_mod <- plot(mod, plot = FALSE)
-#
-plot_mod_1 <- plot_mod[[1]]
-plot_mod_2 <- plot_mod[[2]]
-#
-fig_mod <- plot_grid(plot_mod_1, plot_mod_2, ncol = 1)
-ggsave("./output/figures/suppl/Figure_SX2.png", plot = fig_mod,
-       width = 20, height = 28, units = "cm", dpi = 600, bg = "white")
-knitr::include_graphics("./output/figures/suppl/Figure_SX2.png")
-```
-
-Fig S12. Posterior predictive checks for the model of Detection Latency (t_D) in Visual trials.
-Formula: t_D~ cort * temp + (1|clutch) + (1|lizard_id)
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
+Model convergence was checked through rhat and ess_bulk values. Summary indicates no effect of sex or age, so they were discarded from the final models.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-03-16 .vdoc.r

@@ -19,6 +19,8 @@
 #
 #
 #
+#
+#
 #| label: setup
 pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
 #
@@ -98,6 +100,7 @@ for(k in 1:length(hormone)){
 #
 #
 #
+cat("\\newpage")
 #
 #
 #
@@ -114,11 +117,26 @@ for(k in 1:length(hormone)){
 #
 #
 #
-#| label: fig-Methods
-#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."
-
-knitr::include_graphics("./Others/SPAL_METH.svg")
-
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
 #
 #
 #
@@ -184,94 +202,6 @@ m_def_learn <- fit_m(df = learning_df,
 #
 #
 #
-#| label: organise_posteriors_mit
-#
-# Organising the posteriors of the previous models for tables and figures
-#
-source(here("R", "func.R"))
-#
-#
-post_mit <- data.frame()
-# Names posteriors:
-names <- c("m_def_mit_density", "m_def_mit_potential", "m_def_ROS", "m_def_DNAdamage", "m_def_peroxidation")
-#
-# Organising the results
-for (pos in names) {
-  model <- get(pos)      # Get the model from the global environment
-  post_result <- tidy_post(model)        # Apply tidy_post to each model
-  # Add a new column to the rmodel
-  post_result$Model <- pos
-  post_mit <- bind_rows(post_mit, post_result)
-}
-#
-#
-#
-# Extracting the posteriors for the models with mitochondrial variables and the values of interest. Here, I am creating dfs for each of the variables with the values for all the prenatal conditions to make contrasts easier to write.
-#
-MD <- post_values(m_def_mit_density, "none")
-MP <- post_values(m_def_mit_potential, "none")
-ROS <- post_values(m_def_ROS, "none")
-DNA <- post_values(m_def_DNAdamage, "none")
[…]

### 2025-03-23 ms.qmd
updates

@@
 ---
-title: "Spatial learning and brain mitochondria are robust to prenatal conditions in the delicate skink"
+title: "Prenatal conditions do not affect brain physiology and learning in a lizard"
 docx: 
@@
 crossref:  
-  fig-title: Fig    # (default is "Figure")
+  fig-title: Figure    # (default is "Figure")
   title-delim: —     # (default is ":")
-  fig-prefix: Fig.   # (default is "Figure")
+  fig-prefix: Figure   # (default is "Figure")
   tbl-prefix: Table  # (default is "Table")
@@
 ## Abstract
-Learning is highly sensitive to environmental conditions, including early thermal environments and prenatal exposure to glucocorticoids.
-These factors influence mitochondrial activity, which plays a crucial role in neural development and synaptic plasticity, ultimately shaping cognitive abilities.
-Here, we investigated the interactive effects of prenatal temperature and corticosterone (CORT) — the primary glucocorticoid in reptiles — on spatial learning and brain mitochondrial activity in the delicate skink (_Lampropholis delicata_).
-We supplemented eggs with either CORT or a control vehicle and incubating them under two thermal regimes: cold (23 ± 3°C) or hot (28 ± 3°C).
-Juveniles were then tested in a spatial learning task over 40 days, after which mitochondrial activity in the medial cortex was assessed.
-Our results revealed that both mitochondrial physiology and spatial learning in _L.
-delicata_ are robust to prenatal temperature and CORT exposure.
-In addition, we found no significant relationship between mitochondrial function and cognitive abilities.
-However, mitochondrial potential increased ROS levels, which could impact oxidative stress under high metabolic demands or low antioxidant capacity.
-These findings highlight the cognitive and physiological resilience of _L.
-delicata_ to early-life environmental challenges.
-Future research should explore whether this robustness extends to other brain regions and life stages.
+Early environmental factors like heat or stress hormones can impair learning through brain metabolic function, crucial for neural development and synaptic plasticity.
+However, whether early environments always result in cognitive impairment through changes in neural physiology is not well-established outside of a few model systems.
+Here, we investigated the effects of prenatal temperature and corticosterone (CORT) on brain mitochondrial activity and spatial learning in the delicate skink (_Lampropholis delicata_).
+We treated eggs with either CORT or a control vehicle and incubated at cold (23 ± 3°C) or hot (28 ± 3°C) temperatures.
+Juveniles were tested in a spatial learning task over 40 days after which mitochondrial function in the medial cortex was assessed.
+Despite among-individual variation in learning ability, mitochondrial physiology and spatial learning in _L.
+delicata_ remained robust to prenatal temperature and CORT exposure.
+No significant relationship was found between mitochondrial function and cognitive performance, contrary to predictions.
+Additionally, increased metabolic capacity correlated with higher ROS production but did not affect oxidative damage, possibly due to protective mechanisms.
+These findings highlight the physiological and cognitive resilience of _L.
+delicata_ to early-life challenges.
+Future research should explore whether this robustness extends to other brain regions, cognitive domains, and life stages.
 
@@
 
-Learning - the acquisition and consolidation of new information [@dukas_evolutionary_2004] - enables individuals to create new associations between events, which can be essential for coping with environmental changes [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
+Learning - the acquisition and consolidation of new information [@dukas_evolutionary_2004] - enables animals to create new associations between events, which can be essential for coping with environmental change [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
 However, the capacity to form these associations varies among individuals, potentially affecting their responses to environmental challenges [@ward2016ecological; @welklin2024spatial].
-Animals that learn faster may exploit novel resources better or may start responding to new threats more efficiently than slow learners [@ward2016ecological].
-Consequently, individual differences in learning abilities can have profound effects on survival and reproductive output, ultimately influencing population growth rates and stability [@ward2016ecological; @welklin2024spatial].
+Learning faster may imply exploiting novel resources better or responding to new threats more efficiently [@ward2016ecological].
+Consequently, individual differences in learning abilities can have profound effects on fitness, ultimately influencing population growth rates and stability [@ward2016ecological; @welklin2024spatial].
 Understanding the mechanisms that drive these differences is therefore crucial for predicting how populations will respond to novel conditions.
 
-Factors like age, sex, or early- life conditions can have important effects on learning abilities [@szuran_water_1994; @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @carazo_sex_2014; @noble_age-dependent_2014].
-Early developmental environment is particularly influential, as the brain is highly sensitive to environmental inputs during this period [@zhu_prenatal_2004].
+Factors like age, sex, or early-life conditions can have important effects on learning abilities [@szuran_water_1994; @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @carazo_sex_2014; @noble_age-dependent_2014].
+Adverse developmental environments are particularly influential, as the brain is highly sensitive to environmental inputs during early life [@zhu_prenatal_2004].
 In fact, early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
 Among the physiological mechanisms underlying learning, mitochondrial activity is considered a key factor [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
-The neural processes involved learning impose substantial energetic demands [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], making efficient mitochondrial respiration essential [@du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
-Additionally, learning depends on a dense network of functional neurons [@lefebvre_taxonomic_2011; @amiel_effects_2017], which can be compromised by excessive oxidative stress byproduct of mitochondrial function [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
-Studies on mammals have shown the pervassive effects of mitochondrial physiology on cognitive abilities [@zhu_prenatal_2004; @hara_presynaptic_2014; @cao2014ampk].
+The neural processes involved in learning impose substantial energetic demands [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022], making efficient mitochondrial respiration essential [@du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
+Additionally, learning depends on a dense network of functional neurons [@lefebvre_taxonomic_2011; @amiel_effects_2017], which can be compromised by excessive oxidative stress - a byproduct of mitochondrial metabolism [@finkel_oxidants_2000; @zhu_prenatal_2004; @du_dynamic_2009; @gong_chronic_2011; @hoffmann_mitochondrion_2018].
+Studies on mammals have shown the pervasive effects of mitochondrial physiology on cognitive abilities [@zhu_prenatal_2004; @hara_presynaptic_2014; @cao2014ampk].
 However, the extent to which these effects are generalizable to other taxa remains largely unexplored.
@@
 Since mitochondria are maternally inherited, maternal condition plays a fundamental role in shaping offspring mitochondrial activity [@picard_mitochondria_2014].
-In particular, maternal stress can strongly influence how mitochondria operate in the offspring [@zhu_prenatal_2004].
-Stressful situations elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryo [@uller_sex-specific_2009] altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].
-For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) [@zhu_prenatal_2004; @haussmann_embryonic_2012; @cao2014ampk].
-
-Temperature can be a significant source of maternal stress in ectotherms.
-Thermal environments outside of the optimal range can elevate GCs in the mother, which can be passed to the offspring [see @Crino_2023].
-In addition, temperature can also directly influence offspring development, particularly during early life [@noble_developmental_2018; @crino2024eggs].
+Maternal stress can also influence how mitochondria operate in offspring [@zhu_prenatal_2004].
+Stressful situations experienced by mothers can elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryos [@uller_sex-specific_2009] altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].
+For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) with impacts on spatial learning abilities [@zhu_prenatal_2004; @haussmann_embryonic_2012; @cao2014ampk].
+
+Temperature is also a significant source of maternal stress in ectotherms.
+Thermal environments outside of the optimal range can elevate GCs in mothers, which can be passed to the offspring [see @Crino_2023].
+Additionally, temperature can also directly influence offspring development, particularly during early life [@noble_developmental_2018; @crino2024eggs].
 The prenatal thermal environment plays a crucial role in shaping mitochondrial function, affecting energy metabolism and oxidative stress [@stier2022experimental; @crino2024eggs].
-For instance, hatchlings of the red-eared slider turtle (_Trachemys scripta elegans_) incubated at 26.5 ºC exhibited greater oxidative damage than those incubated at temperatures 0.7 to 1.2 ºC warmer [@treidel2016temperature].
 Thus, the combined effects of prenatal GCs and temperature may profoundly influence mitochondrial function, with important consequences for brain development and cognition.
-However, the extent to which prenatal GCs and temperature interact to shape cognitive abilities via mitochondrial physiology remains largely unknown.
+However, the extent to which prenatal GCs and temperature interact to shape cognitive abilities via mitochondrial physiology remains largely unknown outside of a few model species [@zhu_prenatal_2004; @haussmann_embryonic_2012; @cao2014ampk].
 
@@
 We hypothesized that prenatal CORT and temperature would influence learning abilities through impacts on brain mitochondrial activity.
-Specifically, we predicted that prenatal CORT would decrease energy production while increasing oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
-Similarly, high temperatures would decrease mitochondrial efficienty [@zavorka_climate_2021; @crino2024eggs], but with the potential to reduce oxidative stress [@treidel2016temperature].
+Specifically, we predicted that prenatal CORT would decrease energy production while increasing oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs], especially if CORT can alter cellular components increasing ROS production without enhancing ATP synthesis or making them more vulnerable to oxidative damage.
+Similarly, high temperatures would decrease mitochondrial efficiency [@zavorka_climate_2021; @crino2024eggs], but with the potential to reduce oxidative stress [@treidel2016temperature].
 We further predicted that the combined effects of CORT and temperature would lead to complex interactions, with both factors negatively impacting mitochondrial efficiency but having opposite effects on oxidative stress.
 These effects would lead to differences in learning abilities, that could be affected by the balance between energy production and oxidative stress [@du_dynamic_2009; @picard_energetic_2018; @alexandrov_neuronal_2022].
-By examining these interactions, we aim to clarify how prenatal environmental factors shape learning abilities through mitochondrial function, shedding light on the role of early-life conditions in cognitive development and its mechanismsm.
+By examining these interactions, we aim to clarify how prenatal environmental factors shape learning abilities through mitochondrial function, shedding light on the role of early-life conditions in cognitive development and associated mechanismsm.
 
@@
 
-#### Animals' origin and collection
-Subjects came from a breeding colony established in the laboratory since 2019.
-It consisted in 270 adults living in groups of two males and four females.
-Eggs were collected from these groups between mid-Novemnber 2023 to mid-January 2024.
-After collecting the eggs, we treated them with CORT or vehicle and incubated them under two different temperature regimes (see below).
-Clutch and egg identity were asigned immediately after collection, and the eggs were incubated in individual cups until hatching.
+#### Experimental animals
+Lizards came from a breeding colony established in the laboratory in 2019 from wild populations in Sydney, Australia.
+The colony consisted of 270 adults living in groups of two males and four females.
+Eggs were collected from these groups between November 2023 to January 2024.
+After collecting the eggs, we treated them with CORT or a vehicle control and incubated them under two different temperature regimes (see below).
+Clutch and egg identity were assigned immediately after egg collection, and the eggs were incubated in individual cups until hatching.
 Hatchlings were also kept in individual enclosures until the end of the experiment.
-For details on husbandry and breeding conditions, see Methods 1: Animal husbandry in Supplementary Material.
-
-#### CORT and Temperature manipulation
-We tested the combined effects of prenatal CORT and temperature by elevating CORT concentrations in eggs and then incubating them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC) in a 2x2 factorial design.
+For details on husbandry and breeding conditions see _Methods 1: Animal husbandry in Supplementary Material_.
+
+#### Corticosterone and temperature manipulations
+We tested the combined effects of prenatal CORT and temperature by elevating CORT concentrations in eggs and then incubating them under one of two temperature regimes (Cold - 23 ± 3 ºC or Hot - 28 ± 3 ºC).
 We used a partial split clutch design where eggs from a given clutch were distributed equally across the four treatments when clutch sizes were larger than four and randomly across treatments when less than four.
@@
 No.
-C2505) dissolved in 100% ethanol at a final 10 pg CORT/mL concentration (CORT treatment), while Control eggs received an equal volume of 100% Ethanol.
+C2505) dissolved in 100% ethanol at a final concentration of 10 pg CORT/mL (CORT treatment), while Control eggs received an equal volume of 100% Ethanol.
 CORT dose increased mean yolk CORT levels ~3.7x higher than control eggs in previous studies [@crino2024eggs].
-Eggs were then incubated in one of the two previously mentioned temperature regimes that are within the natural limits in _L.
+Eggs were then incubated in one of the two previously mentioned temperature regimes that are within the natural limits of nest temperatures in _L.
 delicata_ [@cheetham2011embryonic].
 
-#### Learning  
+#### Spatial learning  
 The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact.
-In each trial, the lizards were placed by hand in the center of maze and left to acclimate for two minutes.
-During acclimatation, the central area was surrounded by a yellow device mounted on a pulley system.
+In each trial, lizards were placed by hand in the center of maze and left to acclimatise for two minutes.
+During acclimatisation, the central area was surrounded by a yellow device mounted on a pulley system.
 At the start of each trial, this device was lifted to startle the lizard.
@@
 We assessed lizards' spatial learning using external cues for navigation while avoiding intra-maze cues.
-To avoid subtle intra-maze cues, the maze was replaced every three trials with one of four identical mazes.
+Subtle intra-maze cues were avoided by replacing the maze every three trials with one of four identical mazes.
 Each replacement preserved the correct arm’s orientation and the maze’s position within the room, ensuring consistency for each individual.
@@
 To control for potential side biases, the correct arm was randomly assigned to one of the six arms for each trial.
-We employed four different maze orientations (North-West, South-West, South, and South-East), and the number of lizards assigned to each orientation was counterbalanced across treatments.
+We employed four different maze orientations counterbalancing the number of lizards assigned to each orientation across treatments.
 
@@
 
-#### Brain mitochondrial activity
-Immediately following the completion of the tests, we asessed mitochondrial physiology in the brain of all remaining lizards (see below).
+#### Brain mitochondrial function
+Immediately following the completion of the tests, we quantified mitochondrial physiology in the brain of lizards.
 We euthanized lizards via peritoneal injection 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic), followed by decapitation.
-Before decapitation we evaluated lizard's righting response and the pinching reflex in one of the front limbs.
-
-After decapitation, we extracted the medial cortext in the telencephalon.
-We selected the medial cortex because is considered homologous to the mammalian hippocampus, where spatial cognition is encoded [@rodríguez2002conservation; @naumann2015reptilian].
-The tissue was transferred to 1xPBS solution and then homogenized mechanically using a 100 µm mesh filter (pluriStrainer).
-The resultant homogenate was divided into two aliquotes: one was used fresh for mitochondrial density, membrane potential, and ROS assays; and the other was cryopreserved for later measurements of DNA damage and lipid peroxidation.
+Before decapitation we evaluated lizard righting responses and the pinching reflexes in one of the front limbs.
+
+We extracted the medial cortex in the telencephalon as this brain region is considered homologous to the mammalian hippocampus, where spatial cognition is encoded [@rodríguez2002conservation; @naumann2015reptilian].
+The tissue was transferred to 1X PBS solution and then homogenized mechanically using a 100 µm mesh filter (pluriStrainer).
+The resultant homogenate was divided into two aliquots: one was used fresh for measuring mitochondrial density, membrane potential - a metric of mitochondrial metabolic capacity [@martinez2016tca] - and superoxide (ROS) production; and the other was cryopreserved for later measurements of DNA damage and lipid peroxidation.
 
 Fresh homogenate suspensions were stained with 5 µM MitoTracker Deep Red FM, 2.5 µM MitoTracker Orange CMTMRos, and 50 µM MitoSOX Red.
-We used these fluorescent probes to assess mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
-We also added 5 µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample to differentiate viable cells from debris.
+We used these fluorescent probes to assess mitochondrial density, membrane potential, and ROS, respectively.
+We also added 5 µL of 10 µg/mL Hoechst 33342 Nuclear Viability Dye to each sample to differentiate viable cells from debris.
 These samples were analyzed in a flow cytometer within 2 hours of brain extraction.
 
-The aliquotes reserved for oxidative damage assays were stained with 20 µL of #µM Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
+The aliquots reserved for oxidative damage assays were stained with 20 µL of 10 µg/mL Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
 These dyes were used to measure cell viability and lipid peroxidation, respectively.
@@
 Then, cell membranes were permeabilized in 200 µL of 1X PBS containing 5mM (6.14 mg/mL) digitonin.
-Following membrane permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody - a marker of DNA damage - and allowed them to incubate overnight (~12 hours).
-The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyzed the samples in the the flow cytometer.
+Following permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody - a marker of DNA damage - and allowed them to incubate overnight (~12 hours).
+The following day we counterstained the cells with 20µL of 100 µg/mL H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyzed the samples using the flow cytometer.
 Oxidative damage assays were performed within 6 months of the initial analysis of fresh samples.
@@
 Flow cytometry was performed on a Becton Dickson LSRFortessa X-20 flow cytometer using the default wavelength filters.
-The detectors and voltage settings for each assay were determined during pilot trials and and remained consistent throughout the experiment.
+Detectors and voltage settings for each assay were determined during pilot trials and remained consistent throughout the experiment.
 Data was processed using FlowJo (v.
 10.1) software.
-We obtained the mean fluorescent intensity for mitochondrial density, mitochondrial potential, ROS production, DNA damage, and lipid peroxidation.
-For further details on the homogenization, staining, or flow cytometry assays, see Methods 2: flow cytometry in Supplementary Material.
+We obtained the mean fluorescent intensity for mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation.
+For further details on the homogenization, staining, or flow cytometry assays see _Methods 2: flow cytometry in Supplementary Material_.
+Sample sizes for DNA damage and lipid peroxidation were smaller than for the other variables because of a plater reader malfunction, however, data was missing randomly and we dealt with missing data in our analyses (see below).
 
 #### Statistical analyses
-We performed all the analyses using brm models from stan [@stan] on R version 4.4.0 [@R].
-We run a series of univariate and multivariate models to test the effects of early envirnoment on each variable separately plus the relationships between physiology and learning.
-All models consisted of four MCMC chains of 8000 iterations, with a warm up interval of 2000 iterations.
-
-Univariate models were used to test the effects of early environment on each of the variables recorded: mitochondrial density, mitochondrial potential, ROS, DNA damage, lipid peroxidation, and the number of errors as a measure of learning.
+We performed all analyses using the _brm_ package [@burkner2017brms] which fits Bayesian multilevel models with Stan [@stan] with R version 4.4.0 [@R].
+We ran a series of univariate and multivariate models to test the effects of early envinonment on each variable separately plus quantify the relationships between physiology and learning.
+All models consisted of four MCMC chains of 8000 iterations, with a warmup interval of 2000 iterations.
+
+Univariate models were used to test the effects of early environment on each of the variables recorded: mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and the number of errors as a measure of learning.
 We included in all the models the hormone treatment (CORT vs Control), incubation temperature (Cold vs Hot) and their interaction.
@@
 For all univariate models, clutch identity was included as a random factor.
-Mitochondrial related variables (density, potential, ROS, DNA damage, and lipid peroxidation) where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [@gelman2008scaling].
+Mitochondrial related variables where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [@gelman2008scaling].
 These variables were considered to follow a normal distribution.
 
-The number of errors was modeled as a function of trial, cort, temperature, and the triple interaction.
+Learning was modeled as a function of trial, cort, temperature, and their three-way interaction.
 We included in this model the trial within each level of lizard identity as a random slope.
@@
 Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments.
-Slope estimates <0 were considered as evidence of learning.
-pMCMC were used to test the hypothesis that slopes and slope contrasts are different from zero.
-We considered an effect statistically significant when pMCMC <0.05.
-
-To explore links between mitochondrial function and detection latency, we fitted a multivariate model on brm.
-We included in the model all the variables of interest and their interactions structured following a specific hypothesis (see @fig-sem).
-Individual measurements for learning performance were obtained by modelling errors as a function of trial.
-We included the trial within each level of lizard identity as a random slope, which was used to obtain the learning slopes for each individual.
-These learning slopes, as well as the rest of variables, were standardized as before (see above).
-We assumed a gaussian family for all the variables.
+Slope estimates less than 0 provide evidence of learning.
+_pMCMC_ was used to test the hypothesis that posterior distributions of slopes, and slope contrasts, were different from zero.
+We considered an effect statistically significant when _pMCMC_ <0.05.
+
+To explore direct and indirect links between mitochondrial function and learning, we fitted a multivariate structural equation model (SEM) using _brms_.
+The model was structured based on theoretical expectations shown in @fig-sem.
+Any missing data was imputed during model fitting using data augmentation [@noble2021planned], but this was largely restricted to DNA damage and lipid peroxidation.
+To obtain a measure of learning performance we extracted the posterior distribution of individual learning slopes (i.e., changes in error over time) for each lizard.
+This was achieved by including a random trial slope for each lizard.
+Learning slopes, as well as the rest of variables, were standardized as before (see above).
+We assumed a gaussian error distribution for all the variables in this model.
 Factors found as non-significant in the univariate models were excluded.
-Clutch identity remained a random factor.
-Direct, indirect, and total effects were derived from posterior estimates following a Structural Equations Modelling approach [@kline2005principles] (see Table S4).
-
-One of the plates used to analyze oxidative damage malfunctioned, compromising data quality.
-This data was excluded from the final analysis and, as a result, the sample size for DNA damage and lipid peroxidation is smaller than for the other variables.
-For the multivariate model, we imputed the missing values.
-The final sample sizes are presented below.
+Clutch identity remained a random factor in our SEM model.
+Direct, indirect, and total effects were derived from posterior estimates [@kline2005principles] (see table S4).
 
@@
 We started with 80 lizards - 20 per treatment - coming from `r total_clutches` clutches.
-However, due to natural mortality (n = 2), only 79 lizards were included in the learning analyses, and 78 in the mitochondrial analyses.
-Besides, due to malfunctioning of the flow cytometer, we ended up with 40 inividuals (10 per treatment) for DNA damage and lipid peroxidation.
-Final sample sizes per treatment and species are disclosed on @fig-learning and @fig-results_mit.
+However, due to natural mortality, only 79 lizards were included in the learning analyses, and 78 in the mitochondrial analyses.
+We provide final sample sizes in all our figures.
 
@@
 #| fig.cap: "Results for learning analyses.
-Panel A shows the predicted number of errors over trials.
-The lines represent the mean predicted number of errors for each trial, and the shaded areas indicate the standard deviation of the mean; both were obtained using the slope and intercept estimates from the posterior distributions.
-Panel B shows the distribution of the estimates of slopes per each treatment.
+(a) the predicted number of errors over trials.
+The lines represent the mean predicted number of errors for each trial, and the shaded areas indicate the standard error of the mean; both were obtained using the slope and intercept estimates from the posterior distributions.
+(b) shows the distribution of the estimates of slopes per each treatment.
 The x-axis represents the slope estimate, and in the y-axis are the density of the estimates.
-Points and bars represent the mean and standard deviation of the estimated slopes, respectively.
+Points and bars represent the mean and standard error of the estimated slopes, respectively.
 Dashed lines indicate value 0.
-The different colors in both panels indicate the different treatments."
+The different colors in both panels indicate the different treatments.
+Asterisks in (b) indicate significant differences from 0."
 source(here("R", "func.R"))
@@
   annotate("text", x = 0.045, y = 0.935, label = "A", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
-  annotate("text", x = 0.545, y = 0.935, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") 
+  annotate("text", x = 0.545, y = 0.935, label = "B", hjust = 1, vjust = 1, size = 7, fontface = "bold") +
+  annotate("text", x = 0.705, y = 0.89, label = "*", hjust = 1, vjust = 1, size = 7) +
+  annotate("text", x = 0.68, y = 0.72, label = "*", hjust = 1, vjust = 1, size = 7) +
+  annotate("text", x = 0.7, y = 0.55, label = "*", hjust = 1, vjust = 1, size = 7) +
+  annotate("text", x = 0.69, y = 0.38, label = "*", hjust = 1, vjust = 1, size = 7)
 #
@@
 
-We found that the lizards improved their choices over time, making less mistakes as the trials progressed (see @fig-learning and Table S1).
-However, the learning rate was not influenced by CORT, temperature, or their interaction (see contrasts in Table S2).
-
-Mitochondrial density, mitochondrial potential, ROS, DNA damage, and lipid peroxidation was not affected by prenatal conditions (see @fig-results_mit and Table S2).
-We did not see any significant effects of sex and age on mitochondrial physiology either (see Table S3 and Tables S5-S9).
+We found that the lizards improved their choices over time, making less mistakes as the trials progressed (see @fig-learning and table S1).
+However, the learning rate was not influenced by CORT, temperature, or their interaction (see contrasts in table S2).
+
+Mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation were also not affected by prenatal conditions (see @fig-results_mit and table S2).
+We did not see any significant effects of sex and age on mitochondrial physiology either (see table S3 and tables S5-S9).
 
@@
 #| label: fig-results_mit
-#| fig-cap: "Estimates of mitochondial density (A), mitochondrial potential (B), ROS (C), DNA damage (D), and lipid peroxidation (E) in the medial cortex as a function of the different prenatal conditions.
+#| fig-cap: "Estimates of mitochondrial density (a), metabolic capacity (b), ROS (c), DNA damage (d), and lipid peroxidation (e) in the medial cortex as a function of the different prenatal conditions.
+Note that for (d) and (e) these analyses do not account for missing data resulting from a flow cytometer malfunction that impacted one plate, and so, sample sizes are lower for univariate analyses.
 The x-axis represents the estimated values and in the y-axis is the density of the estimates.
-Points and bars represent the mean and standard deviation of the estimated values, respectively.
+Points and bars represent the mean and standard error of the estimated values, respectively.
 The different colors in both panels indicate the different treatments."
@@
 plot_mit_density <- plotting(MD, "Mit density")
-plot_mit_potential <- plotting(MP, "Mit potential")
+plot_mit_potential <- plotting(MP, "Metabolic capacity")
 plot_ros <- plotting(ROS, "ROS")
@@
   # Insert figure brain
-  draw_image(here("./Others/brain_fig.png"), x = 0.49, y = 0.3, width = 0.5, height = 0.5)
+  draw_image(here("./Others/brain_fig.png"), x = 0.53, y = 0.34, width = 0.45, height = 0.45)
 #
@@
 
-The multivariate model showed that mitochondrial physiology was not related to learning abilities (see @fig-sem and Table S4).
-We did not find significant relationships between ROS and oxidative damage or mitochondrial density and ROS (see @fig-sem and Table S4).
-However, we found that ROS production increased with mitochondrial potential (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], pMCMC `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`).
+Our SEM showed that mitochondrial physiology was not related to learning abilities (see @fig-sem and table S4).
+While we found that ROS production increased with metabolic capacity (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], _pMCMC_ `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`), ROS was not related to  oxidative damage (see @fig-sem and table S4).
 
@@
 #| label: fig-sem
-#| fig-cap: "Schematic representation of the multivariate model and the predicted coefficient for each variable.
-Arrows indicate the directionality of the estimate and the values show the mean predicted coefficient and the 95% CI around the mean."
+#| fig-cap: "Structural equation model testing hypothesized direct and indirect effects of physiology on learning.
+Arrows indicate the directionality of the estimate, and the values show the mean predicted coefficient and its 95% confidence interval."
 #| fig-name: "fig-sem"
@@
 ## Discussion
-
-We found that lizards improved their performance over time, indicating learning.
-However, this improvement was not influenced by prenatal conditions.
-Similarly, mitochondrial physiology was not affected by prenatal conditions, and we found no significant relationships between mitochondrial physiology and learning abilities.
-However, ROS production increased with mitochondrial potential, suggesting that mitochondrial potential may be a good predictor of ROS production in medial cortex of _L.
-delicata_ with potential implications for the balance between energy production and oxidative damage.
-
-#### Spatial learning in _Lampropholis delicata_ is robust to prenatal conditions  
-Lizards exhibited learning, but learning rates were not influenced by prenatal CORT or temperature.
-This contrasts with findings in other taxa, where incubation temperature and prenatal GC elevations affected learning and brain development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017].
-Our results suggest that _L.
-delicata_ buffers cognitive functions against early environmental inputs, consistent with previous findings showing that behavioural flexibility in this species is also unaffected by prenatal CORT or temperature [@recio2025early].
-However, since prenatal CORT and temperature can have region-specific effects on brain development [@coomber_independent_1997; @sakata_neural_2000], other cognitive domains might be more sensitive to prenatal conditions.
-Future research should explore whether different brain regions in _L.
-delicata_ vary in their resilience to prenatal inputs to gain a more comprehensive understanding of cognitive development in this species.
-
-#### Mitochondrial physiology is not affected by prenatal conditions  
+Lizards were able to learn a complex spatial task with marked improvements over time in their ability to choose the correct location.
+However, learning was not influenced by prenatal conditions or linked to mitochondrial function in the medial cortex, counter to what we predicted.
+Despite ROS increasing with greater metabolic capacity as expected, this did not result in a corresponding increase in oxidative damage in the medial cortex in _L.
+delicata_.
+We discuss our results in relation to existing work exploring the underlying mechanisms of learning and how early environments impact it below.
+
+#### Spatial learning is robust to prenatal conditions  
+Lizards exhibited clear evidence for learning but learning rates were not influenced by prenatal CORT or temperature as we predicted.
+This contrasts with findings in other taxa, where incubation temperature and prenatal GC’s have both been shown to affect learning and brain development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017].
+Our results suggest that spatial learning abilities in _L.
+delicata_ are buffered from early environmental insults, consistent with our previous findings showing that behavioural flexibility in this species is also unaffected by prenatal CORT or temperature [@recio2025early].
+_L.
+delicata_'s resilience may stem from the adaptive value of learning in lizards.
+For example, spatial learning in velvet geckos (_Amalosia lesueurii_) enhances survival in the wild [@dayananda_incubation_2017].
+Future studies should investigate the mechanisms underlying this robustness and the role of cognition in _L.
+delicata_'s fitness.
+
+#### Brain metabolic physiology is not affected by prenatal conditions   
 Contrary to our predictions, prenatal temperature and CORT did not significantly influence mitochondrial physiology in the medial cortex of _L.
 delicata_.
-Studies in other taxa have been shown that incubation at high temperatures could decrease energy production and oxidative stress [@treidel2016temperature; @zavorka_climate_2021] while, GCs elevation or maternal stress is related to lower mitochondrial efficiency and higher oxidative stress [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
-
+Studies in other taxa have shown that incubation at high temperatures decrease energy production and oxidative stress [@treidel2016temperature; @zavorka_climate_2021], while elevated GCs or maternal stress is related to lower mitochondrial efficiency and higher oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
 In _L.
-delicata_, previous research has demonstrated that high incubation temperatures decreased mitochondrial efficiency in the liver [@crino2024eggs].
-Our study shows, however, that mitochondrial physiology in _L.
-delicata_'s brain is robust to both incubation temeprature and prenatal CORT.
-Nevertheless, brain effects of prenatal conditions can be region-dependent, with some areas having higher sensitivity to prenatal stressors than others [@coomber_independent_1997; @sakata_neural_2000].
-As such, further investigations are necessary to determine whether mitochondrial responses to prenatal conditions vary between different brain regions in _L.
-delicata_ and their potential effects on behaviour.
-
-#### Mitochondrial physiology is not related to learning abilities
-Mitochondria play a critical role in cognitive abilities by synthesizing energy [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018], and through effects on cell senescence and death caused by oxidative damage [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022].
-For example, donut-shaped mitochondrial density — associated with higher ROS production — on presynaptic neurons impairs visuospatial working-memory in rhesus monkeys (_Macaca mulatta_) [@hara_presynaptic_2014].
-However, we found no significant relationships between mitochondrial physiology and learning rate in _L.
+delicata_, high incubation temperatures are known to decrease mitochondrial efficiency in the liver [@crino2024eggs].
+However, our study shows that mitochondrial physiology in _L.
+delicata_’s brain is robust to both incubation temperature and prenatal CORT.
+This could reflect differential tissue responses to early-life conditions, highlighting the importance of studying these effects across multiple tissues.
+
+#### Variation in brain metabolic physiology is not related to learning abilities
+Mitochondria play a critical role for cognitive function by synthesizing energy [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018], and by affecting the rate of cell senescence and death caused by oxidative damage [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022].
+For example, donut-shaped mitochondria are associated with higher ROS production in presynaptic neurons and this mechanism has been shown to impair visuospatial working-memory in rhesus monkeys (_Macaca mulatta_) [@hara_presynaptic_2014].
+
+Our findings contrast with these studies and suggest that energy limitations do not impact the process of spatial learning in _L.
 delicata_.
-
-Our findings suggest that energy production is not a limiting factor for spatial learning in _L.
-delicata_.
-The lack of association between mitochondrial density, mitochondrial potential, and learning abilities may reflect the species' ability to maintain energy production above the threshold required to prevent cognitive dysfunction.
-Alternatively, _L.
-delicata_ can sustain cognitive function despite fluctuations in energy availability.
-
-However, we found a significant positive relationship between mitochondrial potential and ROS in the medial cortex.
-This aligns with the role of mitochondrial membrane potential in driving ATP synthesis and electron transport, processes that inherently generate ROS as byproducts [@rice2002brain].
+The lack of association between mitochondrial density, mitochondrial capacity, and learning abilities may be the result of lizards being able to maintain energy production above the threshold required to prevent cognitive dysfunction.
+The generally low energetic demands of ectothermic organisms compared to endotherms may make this possible and explain why we found no effect when they seem to be common in mammals [@hara_presynaptic_2014; @cao2014ampk].
+
+Interestingly, despite finding a significant positive relationship between metabolic capacity and ROS in the medial cortex, the heightened ROS production did not lead to greater oxidative stress.
+The strong relationship between ROS and mitochondrial capacity aligns with the role of mitochondrial membrane potential in driving ATP synthesis and electron transport, processes that inherently generate ROS as byproducts [@rice2002brain].
 Normally, moderate ROS levels serve essential signaling functions and are controlled by antioxidants [@rice2002brain; @terman2006oxidative].
 However, excessive ROS production can exceed antioxidant action, leading to oxidative stress and cellular dysfunction [@rice2002brain; @terman2006oxidative].
-In our study, we did not find any significant relationship between ROS and oxidative damage, suggesing that antioxidant action was sufficient to counteract ROS effects.
-However, higher energetic demands in the medial cortex may overcome antioxidant capactities and elevate oxidative stress.
-
-Additionally, DNA damage or lipid peroxidation in the medial cortex were not linked to spatial learning.
-This suggests that oxidative damage levels were not high enough to impair cognitive function or that compensatory mechanisms counteracted any negative effects.
-Nevertheless, since oxidative damage accumulate over time [@terman2006oxidative], the relationships between oxidative damage and cognitive dysfunction could become more pronounced in older individuals [@hara_presynaptic_2014].
+The lack of oxidative damage from increased ROS strongly suggests that antioxidants could deal with these levels of ROS.
+As such, DNA damage and lipid peroxidation in the medial cortex were not linked to spatial learning in _L.
+delicata_ possibly because neurons were effectively buffered from damaging free radicals.
+Nevertheless, since oxidative damage can accumulate over time [@terman2006oxidative], the relationships between oxidative damage and cognitive dysfunction could become more pronounced in older individuals [@hara_presynaptic_2014].
 Future research should investigate the long-term consequences of mitochondrial activity on brain health and the potential mechanisms that sustain cognitive resilience in _L.
@@
 Additionally, there was no relationship between mitochondrial function and learning abilities in a spatial task.
-However, we found that ROS production increased with mitochondrial potential, suggesting that links between energy production, mitochondrial respiration, and oxidative stress may become more pronounced under stressful conditions or with aging, potentially influencing cognitive function over time.
+However, we found that ROS production increased with metabolic capacity, suggesting that links between mitochondrial respiration and oxidative stress may become more pronounced under higher energetic demands or with aging, potentially influencing cognitive function over time.
 
@@
 We used these fluorescent probes as indicators of mitochondrial density, mitochondrial membrane potential, and superoxide (ROS) production, respectively.
-We then added 5µL of #µM Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
+We then added 5µL of 10 µg/mL Hoechst 33342 Nuclear Viability Dye to each sample, which we used to distinguish live, viable, intact cells from cellular debris.
 We then loaded 6 wells with 50µL of homogenate taken from each pooled homogenate suspension (12 wells total), which were to be negative and single-color controls.
-One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5µL of one of 5µM MitoTracker Deep Red FM, 2.5µM MitoTracker Orange CMTMRos, 50µM MitoSOX Red, or #µM Hoechst 33342 Nuclear Viability Dye.
+One well was left unstained as a negative control, one was stained with all the probes to be a positive control, and the remaining four wells were treated with 5µL of one of 5µM MitoTracker Deep Red FM, 2.5µM MitoTracker Orange CMTMRos, 50µM MitoSOX Red, or 10 µg/mL Hoechst 33342 Nuclear Viability Dye.
 Any remaining pooled homogenate was fixed and frozen as previously described.
@@
 
-The aliquots destined to the analysis of oxidative damage were stained with 20µL (10µL for controls) of #µM Hoechst 33342 Nuclear Viability Dye and 20µL of 100µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32°C for 20 minutes.
+The aliquots destined to the analysis of oxidative damage were stained with 20µL (10µL for controls) of 10 µg/mL Hoechst 33342 Nuclear Viability Dye and 20µL of 100µM BODIPY 665/676 Lipid Peroxidation Sensor and incubated at 32°C for 20 minutes.
 Following staining, we washed cells to prevent further binding of unbound fluorescent probes, then resuspended the pellet and continue with cryopreservation.
@@
 We added 20µL of 70µM 8-OHdG Polyclonal Antibody to each sample, and we left the homogenate overnight (~12 hours) for the antibody to bind to 8-OHdG, a marker of oxidative damage on DNA.
-The following day we counterstained the cells with 20µL of #µM (100µg/mL) H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32°C for 20 minutes.
+The following day we counterstained the cells with 20µL of 100 µg/mL H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 at 32°C for 20 minutes.
 After the cells had been tagged with 8-OHdG antibodies and counterstained, we washed the cells once more and resuspended the pellet in 400µL of 1X PBS.
@@
 For analysis, we exported summary statistics of only the area of the fluorescent signal.
-Although we exported robust confidence values for checking repeatability between replicate samples, we used the mean and standard deviation of geometric means among replicates for each individual as our main response variables in analyses.
+Although we exported robust confidence values for checking repeatability between replicate samples, we used the geometric means for each individual as our main response variables in analyses.
 
@@
 
-Fig S1. Raw data of the learning task. Each point represents the mean number of errors per day. The lines represent the modeled data.
-
-```{r, results='asis', echo=FALSE}
-cat("\\newpage")
-```
-
-Table S1. Estimates of the learning slopes per treatment, the 95% CI, and pMCMC values testing the hypothesis that slope is different from 0. In bold, the values that are significant at pMCMC < 0.05.
+Figure S1. Raw data of the learning task. Each point represents the mean number of errors per day. The lines represent the modeled data.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```
+
+_Table S1.
+Estimates of the learning slopes per treatment, the 95% CI, and _pMCMC_ values testing the hypothesis that slope is different from 0.
+In bold, the values that are significant at _pMCMC_ < 0.05._
 ```{r, tbl-learning}
@@
 
-Table S2.
-Estimate of the contrasts between treatments for all the variables analysed, and pMCMC values testing the hypothesis that contrast is different from 0.
-In bold, the values that are significant at pMCMC < 0.05.
+_Table S2.
+Estimate of the contrasts between treatments for all the variables analysed, and _pMCMC_ values testing the hypothesis that contrast is different from 0.
+In bold, the values that are significant at _pMCMC_ < 0.05._
 ```{r tbl-contrasts}
@@
       Variable == "MD" ~ "Mit density",
-      Variable == "MP" ~ "Mit potential",
+      Variable == "MP" ~ "Metabolic capacity",
       Variable == "ROS" ~ "ROS",
@@
 Contrasts were done by:  
-_Temperature_: β~Hot~ - β~Cold~
-_CORT_: β~CORT~ - β~Control~
-_Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
+- _Temperature_: β~Hot~ - β~Cold~  
+- _CORT_: β~CORT~ - β~Control~  
+- _Interaction_: (β~Control-Hot~ - β~CORT-Hot~) - (β~Control-Cold~ - β~CORT-Cold~)
 
@@
 
-Table S3. Estimated direct, indirect, and total coefficients from the multivariate model
+_Table S3. Estimated direct, indirect, and total coefficients from the multivariate model_
 ```{r, table_sem_results}
@@
         predictor_modulator == "mit_density" ~ "Mitochondrial density",
-        predictor_modulator == "mit_potential" ~ "Mitochondrial potential",
+        predictor_modulator == "mit_potential" ~ "Metabolic capacity",
         predictor_modulator == "DNAdamage" ~ "DNA damage",
@@
          Predictor = factor(Predictor, levels = c("Mitochondrial density",
-                                                "Mitochondrial potential",
+                                                "Metabolic capacity",
                                                 "ROS",
@@
 
-Fig S2. Posterior predictive checks for the model of Mitochondrial Density.
+Figure S2. Posterior predictive checks for the model of Mitochondrial Density.
 Formula: mit_density ~ cort * temp + (1|clutch)
@@
 #| label: plotmod_potential
-#| caption: "Posterior predictive checks for the model of Mitochondrial Potential."
+#| caption: "Posterior predictive checks for the model of metabolic capacity."
 #
@@
 
-Fig S3. Posterior predictive checks for the model of Mitochondrial Potential.
+Figure S3. Posterior predictive checks for the model of metabolic capacity.
 Formula: mit_potential ~ cort * temp + (1|clutch)
@@
 
-Fig S4. Posterior predictive checks for the model of ROS.
+Figure S4. Posterior predictive checks for the model of ROS.
 Formula: ROS ~ cort * temp + (1|clutch)
@@
 
-Fig S5. Posterior predictive checks for the model of DNA Damage.
+Figure S5. Posterior predictive checks for the model of DNA Damage.
 Formula: DNAdamage ~ cort * temp + (1|clutch)
@@
 
-Fig S6. Posterior predictive checks for the model of lipid peroxidation.
+Figure S6. Posterior predictive checks for the model of lipid peroxidation.
 Formula: peroxidation ~ cort * temp + (1|clutch)
@@
 
-Fig S7. Posterior predictive checks for the model of spatial learning.
+Figure S7. Posterior predictive checks for the model of spatial learning.
 Formula: errors ~ day * cort * temp + (1 + day|lizard_id) + (1|clutch)
@@
 
-Table S4. Preliminary results of the models testing for Mitochondrial Density.
+_Table S4. Preliminary results of the models testing for Mitochondrial Density._
 ```{r, results_preliminary_mitdensity}
@@
 
-Table S5.  Preliminary results of the models testing for Mitochondrial Potential.
+Table S5.  Preliminary results of the models testing for Metabolic capacity.
 ```{r, results_preliminary_potential}
-#| tbl-cap: "Preliminary results of the models testing for Mitochondrial Potential"
+#| tbl-cap: "Preliminary results of the models testing for Metabolic capacity"
 #| label: results_preliminary_potential

### 2025-03-23 .vdoc.r

@@ -1,1305 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: data_processing
-# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
-refit <- FALSE
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-df_damage <- clean_df %>%
-  filter(DNAdamage != "NA")
-n_damage <- list()
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
[…]

### 2025-03-23 R/func.r

@@ -255,7 +255,7 @@ post_values <- function(df, fac){
 # Function to create the plots for mit variables
 #' @title plotting
 #' @param df to select the df (only posteriors)
-#' @param lab to select the appropriate label for axix Y
+#' @param lab to select the appropriate label for axis Y
 plotting <- function(df, lab){
   if(lab %in% c("DNA damage", "Lipid peroxidation")){
     n_CORTCold <- paste0("CORT-Cold (n=", n_damage$CORT_Cold, ")")
@@ -288,12 +288,13 @@ plotting <- function(df, lab){
   plot <- ggplot(db_violin, aes(x = Treatment, y = Estimate, fill = Treatment)) +
     geom_flat_violin(alpha = 0.5) +
     scale_fill_manual(values = set_names(
-      c("#fa927d", "#b50101", "#68bde1", "#00008B"),
+      c("#00008B", "#b50101", "#68bde1", "#fa927d"),
       unique(db_violin$Treatment))) +
     geom_point(data = db_bars, aes(y = Mean, x = Treatment), position = position_dodge(width = 0.75), color = "black", fill = "black", size = 3) +
     geom_segment(data = db_bars, aes(y = Mean - SD, yend = Mean + SD, x = Treatment, xend = Treatment), size = 1.5, color = "black") +
     ylim(min(db_violin$Estimate), max(db_violin$Estimate)) +
     coord_flip() +
+    guides(fill = guide_legend(reverse = TRUE)) +
     theme_classic() +
     labs(y = lab, x = "Treatments") +
     theme(plot.margin = margin(3, 3, 3, 3, "mm"),

### 2025-04-04 ms.qmd
updates

@@
 ### Acknowledgements
+
 We thank Dr.
@@
 ```
+
+#### Brain validation
+
+To ensure that neurons were not unintentionally lost during homogenization, we performed a small pilot study where we euthanized four other lizards coming from the same breeding colony and prepeared the homogenates from the same brain regions as in the experiment.
+The homogenates here were then dyed with a specific neuronal marker as described in Farrow et al.
+(2021) and Storks et al.
+(2023).
+We employed flow cytometry and **fluorescent optical microscopy** to check for the presence of neurons in the homogenates.
+
+Euthanasia and homogenization for each of brain region followed the procedures outlined above (see Methods: Flow cytometry).
+Before staining the samples, cells were also fixed and permeabilized as described previously (see Methods: Flow cytometry).
+Following permeabilization, we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of a 1:100 NeuN + Alexa488 fluorescent conjugate solution to dye the neuronal nuclei (Farrow et al.
+2021; Storks et al.
+2023).
+The samples were incubated at +4 ºC overnight.
+
+The following day we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of PBS.
+Samples were split in two: 50 µL from each was reserved for examination under a **MICROSCOPE**, while the remaining 50 µL from each sample was pooled into a single tube for flow cytometry analysis.
+
+```{r, cell_size}
+#| label: cell_size
+# Getting the size of the cells in the experiment
+mit_df <- read.csv(here("./data/mit_data.csv"))
+euth <- read.csv(here("./data/Euthanasia.csv"))
+#
+data_cells <- merge(mit_df, euth, by = c("plate", "tube.no")) %>%
+  filter(type != "control", ch == "fsc") %>% 
+  dplyr::select(lizard_id, sex, ch, geo.mean) %>%
+  group_by(lizard_id, ch) %>%
+  summarize(geo.mean = mean(geo.mean, na.rm = TRUE),
+            .groups = 'drop') %>%
+  pivot_wider(
+    names_from = ch,
+    values_from = geo.mean
+  ) %>%
+data.frame()
+#
+mean_cells <- format_dec(mean(data_cells$fsc), 3)
+quant2_5_cells <- format_dec(quantile(data_cells$fsc, 0.025), 3)
+quant97_5_cells <- format_dec(quantile(data_cells$fsc, 0.975), 3)
+```
+
+Samples examined under the microscope showed a clear presence of neuronal nuclei in the homogenates (Figure S13).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r mean_cells`, 95% CI = [`r quant2_5_cells`, `r quant97_5_cells`], n = 80; size neurons = ).
+
+```{r, fig-S8}
+#| label: fig-S13
+#| fig-cap: "Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells."
+#
+knitr::include_graphics("./output/figures/microscope/neurons_black_background.jpg")
+```
+
+Figure S8. Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells.
+
+**REFERENCES**
+
+Farrow LF, Andronicos NM, McDonald PG, Hamlin AS. 2021. Quantitative determination of neuronal size and density using flow cytometry. _Journal of Neuroscience Methods_, **352**, 109081.
+
+Storks L, Powell BJ, Leal M.
+2023.
+Peeking inside the lizard brain: neuron numbers in Anolis and its implications for cognitive performance and vertebrate brain evolution.
+_Integrative and Comparative Biology_, **63(1)**, 223-237.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-04-04 .vdoc.r

@@ -707,6 +707,7 @@ knitr::include_graphics("./output/figures/text/SEM.png")
 #
 #
 #
+#
 cat("\\newpage")
 #
 #
@@ -1291,6 +1292,58 @@ flextable(sum_m_learn_prel)
 cat("\\newpage")
 #
 #
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: cell_size
+# Getting the size of the cells in the experiment
+mit_df <- read.csv(here("./data/mit_data.csv"))
+euth <- read.csv(here("./data/Euthanasia.csv"))
+#
+data_cells <- merge(mit_df, euth, by = c("plate", "tube.no")) %>%
+  filter(type != "control", ch == "fsc") %>% 
+  dplyr::select(lizard_id, sex, ch, geo.mean) %>%
+  group_by(lizard_id, ch) %>%
+  summarize(geo.mean = mean(geo.mean, na.rm = TRUE),
+            .groups = 'drop') %>%
+  pivot_wider(
+    names_from = ch,
+    values_from = geo.mean
+  ) %>%
+data.frame()
+#
+mean_cells <- format_dec(mean(data_cells$fsc), 3)
+quant2_5_cells <- format_dec(quantile(data_cells$fsc, 0.025), 3)
+quant97_5_cells <- format_dec(quantile(data_cells$fsc, 0.975), 3)
+#
+#
+#
+#
+#
+#| label: fig-S13
+#| fig-cap: "Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells."
+#
+knitr::include_graphics("./output/figures/microscope/neurons_black_background.jpg")
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+cat("\\newpage")
 #
 #
 #

### 2025-04-07 ms.qmd
updates

@@
 #
-mean_cells <- format_dec(mean(data_cells$fsc), 3)
-quant2_5_cells <- format_dec(quantile(data_cells$fsc, 0.025), 3)
-quant97_5_cells <- format_dec(quantile(data_cells$fsc, 0.975), 3)
+mean_cells <- format_dec(log(mean(data_cells$fsc)), 3)
+quant2_5_cells <- format_dec(log(quantile(data_cells$fsc, 0.025)), 3)
+quant97_5_cells <- format_dec(log(quantile(data_cells$fsc, 0.975)), 3)
+#
+# Getting the size of the neurons in the validation
+data_neurons <- read.csv(here("./data/neuron_size.csv")) %>%
+  filter(X == "Medial Cortex_Tube_002_003.fcs")
+neuron_size <- log(as.numeric(data_neurons$cells.Neurons...Geometric.Mean..FSC.A.))
 ```
@@
 Samples examined under the microscope showed a clear presence of neuronal nuclei in the homogenates (Figure S13).
-Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r mean_cells`, 95% CI = [`r quant2_5_cells`, `r quant97_5_cells`], n = 80; size neurons = ).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were within the size range of the cells employed in the experiment (mean size cells experiment = `r mean_cells`, 95% CI = [`r quant2_5_cells`, `r quant97_5_cells`], n = 80; size neurons = `r neuron_size`).
 

### 2025-04-07 .vdoc.r

@@ -1,1350 +1,0 @@
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#| label: setup
-pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
-#
-#
-#
-#| label: data_processing
-# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
-refit <- FALSE
-source(here("R", "data_process.R"))
-#
-#
-#
-#| label: data_metrics
-# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
-data <- read.csv(here("./data/Spatial_learn.csv"))
-#
-# Total number of clutches
-total_clutches <- data %>% 
-  distinct(clutch) %>% 
-  count() %>% 
-  pull(n)
-# Calculate total number of individuals
-total_individuals <- data %>% 
-  distinct(lizard_id) %>% 
-  count() %>% 
-  pull(n)
-# Calculate number of individuals per clutch
-individuals_per_clutch <- data %>%
-  distinct(lizard_id, clutch) %>%
-  group_by(clutch) %>%
-  summarize(individual_count = n())
-#
-#
-#
-#| label: sampleSize
-# List with the sample sizes from the main database.
-source(here("R", "func.R"))
-#
-hormone <- c("CORT", "Control")
-temperature <- c("Cold", "Hot")
-#
-n_list <- list()
-#
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
-    list_name <- paste0(hormone[k], "_", temperature[l])
-    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
-  }
-}
-#
-df_damage <- clean_df %>%
-  filter(DNAdamage != "NA")
-n_damage <- list()
-for(k in 1:length(hormone)){
-  for(l in 1:length(temperature)){
[…]

### 2025-04-11 ms.qmd
updates

--- 2025-03-23 ms.qmd
@@
 
-Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3^, Daniel W.A. Noble^1^
+Pablo Recio^1^,$\ddagger$ , Dalton C. Leibold^1^, Ondi L. Crino ^1,2^, Christopher R. Friesen^3,4^, Daniel W.A. Noble^1^
 
 ^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2601, Australia  
-^2^ Flinder's University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
+^2^ Flinders University, College of Science and Engineering, Bedford Park, SA 5042, Australia  
 ^3^ University of Wollongong, Wollongong, NSW 2500, Australia  
+^4^ Environmental Futures University of Wollongong, Wollongong, NSW 2500, Australia
 $\ddagger$ Corresponding author: pablo.reciosantiago@anu.edu.au
-
-ORCID:
 
@@
 No significant relationship was found between mitochondrial function and cognitive performance, contrary to predictions.
-Additionally, increased metabolic capacity correlated with higher ROS production but did not affect oxidative damage, possibly due to protective mechanisms.
+Increased metabolic capacity correlated with higher ROS production but did not affect oxidative damage, possibly due to protective mechanisms.
 These findings highlight the physiological and cognitive resilience of _L.
@@
 
-Learning - the acquisition and consolidation of new information [@dukas_evolutionary_2004] - enables animals to create new associations between events, which can be essential for coping with environmental change [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
-However, the capacity to form these associations varies among individuals, potentially affecting their responses to environmental challenges [@ward2016ecological; @welklin2024spatial].
-Learning faster may imply exploiting novel resources better or responding to new threats more efficiently [@ward2016ecological].
-Consequently, individual differences in learning abilities can have profound effects on fitness, ultimately influencing population growth rates and stability [@ward2016ecological; @welklin2024spatial].
-Understanding the mechanisms that drive these differences is therefore crucial for predicting how populations will respond to novel conditions.
+Learning - the acquisition and consolidation of new information - enables animals to create new associations between events, which can be essential for coping with environmental change [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013].
+However, the capacity to form new associations varies among individuals, potentially affecting their responses to environmental challenges [@ward2016ecological; @welklin2024spatial].
+Learning faster may imply better exploitation of resources or responding more efficiently to novel threats [@ward2016ecological].
+Consequently, individual differences in learning abilities can profoundly affect fitness, ultimately influencing population growth rates and stability [@ward2016ecological; @welklin2024spatial].
+Therefore,understanding the mechanisms that drive these differences is crucial for predicting how populations will respond to novel conditions.
 
@@
 Adverse developmental environments are particularly influential, as the brain is highly sensitive to environmental inputs during early life [@zhu_prenatal_2004].
-In fact, early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
+For example, early experiences can alter neurotransmitter production [@amani2021perinatal], gene expression [@zhou2020effects], or brain structure [@amiel_effects_2017], with lasting effects on cognition.
 Among the physiological mechanisms underlying learning, mitochondrial activity is considered a key factor [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018].
@@
 Studies on mammals have shown the pervasive effects of mitochondrial physiology on cognitive abilities [@zhu_prenatal_2004; @hara_presynaptic_2014; @cao2014ampk].
-However, the extent to which these effects are generalizable to other taxa remains largely unexplored.
+However, the extent to which these effects are generalisable to other taxa remains largely unexplored.
 
@@
 Maternal stress can also influence how mitochondria operate in offspring [@zhu_prenatal_2004].
-Stressful situations experienced by mothers can elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can be transmitted to the developing embryos [@uller_sex-specific_2009] altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].
+Stressful situations experienced by mothers can elevate glucocorticoids (GCs) [@sapolsky_how_2000], which can affect developing embryos [@uller_sex-specific_2009] altering mitochondrial physiology through transgenerational effects [@picard_mitochondria_2014].
 For instance, maternal stress has been shown to contribute significantly to oxidative stress in the brain of rats (_Rattus norvegicus_) with impacts on spatial learning abilities [@zhu_prenatal_2004; @haussmann_embryonic_2012; @cao2014ampk].
@@
 Temperature is also a significant source of maternal stress in ectotherms.
-Thermal environments outside of the optimal range can elevate GCs in mothers, which can be passed to the offspring [see @Crino_2023].
-Additionally, temperature can also directly influence offspring development, particularly during early life [@noble_developmental_2018; @crino2024eggs].
-The prenatal thermal environment plays a crucial role in shaping mitochondrial function, affecting energy metabolism and oxidative stress [@stier2022experimental; @crino2024eggs].
+Thermal environments outside the optimal range can elevate GCs in mothers, which can be passed to the offspring [see @Crino_2023].
+Temperature can also directly influence offspring development, particularly during early life [@noble_developmental_2018; @crino2024eggs].
+The prenatal thermal environment is crucial in shaping mitochondrial function, affecting energy metabolism and oxidative stress [@stier2022experimental; @crino2024eggs].
 Thus, the combined effects of prenatal GCs and temperature may profoundly influence mitochondrial function, with important consequences for brain development and cognition.
@@
 Here, we examined how prenatal temperature and corticosterone (CORT) — the primary GC in reptiles — affect brain mitochondrial physiology and spatial learning in the delicate skink (_Lampropholis delicata_).
-We hypothesized that prenatal CORT and temperature would influence learning abilities through impacts on brain mitochondrial activity.
+We hypothesized that prenatal CORT and temperature would influence learning abilities by impacting brain mitochondrial activity.
 Specifically, we predicted that prenatal CORT would decrease energy production while increasing oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs], especially if CORT can alter cellular components increasing ROS production without enhancing ATP synthesis or making them more vulnerable to oxidative damage.
-Similarly, high temperatures would decrease mitochondrial efficiency [@zavorka_climate_2021; @crino2024eggs], but with the potential to reduce oxidative stress [@treidel2016temperature].
+Similarly, we predicted that high temperatures would decrease mitochondrial efficiency [@zavorka_climate_2021; @crino2024eggs], but would reduce oxidative stress [@treidel2016temperature].
 We further predicted that the combined effects of CORT and temperature would lead to complex interactions, with both factors negatively impacting mitochondrial efficiency but having opposite effects on oxidative stress.
-These effects would lead to differences in learning abilities, that could be affected by the balance between energy production and oxidative stress [@du_dynamic_2009; @picard_energetic_2018; @alexandrov_neuronal_2022].
-By examining these interactions, we aim to clarify how prenatal environmental factors shape learning abilities through mitochondrial function, shedding light on the role of early-life conditions in cognitive development and associated mechanismsm.
+These effects would lead to differences in learning abilities, which could be affected by the balance between energy production and oxidative stress [@du_dynamic_2009; @picard_energetic_2018; @alexandrov_neuronal_2022].
+By examining these interactions, we aim to clarify how prenatal environmental factors shape learning abilities through mitochondrial function, clarifying the mechanisms that mediate the role of early-life conditions on cognitive development.
 
@@
 Lizards came from a breeding colony established in the laboratory in 2019 from wild populations in Sydney, Australia.
-The colony consisted of 270 adults living in groups of two males and four females.
+The colony consisted of 270 adults in groups of two males and four females.
 Eggs were collected from these groups between November 2023 to January 2024.
-After collecting the eggs, we treated them with CORT or a vehicle control and incubated them under two different temperature regimes (see below).
+After collecting the eggs, we treated them with CORT or vehicle control and incubated them under two different temperature regimes (see below).
 Clutch and egg identity were assigned immediately after egg collection, and the eggs were incubated in individual cups until hatching.
@@
 The spatial learning task involved training lizards to navigate a 6-arm maze to reach an exit connected to a transport box that allowed us to return the lizards to their enclosure without further contact.
-In each trial, lizards were placed by hand in the center of maze and left to acclimatise for two minutes.
+In each trial, lizards were placed by hand in the center of the maze and left to acclimatise for two minutes.
 During acclimatisation, the central area was surrounded by a yellow device mounted on a pulley system.
@@
 If the lizard did not immediately choose an arm, it was gently prodded with a brush at the end of the tail.
-Once the lizard made a choice, the brush was used to encourage movement but without guiding it towards any specific arm.
+Once the lizard made a choice, the brush encouraged movement but without guiding it towards any specific arm.
 If the lizard did not choose correctly after 20 errors — at which point it typically ceased responding — the lizard was gently guided to the correct arm.
@@
 Additionally, the maze was cleaned with 70% ethanol between trials to avoid the influence of chemical cues.
-To control for potential side biases, the correct arm was randomly assigned to one of the six arms for each trial.
-We employed four different maze orientations counterbalancing the number of lizards assigned to each orientation across treatments.
+The correct arm was randomly assigned to one of the six arms for each trial to control for potential side biases.
+We employed four maze orientations to counterbalance the number of lizards assigned to each orientation across treatments.
 
 The task was repeated once daily for 40 consecutive days, and the number of errors made by each lizard was recorded.
-An error was defined as the lizard inserting its head into one of the incorrect arms, choosing the same arm two or more times consecutively was counted as independent errors.
+An error was defined as the lizard inserting its head into one of the incorrect arms with each incorrect choice, including consecutive choices of the same arm, counted as independent errors.
 
 #### Brain mitochondrial function
-Immediately following the completion of the tests, we quantified mitochondrial physiology in the brain of lizards.
-We euthanized lizards via peritoneal injection 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic), followed by decapitation.
-Before decapitation we evaluated lizard righting responses and the pinching reflexes in one of the front limbs.
+Immediately after completing the tests, we quantified mitochondrial physiology in the brains of lizards.
+We euthanized lizards via intraperitoneal injection of 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic), followed by decapitation.
+Before decapitation, we evaluated lizard righting responses and the pinching reflexes in one of the front limbs.
 
@@
 
-The aliquots reserved for oxidative damage assays were stained with 20 µL of 10 µg/mL Hoechst 33342 Nuclear Viability Dye and 20 µL of 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
+The aliquots reserved for oxidative damage assays were stained with 10 µg/mL Hoechst 33342 Nuclear Viability Dye and 100 µM BODIPY 665/676 Lipid Peroxidation Sensor before cryopreservation.
 These dyes were used to measure cell viability and lipid peroxidation, respectively.
@@
 On the day of the oxidative damage assays, the samples were thawed, and DMSO was removed.
-Then, cell membranes were permeabilized in 200 µL of 1X PBS containing 5mM (6.14 mg/mL) digitonin.
-Following permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody - a marker of DNA damage - and allowed them to incubate overnight (~12 hours).
+Then, cell membranes were permeabilized in 200 µL of 1X PBS containing 20µM digitonin.
+Following permeabilization we stained the samples with 20µL of 70µM 8-OHdG Polyclonal Antibody - a marker of DNA damage - and allowed them to incubate overnight at 4 ºC.
 The following day we counterstained the cells with 20µL of 100 µg/mL H+G Goat Anti-Rabbit Conjugate Antibody with Alexa-Fluor 488 and analyzed the samples using the flow cytometer.
-Oxidative damage assays were performed within 6 months of the initial analysis of fresh samples.
-
-Flow cytometry was performed on a Becton Dickson LSRFortessa X-20 flow cytometer using the default wavelength filters.
+Oxidative damage assays were performed within 6 months of the initial preparation of fresh samples.
+
+Flow cytometry was performed on a 5-lasers flow cytometer (Becton Dickson LSRFortessa X-20) using the default wavelength filters and a high-throughput plate reader.
 Detectors and voltage settings for each assay were determined during pilot trials and remained consistent throughout the experiment.
@@
 We obtained the mean fluorescent intensity for mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation.
-For further details on the homogenization, staining, or flow cytometry assays see _Methods 2: flow cytometry in Supplementary Material_.
-Sample sizes for DNA damage and lipid peroxidation were smaller than for the other variables because of a plater reader malfunction, however, data was missing randomly and we dealt with missing data in our analyses (see below).
+For further details on the homogenization, staining, or flow cytometry assays, see _Methods 2: flow cytometry in Supplementary Material_.
+Sample sizes for DNA damage and lipid peroxidation were smaller than for the other variables because of a plater reader malfunction; however, data was missing randomly and we dealt with missing data in our analyses (see below).
+We validated that our homogenates contained neurons in a pilot study that used dyes specifically targeting neuronal nuclei (See _Brain validation in Supplementary Material_).
+This pilot study also ensured that our gating strategy identified these neurons using Flow cytometry.
 
 #### Statistical analyses
-We performed all analyses using the _brm_ package [@burkner2017brms] which fits Bayesian multilevel models with Stan [@stan] with R version 4.4.0 [@R].
+We performed all analyses using the _brm_ package [@burkner2017brms], which fits Bayesian multilevel models with _Stan_ [@stan] with R version 4.4.0 [@R].
 We ran a series of univariate and multivariate models to test the effects of early envinonment on each variable separately plus quantify the relationships between physiology and learning.
@@
 
-Univariate models were used to test the effects of early environment on each of the variables recorded: mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and the number of errors as a measure of learning.
-We included in all the models the hormone treatment (CORT vs Control), incubation temperature (Cold vs Hot) and their interaction.
+Univariate models were used to test the effects of early environment on each variable recorded: mitochondrial density, metabolic capacity, ROS, DNA damage, lipid peroxidation, and the number of errors as a measure of learning.
+In all the models, we included hormone treatment (CORT vs Control), incubation temperature (Cold vs Hot), and their interaction.
 Sex and age were included in preliminary models and excluded from the final models when they were not significant.
 For all univariate models, clutch identity was included as a random factor.
-Mitochondrial related variables where log-transformed when necessary and all were mean-centered and standardized by dividing them by two times the standard deviation [@gelman2008scaling].
+Mitochondrial-related variables where log-transformed when necessary, and all were mean-centered and standardized by dividing them by two times the standard deviation [@gelman2008scaling].
 These variables were considered to follow a normal distribution.
 
-Learning was modeled as a function of trial, cort, temperature, and their three-way interaction.
-We included in this model the trial within each level of lizard identity as a random slope.
+Learning was modeled as a function of trial, CORT, temperature, and their three-way interaction.
+In this model, we included the trial within each level of lizard identity as a random slope.
 The error structure was modelled using a negative binomial distribution with a logit link function [negbinomial(link = "log")].
@@
 Slope estimates less than 0 provide evidence of learning.
-_pMCMC_ was used to test the hypothesis that posterior distributions of slopes, and slope contrasts, were different from zero.
+_pMCMC_ was used to test the hypothesis that posterior distributions of slopes and slope contrasts differed from zero.
 We considered an effect statistically significant when _pMCMC_ <0.05.
 
-To explore direct and indirect links between mitochondrial function and learning, we fitted a multivariate structural equation model (SEM) using _brms_.
+We fitted a multivariate structural equation model (SEM) using _brms_ to explore direct and indirect links between mitochondrial function and learning.
 The model was structured based on theoretical expectations shown in @fig-sem.
 Any missing data was imputed during model fitting using data augmentation [@noble2021planned], but this was largely restricted to DNA damage and lipid peroxidation.
-To obtain a measure of learning performance we extracted the posterior distribution of individual learning slopes (i.e., changes in error over time) for each lizard.
-This was achieved by including a random trial slope for each lizard.
-Learning slopes, as well as the rest of variables, were standardized as before (see above).
-We assumed a gaussian error distribution for all the variables in this model.
-Factors found as non-significant in the univariate models were excluded.
+To obtain a measure of learning performance we extracted the posterior distribution of individual learning slopes (i.e., changes in error over time) by including a random trial slope for each lizard.
+Learning slopes and the rest of variables, were standardized as before (see above).
+We assumed a Gaussian error distribution for all the variables in this model.
+Factors found to be non-significant in the univariate models were excluded.
 Clutch identity remained a random factor in our SEM model.
@@
 
-We started with 80 lizards - 20 per treatment - coming from `r total_clutches` clutches.
+We started with 80 lizards - 20 per treatment - from `r total_clutches` clutches.
 However, due to natural mortality, only 79 lizards were included in the learning analyses, and 78 in the mitochondrial analyses.
@@
 #| label: fig-learning
-#| fig.cap: "Results for learning analyses.
+#| fig.cap: "Model estimates for learning analyses.
 (a) the predicted number of errors over trials.
@@
 Mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation were also not affected by prenatal conditions (see @fig-results_mit and table S2).
-We did not see any significant effects of sex and age on mitochondrial physiology either (see table S3 and tables S5-S9).
+We did not find effects of sex and age on mitochondrial physiology (see table S3 and tables S5-S9).
 
@@
 #| label: fig-results_mit
-#| fig-cap: "Estimates of mitochondrial density (a), metabolic capacity (b), ROS (c), DNA damage (d), and lipid peroxidation (e) in the medial cortex as a function of the different prenatal conditions.
-Note that for (d) and (e) these analyses do not account for missing data resulting from a flow cytometer malfunction that impacted one plate, and so, sample sizes are lower for univariate analyses.
-The x-axis represents the estimated values and in the y-axis is the density of the estimates.
+#| fig-cap: "Model estimates of mitochondrial density (a), metabolic capacity (b), ROS (c), DNA damage (d), and lipid peroxidation (e) in the medial cortex as a function of the different prenatal conditions.
+Note that for (d) and (e), these analyses do not account for missing data resulting from a flow cytometer malfunction that impacted one plate, so sample sizes are lower for univariate analyses.
+The x-axis represents the estimated values while the y-axis shows the density of the estimates.
 Points and bars represent the mean and standard error of the estimated values, respectively.
@@
 
-Our SEM showed that mitochondrial physiology was not related to learning abilities (see @fig-sem and table S4).
+Our SEM showed that mitochondrial physiology was unrelated to learning abilities (see @fig-sem and table S4).
 While we found that ROS production increased with metabolic capacity (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], _pMCMC_ `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`), ROS was not related to  oxidative damage (see @fig-sem and table S4).
@@
 Lizards were able to learn a complex spatial task with marked improvements over time in their ability to choose the correct location.
-However, learning was not influenced by prenatal conditions or linked to mitochondrial function in the medial cortex, counter to what we predicted.
+However, learning was neither influenced by prenatal conditions nor linked to mitochondrial function in the medial cortex, counter to what we predicted.
 Despite ROS increasing with greater metabolic capacity as expected, this did not result in a corresponding increase in oxidative damage in the medial cortex in _L.
@@
 #### Spatial learning is robust to prenatal conditions  
-Lizards exhibited clear evidence for learning but learning rates were not influenced by prenatal CORT or temperature as we predicted.
-This contrasts with findings in other taxa, where incubation temperature and prenatal GC’s have both been shown to affect learning and brain development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017].
+Lizards exhibited clear evidence for learning, but learning rates were not influenced by prenatal CORT or temperature as we predicted, contrasting with findings in other taxa where incubation temperature and prenatal GCs impacted learning and brain development [@lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_effects_2017; @dayananda_incubation_2017].
 Our results suggest that spatial learning abilities in _L.
-delicata_ are buffered from early environmental insults, consistent with our previous findings showing that behavioural flexibility in this species is also unaffected by prenatal CORT or temperature [@recio2025early].
+delicata_ are buffered from adverse early environmental conditions, consistent with our previous findings showing that behavioural flexibility in this species is also unaffected by prenatal CORT or temperature [@recio2025early].
 _L.
-delicata_'s resilience may stem from the adaptive value of learning in lizards.
+delicata_'s resilience may stem from the relevance of spatial cognition in lizards.
 For example, spatial learning in velvet geckos (_Amalosia lesueurii_) enhances survival in the wild [@dayananda_incubation_2017].
@@
 delicata_.
-Studies in other taxa have shown that incubation at high temperatures decrease energy production and oxidative stress [@treidel2016temperature; @zavorka_climate_2021], while elevated GCs or maternal stress is related to lower mitochondrial efficiency and higher oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
+Studies in other taxa have shown that incubation at high temperatures decreases energy production and oxidative stress [@treidel2016temperature; @zavorka_climate_2021], while elevated GCs or maternal stress is related to lower mitochondrial efficiency and higher oxidative damage [@costantini2011meta; @gong_chronic_2011; but see @crino2024eggs].
 In _L.
@@
 delicata_’s brain is robust to both incubation temperature and prenatal CORT.
-This could reflect differential tissue responses to early-life conditions, highlighting the importance of studying these effects across multiple tissues.
+Differences between liver and brain could reflect tissue-dependent responses to early-life conditions, highlighting the importance of studying these effects across multiple tissues [@casagrande2023mitochondrial].
 
 #### Variation in brain metabolic physiology is not related to learning abilities
-Mitochondria play a critical role for cognitive function by synthesizing energy [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018], and by affecting the rate of cell senescence and death caused by oxidative damage [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022].
-For example, donut-shaped mitochondria are associated with higher ROS production in presynaptic neurons and this mechanism has been shown to impair visuospatial working-memory in rhesus monkeys (_Macaca mulatta_) [@hara_presynaptic_2014].
-
-Our findings contrast with these studies and suggest that energy limitations do not impact the process of spatial learning in _L.
+Mitochondria play a critical role in cognitive function by synthesizing energy [@siegel1994basic; @du_dynamic_2009; @picard_mitochondria_2014; @picard_energetic_2018], and by affecting the rate of cell senescence and death caused by oxidative damage [@mcnay_decreases_2000; @mann_coupling_2021; @alexandrov_neuronal_2022].
+Our findings contrast with past studies and suggest that energy limitations do not impact spatial learning in _L.
 delicata_.
-The lack of association between mitochondrial density, mitochondrial capacity, and learning abilities may be the result of lizards being able to maintain energy production above the threshold required to prevent cognitive dysfunction.
-The generally low energetic demands of ectothermic organisms compared to endotherms may make this possible and explain why we found no effect when they seem to be common in mammals [@hara_presynaptic_2014; @cao2014ampk].
-
-Interestingly, despite finding a significant positive relationship between metabolic capacity and ROS in the medial cortex, the heightened ROS production did not lead to greater oxidative stress.
+The lack of association between mitochondrial density, mitochondrial capacity, and learning abilities may result from lizards being able to maintain energy production above the threshold required to prevent cognitive dysfunction.
+The generally low energetic demands of ectothermic organisms compared to endotherms may make this possible and explain why we found no effect when they seem common in mammals [@hara_presynaptic_2014; @cao2014ampk].
+Alternatively, individuals with higher respiration and ROS may be upregulating antioxidant production [@rice2002brain].
+
+Interestingly, despite finding a significant positive relationship between metabolic capacity and ROS in the medial cortex, the heightened ROS production did not lead to greater oxidative damage.
 The strong relationship between ROS and mitochondrial capacity aligns with the role of mitochondrial membrane potential in driving ATP synthesis and electron transport, processes that inherently generate ROS as byproducts [@rice2002brain].
@@
 ### Ethics
-Both the breeding animals and the experimental lizards were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate levels of humidity (see Methods 1: Animal husbandry in Supplementary Material for details).
-Euthanasia was performed by peritoneal injection of a 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
-We monitored the animals to ensure there was no irritation from the agent as indicated by distressed animals.
+Both the breeding animals and the experimental lizards were provided humane laboratory housing, with thermoregulatory opportunities, light (UV and heat) and moderate humidity levels (see Methods 1: Animal husbandry in Supplementary Material for details).
+Euthanasia was performed by intraperitoneal injection of 10 mg/kg of a 10 mg/mL alfaxan solution (a potent anesthetic) followed by decapitation.
+We monitored the animals to ensure no irritation from the agent as indicated by distressed animals.
 Before disposing of the lizard, we confirmed the absence of righting response and the pinching reflex in one of the front limbs.
@@
 ### Declaration of AI use
-We declare Chat GPT was used mainly for questions related to coding and data analyses.
-All other parts of the manuscript were written by the authors and those parts where AI was used were checked and modified accordingly by the authors.
+We declare Chat GPT-4 was used for questions related to code checking when there were unknown errors, but we did not rely on AI for the creation of any of the content in the manuscript, which was written fully by the authors.
+Last time Chat GPT-4 was used was February 2025.
 
@@
 ### Acknowledgements
-We thank Dr.
-Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
+
+We thank Mick Devoy and the Cytometry, Histology and Spatial Multiomics team for their advice and help througout the flow cytometry protocols.
 We thank the help and assistance of our lab technicians Benjamin Durant and Michelle Stephens for taking care of the lizards.
 We also thank ANU MakerSpace, where we designed and built the prototypes of the 3D printed feeders.
+The authors acknowledge Microscopy Australia (ROR: 042mm0k03) at the Centre for Advanced Microscopy, The Australian National University, a facility enabled by NCRIS and university support.
+We specifically thank Dr Angus Rae and Dr Daryl Webb for their help with microscopy protocols.
 Finally, we wish to acknowledge the anonymous reviewers for their valuable feedback on the manuscript.
@@
 #### Brain mitochondrial activity
-_Homogenates_
-After the medial cortex was extracted, it was transferred immediately to 1.5mL centrifuge tubes containing 100µL of cold 1X PBS and kept on ice until further processing.
+**Homogenates**: After the medial cortex was extracted, it was transferred immediately to 1.5mL centrifuge tubes containing 100µL of cold 1X PBS and kept on ice until further processing.
 The tissue was mechanically homogenized by placing the tissue in the well of a 100µm mesh filter (pluriStrainer) affixed atop a 1.5mL centrifuge tube, then mashed with the rubber end of an insulin syringe stopper.
@@
 Following homogenization, we centrifuged each sample at 1000 RCF for 10 minutes to pellet cells, then removed the supernatant (hereafter, this process referred to as ‘washing’) and resuspended the cells in 500µL 1x PBS.
-This step was to performed to remove cellular debris from homogenates.
+This step was performed to remove cellular debris from homogenates.
 
@@
 
-_Staining fresh samples_
-From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50µL of homogenate in duplicates (2 wells per homogenate).
+**Staining fresh samples**: From fresh homogenate suspensions, we loaded the wells of a 96-well flat-bottom plate (Nunclon) with 50µL of homogenate in duplicates (2 wells per homogenate).
 To each replicate well, we added 5µL of a fluorescent probe mix containing equal parts 5µM MitoTracker Deep Red FM, 2.5µM MitoTracker Orange CMTMRos, and 50µM MitoSOX Red.
@@
 Upon the completion of staining, samples were immediately transferred to flow cytometry facilities for data collection and were sampled within 2 hours.
-Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (unpublished data).
+Samples prepared this way remained viable for flow cytometry for approximately 5 hours post-staining at room temperature (~19°C) before cells began rapidly degrading (DCL, personal observation).
 
@@
 
-_Cryopreservation_
-Aliquots were fixed first by adding the samples into a 1mL solution of 1% Neutral-Buffered Formalin (as a fixative agent) and incubating them at 32°C for 20 min.
-Then we centrifuged the samples at 800-1200xg for 10 min to pellet cells, removed the supernatant, and resuspended the cells (from now on this process will be called as 'washing') in 1mL cold 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant).
+**Cryopreservation**: Aliquots were fixed first by adding the samples into a 1mL solution of 1% Neutral-Buffered Formalin (as a fixative agent) and incubating them at 32°C for 20 min.
+Then washed the samples and resuspended the cells in 1mL cold 1X Tris-EDTA (chelates metals that can damage DNA during freezing) and 10% DMSO (a cryoprotectant).
 Samples were stored at -20°C until oxidative damage assays.
 
-_Staining cryopreserved samples_
-Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
+**Staining cryopreserved samples**: Assays of oxidative damage from cryopreserved samples were performed within 6 months of the initial processing and analysis of fresh samples.
 On the day of oxidative damage assays, we rapidly thawed frozen samples by briefly (1-2 minutes) submerging them in hot water.
-We washed each thawed sample twice, the first time resuspending the pelleted cells in 1000µL warm 1X Tris-EDTA, and the second time in 200µL warm 1X PBS containing 5mM (6.14 mg/mL) digitonin.
+We washed each thawed sample twice, the first time resuspending the pelleted cells in 1000µL warm 1X Tris-EDTA, and the second time in 200µL warm 1X PBS containing 20µM digitonin.
 We incubated the samples at 32°C for 20 minutes to permeabilize the cell membrane, after which we washed the homogenate and resuspended the pelleted cells in 200µL 1X PBS.
@@
 Unstained and single-color controls were treated identically to samples, but stained with only up to one of BODIPY 665/676, Hoechst 33342, 8-OHdG antibody, or Alexa-Fluor 488 conjugate.
-Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugated.
+Additionally, one control was stained with both 8-OHdG antibody and the Alexa-Fluor 488 conjugate.
 We then loaded a 96-well plate with 100µL of each single-color control and 100µL in triplicate of each sample.
@@
 
-_Flow cytometry_
-All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
+**Flow cytometry**: All flow cytometry assays were performed using a flow cytometer with 5-lasers (blue, red, yellow-green, violet, and ultraviolet), 20 detectors, and a high-throughput plate reader (Becton Dickson LSRFortessa X-20) using the default wavelength filters on detectors.
 Immediately prior to all assays, we performed a quality-control check and laser alignment using the CS&T function of BD FACSDiva (v.
 8.0.1) and BD CS&T fluorescent beads (Lot No.
-30664) diluted at 2 drops to 150µL 1X PBS.
+30664) diluted at 1 drop to 150µL 1X PBS.
 During data collection, data for single-color controls was filtered using a liberal threshold of 200 on the FSC (roughly, cell size) detector, while data from samples was filtered using a threshold of 200 on the BUV-496 (Hoechst 33342) detector.
@@
 The detectors and voltage settings used in data acquisition for each assay type (mitochondrial function, oxidative damage) were determined during pilot trials prior to assays and were not changed during assays to allow for comparison among different plates and samples throughout the experiment.
-Voltages were chosen to center the distribution of observations in each channel at 103 fluorescent intensity and reduce observations of off-scale (<101 or >105) events.
+Voltages were chosen to center the distribution of observations in each channel at 10^3^ fluorescent intensity and reduce observations of off-scale (<10^1^ or >10^5^) events.
 For the mitochondrial function assay, we recorded data from the following channels (in brackets: voltage; parameter): FSC (44; forward scatter), SSC (180; side scatter), Alexa-Fluor 488 (544; autofluorescence), BUV-496 (450; Hoechst 33342), APC (647; MitoTracker Deep Red FM), PE (522; MitoTracker Orange CMTMRos), and PerCP-Cy5-5 (592; MitoSOX Red) channels.
@@
 
-_Data processing_
-In FlowJo v.
-10.1 we first applied transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (101-105) to remove any off-scale events.
+**Data processing**: In FlowJo v.
+10.1 we first transformed all fluorescent data to a logarithmic base 10 scale, then applied a basic gating process across all channels by filtering to observations within the detectable range (10^1^-10^5^) to remove any off-scale events.
 We then used a backgating process wherein we aimed to identify the approximate FSC (cell size) and SSC (cell complexity) range of viable cells that were positive for all stains.
-We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 103).
-For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 103).
-For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 103).
+We primarily used the BUV-496 channel (Hoechst 33342) in the backgating process to identify intact, nucleated cells (BUV-496 > 10^3^).
+For the mitochondrial function assay, we aimed to identify populations of viable cells containing mitochondria (APC > 103) and actively respiring (PE and PerCP-Cy5-5 > 10^3^).
+For the oxidative damage assay, we aimed to identify populations of cells exhibiting both DNA damage (Alexa-Fluor 488 > 103) and lipid peroxidation (PerCP > 10^3^).
 When backgating was done, we filtered the data to the FSC by SSC range that captured the ideal population.
@@
 ```
+
+#### Brain validation
+
+To ensure that neurons were not unintentionally lost during homogenization, we performed a pilot study where we euthanized four lizards and prepared medial cortex homogenates using the same procedures as before.
+However, to ensure we could identify neurons, homogenates here were dyed with marker that specifically targeted neuron nuclei (Farrow et al.
+2021; Storks et al.
+2023).
+We employed fluorescence microscopy and flow cytometry to check for the presence of neurons in the homogenates.
+
+Euthanasia and homogenization for each of brain region followed the procedures outlined above (see _Methods 2: Flow cytometry_).
+Before staining the samples, cells were also fixed and permeabilized as described previously (see _Methods 2: Flow cytometry_).
+Following permeabilization, we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of a 1:100 dilution of NeuN + Alexa488 fluorescent conjugate to dye the neuronal nuclei (Farrow et al.
+2021; Storks et al.
+2023).
+The samples were incubated at +4 ºC overnight.
+
+The following day we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of PBS.
+Samples were split in two: 50 µL from each was reserved for examination under a Zeiss AxioObserver Z1 microscope, while the remaining 50 µL from each sample was pooled into a single tube for flow cytometry analysis.
+.
+Flow cytometry of stained cells allowed us to validate that our gating strategy was correct.
+
+```{r, cell_size}
+#| label: cell_size
+# Getting the size of the cells in the experiment
+mit_df <- read.csv(here("./data/mit_data.csv"))
+euth <- read.csv(here("./data/Euthanasia.csv"))
+#
+data_cells <- merge(mit_df, euth, by = c("plate", "tube.no")) %>%
+  filter(type != "control", ch == "fsc") %>% 
+  dplyr::select(lizard_id, sex, ch, geo.mean) %>%
+  group_by(lizard_id, ch) %>%
+  summarize(geo.mean = mean(geo.mean, na.rm = TRUE),
+            .groups = 'drop') %>%
+  pivot_wider(
+    names_from = ch,
+    values_from = geo.mean
+  ) %>%
+data.frame()
+#
+mean_cells <- format_dec(mean(data_cells$fsc), 3)
+quant2_5_cells <- format_dec(quantile(data_cells$fsc, 0.025), 3)
+quant97_5_cells <- format_dec(quantile(data_cells$fsc, 0.975), 3)
+#
+# Getting the size of the neurons in the validation
+data_neurons <- read.csv(here("./data/neuron_size.csv")) %>%
+  filter(X == "Medial Cortex_Tube_002_003.fcs")
+neuron_size <- log10(as.numeric(data_neurons$cells.Neurons...Geometric.Mean..FSC.A.))
+```
+
+Samples examined under the microscope showed a clear presence of neuronal nuclei in the homogenates (Figure S13).
+Flow cytometry analysis also confirmed the presence of neuronal nuclei in the homogenates, that were similar to the size of the cells employed in the experiment (mean size cells experiment = `r mean_cells`; 95% CI = [`r quant2_5_cells`, `r quant97_5_cells`], n = 80; size neurons = `r neuron_size`).
+
+```{r, fig-S8}
+#| label: fig-S13
+#| fig-cap: "Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488. The image was excited using a 488 nm wavelength, and green fluorescence corresponds to NeuN-positive cells."
+#
+knitr::include_graphics("./output/figures/microscope/neurons_black_background.jpg")
+```
+
+Figure S8.
+Fluorescence microscopy image of neuron nuclei stained with NeuN-Alexa 488.
+Images were taken on a Zeiss AxioObserver Z1, equipped with Zeiss Axiocam 506 monochrome camera.
+A Zeiss 38HE fluorescent  filter set (450-490nm Ex, 500-550nm Em.) was used  in conjunction with a 63x 1.4 NA.
+
+**REFERENCES**
+
+Farrow LF, Andronicos NM, McDonald PG, Hamlin AS. 2021. Quantitative determination of neuronal size and density using flow cytometry. _Journal of Neuroscience Methods_, **352**, 109081.
+
+Storks L, Powell BJ, Leal M.
+2023.
+Peeking inside the lizard brain: neuron numbers in Anolis and its implications for cognitive performance and vertebrate brain evolution.
+_Integrative and Comparative Biology_, **63(1)**, 223-237.
+
+```{r, results='asis', echo=FALSE}
+cat("\\newpage")
+```

### 2025-04-11 .vdoc.r

@@ -1,0 +1,1349 @@
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#| label: setup
+pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes, ggh4x, cowplot, fitdistrplus, MASS, goftest, forcats, nortest, fitdistrplus, ggh4x, PupillometryR, png, grid, remotes, ggthemes, bayestestR, HDInterval, DiagrammeR, magick)
+#
+#
+#
+#| label: data_processing
+# The result will be the final df with the data for the analysis. To do so, we first estimate the learning slope (choice and errors) for each individual and then merge the data with the mitochondrial data, extracted using the script in extract.R and extraction_finc.R (see R folder). The final df will be saved in (here("output/databases_clean/data_complete.csv") 
+refit <- FALSE
+source(here("R", "data_process.R"))
+#
+#
+#
+#| label: data_metrics
+# Using the raw learning database, I calculate here some basic metrics mentioned later in the Methdos (see below).
+data <- read.csv(here("./data/Spatial_learn.csv"))
+#
+# Total number of clutches
+total_clutches <- data %>% 
+  distinct(clutch) %>% 
+  count() %>% 
+  pull(n)
+# Calculate total number of individuals
+total_individuals <- data %>% 
+  distinct(lizard_id) %>% 
+  count() %>% 
+  pull(n)
+# Calculate number of individuals per clutch
+individuals_per_clutch <- data %>%
+  distinct(lizard_id, clutch) %>%
+  group_by(clutch) %>%
+  summarize(individual_count = n())
+#
+#
+#
+#| label: sampleSize
+# List with the sample sizes from the main database.
+source(here("R", "func.R"))
+#
+hormone <- c("CORT", "Control")
+temperature <- c("Cold", "Hot")
+#
+n_list <- list()
+#
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
+    list_name <- paste0(hormone[k], "_", temperature[l])
+    n_list[[list_name]] <- sample(df = clean_df, corti = hormone[k], therm = temperature[l])
+  }
+}
+#
+df_damage <- clean_df %>%
+  filter(DNAdamage != "NA")
+n_damage <- list()
+for(k in 1:length(hormone)){
+  for(l in 1:length(temperature)){
[…]

### 2025-04-20 ms.qmd
updates

@@
 bibliography: "./bib/refs.bib"
-csl: "./bib/biology-letters.csl"
+csl: "./bib/the-journal-of-experimental-biology.csl"
 reference-doc: "./bib/tmpl.docx"
@@
 crossref:  
-  fig-title: Figure    # (default is "Figure")
+  fig-title: Fig.    # (default is "Figure")
   title-delim: —     # (default is ":")
-  fig-prefix: Figure   # (default is "Figure")
+  fig-prefix: Fig.   # (default is "Figure")
   tbl-prefix: Table  # (default is "Table")
@@
 
+## Keywords
+cognition, corticosterone, learning, mitochondria, reptiles, incubation temperature
+
 ## Abstract
@@
 Clutch identity remained a random factor in our SEM model.
-Direct, indirect, and total effects were derived from posterior estimates [@kline2005principles] (see table S4).
+Direct, indirect, and total effects were derived from posterior estimates [@kline2005principles] (see Table S4).
 
@@
 
-We found that the lizards improved their choices over time, making less mistakes as the trials progressed (see @fig-learning and table S1).
-However, the learning rate was not influenced by CORT, temperature, or their interaction (see contrasts in table S2).
-
-Mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation were also not affected by prenatal conditions (see @fig-results_mit and table S2).
-We did not find effects of sex and age on mitochondrial physiology (see table S3 and tables S5-S9).
+We found that the lizards improved their choices over time, making less mistakes as the trials progressed (see @fig-learning and Table S1).
+However, the learning rate was not influenced by CORT, temperature, or their interaction (see contrasts in Table S2).
+
+Mitochondrial density, metabolic capacity, ROS, DNA damage, and lipid peroxidation were also not affected by prenatal conditions (see @fig-results_mit and Table S2).
+We did not find effects of sex and age on mitochondrial physiology (see Table S3 and Tables S5-S9).
 
@@
 Our SEM showed that mitochondrial physiology was unrelated to learning abilities (see @fig-sem and table S4).
-While we found that ROS production increased with metabolic capacity (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], _pMCMC_ `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`), ROS was not related to  oxidative damage (see @fig-sem and table S4).
+While we found that ROS production increased with metabolic capacity (β = `r format_dec(mean(coeff_potential_ros), 3)`, 95% CI = [`r format_dec(quantile(coeff_potential_ros, 0.025), 3)`, `r format_dec(quantile(coeff_potential_ros, 0.975), 3)`], _pMCMC_ `r format_p(pmcmc(coeff_potential_ros), 3, equal = TRUE)`), ROS was not related to  oxidative damage (see @fig-sem and Table S4).
 
@@
 To ensure that neurons were not unintentionally lost during homogenization, we performed a pilot study where we euthanized four lizards and prepared medial cortex homogenates using the same procedures as before.
-However, to ensure we could identify neurons, homogenates here were dyed with marker that specifically targeted neuron nuclei (Farrow et al.
-2021; Storks et al.
-2023).
+However, to ensure we could identify neurons, homogenates here were dyed with marker that specifically targeted neuron nuclei (Farrow et al., 2021; Storks et al., 2023).
 We employed fluorescence microscopy and flow cytometry to check for the presence of neurons in the homogenates.
@@
 Before staining the samples, cells were also fixed and permeabilized as described previously (see _Methods 2: Flow cytometry_).
-Following permeabilization, we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of a 1:100 dilution of NeuN + Alexa488 fluorescent conjugate to dye the neuronal nuclei (Farrow et al.
-2021; Storks et al.
-2023).
+Following permeabilization, we centrifuged the samples at 1000 RCF for 10 minutes and resuspended the pellet in 100 µL of a 1:100 dilution of NeuN + Alexa488 fluorescent conjugate to dye the neuronal nuclei (Farrow et al., 2021; Storks et al., 2023).
 The samples were incubated at +4 ºC overnight.
@@
 
-Farrow LF, Andronicos NM, McDonald PG, Hamlin AS. 2021. Quantitative determination of neuronal size and density using flow cytometry. _Journal of Neuroscience Methods_, **352**, 109081.
-
-Storks L, Powell BJ, Leal M.
-2023.
+**Farrow, L.
+F., Andronicos, N.
+M., McDonald, P.
+G.
+and Hamlin, A.
+S.** (2021).
+Quantitative determination of neuronal size and density using flow cytometry.
+_Journal of Neuroscience Methods_, **352**, 109081.
+
+**Storks, L., Powell, B.
+J.
+and Leal, M.
+(2023)**.
 Peeking inside the lizard brain: neuron numbers in Anolis and its implications for cognitive performance and vertebrate brain evolution.

### 2025-04-20 R/data_process.R

@@ -71,8 +71,8 @@ learning_slopes <- merge(mod_learning, posteriors_errors, by = "lizard_id")
 mit_df <- read.csv(here("./data/mit_data.csv"))
 euth <- read.csv(here("./data/Euthanasia.csv"))
 mit_damage <- read.csv(here("./data/mit_damage.csv")) %>%
-  filter(startsWith(X, "Specimen_")) %>%
-  mutate(tube.dam = sapply(strsplit(X, "_"), function(x) x[4])) %>%
+  filter(startsWith(specimen, "Specimen_")) %>%
+  mutate(tube.dam = sapply(strsplit(specimen, "_"), function(x) x[4])) %>%
 data.frame()
 #
 #### B.2) Merge mito_df and mit_damage and modify the resultant df to make it clearer
